### 2017-08-31 paper/coverletter.tex
finish redrafting cover letter

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt,letterpaper]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
%\usepackage[urw-garamond]{mathdesign}
\usepackage{graphicx} % Required for including pictures
\usepackage{microtype} % Improves typography
%\usepackage{gfsdidot} % Use the GFS Didot font: http://www.tug.dk/FontCatalogue/gfsdidot/
\usepackage[T1]{fontenc} % Required for accented characters
\usepackage{url}
% Create a new command for the horizontal rule in the document which allows thickness specification
\makeatletter
\def\vhrulefill#1{\leavevmode\leaders\hrule\@height#1\hfill \kern\z@}
\makeatother

%----------------------------------------------------------------------------------------
%	DOCUMENT MARGINS
%----------------------------------------------------------------------------------------

\textwidth 6.75in
\textheight 9.25in
\oddsidemargin -.25in
\evensidemargin -.25in
\topmargin -1in
\longindentation 0.50\textwidth
\parindent 0.4in

%----------------------------------------------------------------------------------------
%	SENDER INFORMATION
%----------------------------------------------------------------------------------------

\def\Who{} % Your name
\def\What{} % Your title
\def\Autha{Michael Henry Tessler} % Your name
\def\Authb{Noah D. Goodman} % Your name
\def\Where{Department of Psychology} % Your department/institution
\def\Address{450 Serra Mall, Bldg 420} % Your address
\def\CityZip{Stanford, CA 94305} % Your city, zip code, country, etc
\def\Email{E-mail: mtessler@stanford.edu} % Your email address
\def\TEL{Phone: (757) 561-7971} % Your phone number

%----------------------------------------------------------------------------------------
%	HEADER AND FROM ADDRESS STRUCTURE
%----------------------------------------------------------------------------------------

\address{
\includegraphics[width=0.8in]{logo-simple.pdf} % Include the logo of your institution
\hspace{5.1in} % Position of the institution logo, increase to move left, decrease to move right
\vskip -0.85in~\\ % Position of the text in relation to the institution logo, increase to move down, decrease to move up
\Large\hspace{1.5in}STANFORD UNIVERSITY \hfill %~\\[0.05in] % First line of institution name, adjust hspace if your logo is wide
%\hspace{1.5in}University 
\hfill 
\normalsize % Second line of institution name, adjust hspace if your logo is wide
\makebox[0ex][r]{\bf \Who \What }\hspace{0.08in} % Print your name and title with a little whitespace to the right
~\\[-0.11in] % Reduce the whitespace above the horizontal rule
\hspace{1.5in}\vhrulefill{1pt} \\ % Horizontal rule, adjust hspace if your logo is wide and \vhrulefill for the thickness of the rule
\hspace{\fill}\parbox[t]{2.85in}{ % Create a box for your details underneath the horizontal rule on the right
%\flushright 
\footnotesize % Use a smaller font size for the details
\flushright
%\Who \\  % Your name, all text after this will be italicized
\Autha\\
\Authb\\
\Where\\ % Your department
\Address\\ % Your address
\CityZip\\ % Your city and zip code
%\TEL\\ % Your phone number
%\Email\\ % Your email address
%\URL % Your URL
}
\hspace{0in} % Horizontal position of this block, increase to move left, decrease to move right
\vspace{-1in} % Move the letter content up for a more compact look
}

%----------------------------------------------------------------------------------------
%	TO ADDRESS STRUCTURE
%----------------------------------------------------------------------------------------

\def\opening#1{\thispagestyle{empty}
{\centering\fromaddress \vspace{0.6in} \\ % Print the header and from address here, add whitespace to move date down
\hspace*{\longindentation}\hspace*{\fill}\par} % Print today's date, remove \today to not display it
%{\raggedright \toname \\ \toaddress \par} % Print the to name and address
\vspace{0.4in} % White space after the to address
\noindent #1 % Print the opening line
% Uncomment the 4 lines below to print a footnote with custom text
%\def\thefootnote{}
%\def\footnoterule{\hrule}
%\footnotetext{\hspace*{\fill}{\footnotesize\em Footnote text}}
%\def\thefootnote{\arabic{footnote}}
}

%----------------------------------------------------------------------------------------
%	SIGNATURE STRUCTURE
%----------------------------------------------------------------------------------------

\signature{\Autha \\ \Authb \What} % The signature is a combination of your name and title

\long\def\closing#1{
\vspace{0.1in} % Some whitespace after the letter content and before the signature
\noindent % Stop paragraph indentation
\hspace*{\longindentation} % Move the signature right
\parbox{\indentedwidth}{\raggedright
#1 % Print the signature text
\vskip 0.1in % Whitespace between the signature text and your name
\fromsig}} % Print your name and title

%----------------------------------------------------------------------------------------

\begin{document}

\begin{letter}

\opening{Dear Editor,}

We enclose the manuscript \emph{Communicating Generalizations} to be considered for publication in \emph{Psychological Review}. 
This paper explores the linguistic means by which generalizations about categories, events, and causality are communicated.  
We present a Bayesian model of understanding generalizations in language, and show that it predicts human judgments with high quantitative accuracy across several experiments.
We probe the theory in more detail, and find that the semantics of generic language depend in a fundamental way on subjective beliefs.

Understanding how the world works requires learning generalizations. 
The kinds of observations that would facilitate such learning, however, can be rare (e.g., \emph{figuring out that lightning strikes tall objects}) or costly (e.g., \emph{learning that certain plants are poisonous}).
We don't need to need taste every plant to figure out which ones are poisonous; we don't need to stand under objects of various heights to learn that lightning strikes tall objects.
Language allow us to communicate generalizations to each other.

The language of generalizations is often studied in what's called \emph{generic language} (e.g., ~\emph{Birds fly}).
Generic utterances are common in everyday conversation (\emph{Movie theaters are cold.}), the speech of young children (\emph{Slides aren't for grown ups!}), child-directed speech (\emph{Dogs are friendly.}), political and scientific discourse (\emph{Politicians make promises.}, \emph{Psychology experiments don't replicate.}), stereotypes (\emph{Boys are good at math.}), motivation (\emph{Big girls eat broccoli!}), and many other facets of experience.
Generalizations can be transmitted about more than just categories.
\emph{Habitual language} conveys generalizations about events (e.g., ~\emph{Mary swims after work}).
\emph{Causal language} conveys generalizations about causal events (e.g., ~\emph{Staring at the sun makes you go blind}).

Though this type of language is found everywhere, the meaning of generalizations in language is philosophically puzzling and has resisted precise formalization. 
The major issue in formalizing generalizations in language is determining what makes such statements true or false.
Consider generic utterances:
Generics are often about characteristic properties contained by a majority (e.g., \emph{Dogs bark}) but still can be felicitous based on only a minority displaying the property (e.g., \emph{Robins lay eggs}, even though only adult, female, fertile robins do). 
Indeed, \emph{Mosquitos carry malaria} sounds true, even though only 1\% of them do, highlighting that the acceptability of generics can be so flexible as to include cases where only the weakest quantifier (i.e., \emph{some}) would suffice. 
Some have taken this evidence to suggest that generics must express special mental relationships between kinds and properties, but arbitrary generalizations --- which seem only to depend on the actual statistics of the world --- can also be conveyed with generics (e.g., \emph{Ravens are bigger than toasters}).
There have been a few formal models proposed for generic language, but none that have made precise, quantitative predictions about human endorsement patterns.

Analogous phenomena can be observed in habitual and causal language.
If John smokes once a year, one might hesitate to endorse the statement \emph{John smokes}.
Yet, climbing a mountain once a year makes John really seem like he \emph{climbs mountains}.
\emph{Drinking moonshine makes you go blind} and so does \emph{staring at the sun}, but neither statement conveys precise information about the relative risk.

In this paper, we propose that the core meaning of generalizations in language is \emph{simple but underspecified}.
We adopt a formalism for the truth conditions of this language which has antecedents in formal semantics: a threshold function (e.g., the statement is true if the frequency is greater than some criterion $\theta$).
The critical novelty of our approach is that the criterion $\theta$ is left underspecified in the language, and must be resolved in context.
We instantiate this idea in a state-of-the-art probabilistic model of communication, and operationalize \emph{endorsement} (i.e., felicity or agreement with a statement) as a speaker's decision about whether or not to say the generalization to a naive listener.
We show how this simple model naturally explains \emph{flexible truth conditions} via interlocutors' beliefs about the property or event in the question.
Across our three case studies, we both \emph{measure and manipulate} prior knowledge, showing that it is strongly and causally related (via our model) to endorsement. 
Using the case study of habitual language, we go on to further interrogate the semantic scale from which generalizations are evaluated.
We find that speakers' subjective beliefs about what is likely to be the case in the future, and not the objective truth about what has been the case in the past is what drives endorsements. 
Across our case studies, we show very strong agreement between the predictions of our model with empirical human data; additionally, we do so in situations where simpler models would not suffice. 
This novel approach for generalizations in language opens up a number of interesting questions about the nature of conceptual structure and its linguistic traces, which we discuss at the end of the paper.

This work will be of strong interest to psychologists, as well as linguists, philosophers, and computer scientists interested in language and interaction.
The puzzles of generalizations in language are immediately compelling to a general audience, the more so because of their connections to cognitive development, cultural transmission of knowledge, and propagation of stereotypes.
Our solution represents the first formal model of generic language to be validated by quantitative agreement with human behavioral data.
It is not, however, uncontroversial: we champion a statistical view that has been often dismissed in the literature.
Indeed, in presentations around the world we have found this work to excite broad interest and engagement.

The following reviewers have expertise in the phenomenology of generic language, experimental studies of pragmatics, and/or the Bayesian modeling tools we use:

\begin{itemize}
\item Susan A. Gelman, Department of Psychology, University of Michigan gelman@umich.edu
\item Greg N. Carlson, Department of Linguistics, University of Rochester carlson@ling.rochester.edu
\item Ted Gibson, Department of Brain and Cognitive Sciences, MIT egibson@mit.edu
\item Gerhard J{\"a}ger, Department of General Linguistics, University of T{\"u}bingen gerhard.jaeger@uni-tuebingen.de
%\item Mark Steedman, School of Informatics, University of Edinburgh  steedman@inf.ed.ac.uk
\item Naomi Feldman, Department of Linguistics, University of Maryland nhf@umd.edu
\item Amy Perfors, Department of Psychology, University of Adelaide amy.perfors@adelaide.edu.au
\item Klinton Bicknell, Department of Linguistics, Northwestern University kbicknell@northwestern.edu
\item Michael Franke, University of T{\"u}bingen, mchfranke@gmail.com
\item Tamar Kushnir, Department of Psychology, Cornell University, tk397@cornell.edu
%\item Michael Lee, Department of Cognitive Sciences, University of California, Irvine mdlee@uci.edu
\end{itemize}

We look forward to receiving your opinion.

\closing{Sincerely yours,}

\end{letter}

\end{document}