### 2017-12-20 paper/appendix-bda.Rmd
lots of edits to appendices

```{r echo = F}
library(langcog)
library(tidyverse)
library(ggthemes)
library(jsonlite)
theme_set(theme_few())
project.path <- "../"
data.path <- "data/generics/endorsement/"
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}
```

# Appendix C: Bayesian Data Analysis

<!-- - add issues with doing BDA on prevalence and cue validity models -->
<!--   - look at scatter plots of MAP estimates of referent prevalence parameters (2 per item) for all 4 models -->
<!--   - look at scatterplots of bootstrapped cue validity with inferred cv from cue validity + prev regression model  -->

<!-- Add scatterplots of inferred parameters from prior or prevalence alone vs. joint inference model -->

In our three case studies, we compare an information-theoretic, computational model of endorsement to human endorsements of generalizations in language. 
The model has a single free parameter: the optimality parameter in Eq. \ref{eq:S1}.
Other model parameters (governing the prevalence prior $P(h)$ and the referent-prevalence $h'$) are estimated using additional behavioral data (e.g., prior elicitation tasks). 
Througout the paper, we model all sources of data (prevalence prior data, endorsement data) using a single joint-inference Bayesian data analytic model.
In this appendix, we describe this procedure in more detail for each case study.

## Modeling prevalence priors

In Case Study 1: Generic Language (Expt. 1b), we elicited the prevalence prior by asking about individual about the prevalence of features for individual categories.
We performed an analagous elicitation in Case Study 3: Causal Language (Expt. 3a).
We describe the analysis using generics as our running example, but a parallel analysis was done for causals.

Participants' responses can be thought of as *samples* from the prevalence prior distribution. 
Formally, we assume the prior data (analyzed independently for each property) was generated from one of two distributions: a distribution corresponding to those kinds with a stable causal mechanism that *could* give rise to the property ($\mathcal{D}_{stable}$) and a "transient cause" distribution corresponding to those kinds without a stable mechanism ($\mathcal{D}_{transient}$).
The "transient" distribution intuitively corresponds to accidental causes of the feature (e.g., a lion, who through some genetic mutation, reproduces by laying eggs).
We model this distribution as a Beta distribuition that heavily favors probabilities near 0: $\text{Beta}(\gamma = 0.01, \delta = 100)$.^[
  Note that we use the noncanonical mean $\gamma$ and concentration $\xi$ (or, inverse-variance) parameterization of the Beta distribution rather than the canonical shape (or pseudocount) parameterization for ease of posterior inference. The shape parameterization can be recovered using: $\alpha = \gamma \cdot \xi; \beta = (1 - \gamma) \cdot \xi$.
]
The "stable" distribution is modeled as a Beta distribution with unknown parameters $\text{Beta}(\gamma, \xi)$.^[
Because the Beta distribution is not defined at the points 0 and 1, we add $\epsilon$ to the 0 responses and round 1 to 0.99.
Similar results can be obtained by rounding 0 to 0.01. 
Alternatively, the "transient" distribution could be defined as a Delta distribution at 0, and 0 responses could remain in their raw form.
Adjusting 1 to $1- \epsilon$ leads to improper inferences for this model, as $1 - \epsilon$ is only likely under a highly right-skewed distribution; treating 1 as $1- \epsilon$ disproportionately influences the shape of $\mathcal{D}_{stable}$ relative to other responses.
This problem does not appear for 0 being adjusted to $\epsilon$ because the "transient" distribution already expects such low values.
]
Finally, we assume that these two components combine with mixture weighting $\phi$ such that the data we observe is $$P(d) = \phi\cdot \text{Beta} (d \mid \gamma, \xi) + (1 -  \phi) \cdot \text{Beta}(d \mid \gamma = 0.01, \xi = 100) $$.
We put the following priors over the latent parameters of the model:
\begin{eqnarray*}
\phi_i & \sim & \text{Uniform}(0, 1) \\
\gamma_i & \sim & \text{Uniform}(0, 1) \\
\xi_i & \sim & \text{Uniform}(0, 100)
\end{eqnarray*}
where $i$ ranges over the different properties (e.g., *lays eggs*, *carries malaria*).

To learn about the credible values of the parameters, we ran separate MCMC chains for each item, collecting 75,000 samples, removing the first 25,000 for burn-in.
To see how well the mixture model fits the prevalence prior data, we use the inferred parameters to generate new data.
The data generated from the model's posterior is called the *posterior predictive distribution* and is an important step in model criticism. 
If the model is a good representation of the data, the posterior predictive data will align with the observed experimental data. 
We construct a posterior predictive distribution by "forward sampling" the model.^[
This can be described by the following algorithm: First, flip a coin weighted by $\phi$.
If it comes up heads, we then sample from the "stable" component: $\text{Beta}(\gamma, \xi)$.
If it comes up tails, we sample from the "transient" component: $\text{Beta}(0.01, 100)$.
We do this many times using the posterior distibution to generate a distribution over predicted prevalence ratings.
]

In Case Study 2 (Expt. 2a), we asked participants about parameters of this mixture model (by having participants answer questions abouut *different kinds of people*) rather than having participants give samples (e.g., by listing their friends and rating how often their friends did certain actions).
In pilot testing, we found these different methodologies to give similar results.
We made the decision to ask more abstract information about the events to get at participants' abstract beliefs about the property rather than their knowledge of true frequencies. 
As well, by querying about abstract characters, we were able to include items that participants might not otherwise feel comfortable reporting off (e.g., how often their friends smoke marijuana).
The questions used to elicit information about the parameters are described in the main text.

## Modeling referent-prevalence

In Case Study 1, we used participants prevalence ratings for the category-of-interest in our generic sentences as the referent-prevalence that is used in the endorsement model (Eq. \ref{eq:S1}).
For a given generic sentence (e.g., "Robins lay eggs"), we took the prevalence ratings for the referent-category (e.g., the percentage of robins that lay eggs) from the prior elicitation task (Expt. 1b) and assumed those were generated from a single Beta distribution.
We used the same priors as we did for the *stable cause* component of the prevalence prior:
\begin{eqnarray*}
\gamma_i & \sim & \text{Uniform}(0, 1) \\
\xi_i & \sim & \text{Uniform}(0, 100)
\end{eqnarray*}
We took samples from the posterior predictive of this Beta distribution (i.e., reconstructed prevalence ratings) as the *referent-prevalence* used in the model.

In Expt. 2b (Habitual endorsement), we used the frequency given to participants in the experimental prompt (e.g., 3 times in the past week) as the referent-prevalence.
In Expt. 2c ("What is prevalence?"), we compared two endorsement models that differed in their representation of referent-prevalence. For one model (past frequency model), it was the frequency given to participants in the experimental prompt (same as in Expt. 2b); for the other model (predictive frequency model), we used the mean elicited frequency from the *predictive frequency* condition (see main text).
In Case Study 3 (Causal endorsement), we used the proportion of successful causal events  given to participants in the experimental prompt (e.g., 70 out of 100 uses of Herb C made animals sleepy).

## Jointly modeling referent-prevalence, prevalence priors, and generic endorsements

To fit the generic endorsement models, we incorporate them into the Bayesian data analytic model of the prevalence prior data (described above) to create a single, joint-inference model where the optimality parameter $\lambda_1$ (Eq. \ref{eq:S1}) is inferred jointly with all the other latent parameters of the full model (the referent-prevalence $h'_{k, f}$ for each category $k$ and property $f$ and the parameters of the prevalence priors $P_f(h)$ for each property $f$) using data from Expt. 1a \& b (Figure \ref{fig:genericsModelDiagram}).
For the parameters of the prevalence prior distributions, we use the same priors described in Expt. 1b; for the speaker optimality parameter, we use a prior with a range consistent with previous literature using the same model class: $\lambda_1 \sim \text{Uniform}(0,5)$.
We learn about the \emph{a posteriori} credible values of the joint inference models by collecting samples from 3 MCMC chains of 100,000 iterations removing the first 50,000 iterations for burn-in, using an incrementalized version of the Metropolis-Hastings algorithm [@Ritchie2016].

```{r genericsModelDiagram, fig.cap="Graphical model corresponding to the fully Bayesian data analysis of the RSA model. The feature-probability prior data $d^{pr}_{f}$ is assumed to be generated from the mixture model validated in Expt. 1a. The referent-category feature-probability $d^{pr}_{k,f}$ is generated from a latent feature-probability $h_{k,f}$. The fixed-threshold (control model) or uncertain-threshold RSA model takes the feature-probability prior $P_f(h)$, the referent-category feature-probability $h$, and a single speaker optimality parameter $\\lambda$ as input. This overall structured is repeated (except $\\lambda$) for each of the unique properties $f$ and categories $k$ that correspond to the generic sentences in our stimulus set. Note that RSA corresponds to a probabilistic *function* and not a *random variable* that is standard in graphical model notation; RSA cannot be represented by a graphical model because it has recursion. "}
knitr::include_graphics("figs/generics_bayesnetmodel.pdf", dpi = 108)
```

In addition to examining the posterior predictive distribution on endorsement judgments (presented in main text), we examined the marginal posteriors on parameters of the prevalence priors and referent-prevalence.
These marginal distributions are important to examine to confirm that they have not changed significantly from the parameters inferred from their private data sources in isolation.
For example, when modeling the referent-prevalence data in isolation, the model infers that roughly 50\% of robins lay eggs, as that is what participants tend to say in the prevalence elicitation task.^[
  Some participants do report that 100\% of robins lay eggs so the average prevalence rating is closer to 65\%..
].
The question for this joint inference model is whether or not it still believes that roughly 50\% of robins lay eggs.
We found that the referent-prevalences and prevalence priors inferred under the joint model were almost indistinguishable from those inferred using only their data, suggesting that the modeling all three data sources together does not come at the cost of modeling any one of them accurately.