### 2018-04-10 paper/case3-causals.Rmd
refresh casauls figure, refactor causals code to analysis script, save cached results

# Case Study 3: Causal Language

```{r load_causals_results}
load(file = "cached_results/case3_endorsementDataModel.RData")
#md.cas.fullmodel.endorsement,
#      md.caus.endorse, 
#      n.subj.cas.endorse, 
#      ave.minutes.cas.endorse,
#      n.subj.cas.endorse.failed,

load(file = "cached_results/case3_priorRatings.RData")
# d.cas.priors, 
   #  n.subj.cas,
  #   n.subj.cas.failed,
 #    ave.minutes.cas,
```

<!-- delta_p = presence-absence -->
<!-- cp = delta_p / normalization ... subset of relevant cases -->
<!-- caus_supp ~ structure induction problem. -->
<!--   - get to delta_p & cp assuming a different integration fn's -->

<!-- caus_support = "" -->
<!-- caus_strength = "This virus causes the disease." -->

Language about causal relationships manifests in generalization.
The utterance "Fire causes smoke" relates to "This fire caused this smoke" in a way analogous to how "John runs" relates to  "John ran yesterday". 
<!-- \ndg{"my grandpa's..." is itself a generic... is there a simple example to start with? e.g. "fire causes smoke" generalizes "this fire caused this smoke"? Also, somewhere (probably GD) we should discuss how many sentences have two ore more "dimensions of generalization".} -->
We explore this hypothesis in our third case study: causal language or *causals* (e.g., "A causes B").

The problem of *causal induction*---knowing that one thing causes another---has been studied extensively in human psychology  [@Cheng1992; @Cheng1997; @Griffiths2005; @Griffiths2009].
Classically, this is cast as a problem of inducing an unobservable relation (*type causation*) from observable events or contingency data. 
We take a different approaching, examining *type causation* by the language used to describe it (e.g., "A causes B").
We explore the idea that such language conveys a generalization about *token* or *actual causation* (e.g., "A caused B, in this instance") and that our theory of the language of generalizations extends in a natural way to describe *causal language*.
Ascribing causation to an individual event (*token or actual causation*) is itself a complex, inferential process [e.g., depending on counterfactual reasoning; @Gerstenberg2015how], which we do not try to model here.

In this paper, we posit that prevalence priors are a mediating representation between abstract conceptual structure and generalizations in language.
In this last set of experiments, we explicitly test the relationship between the prevalence priors and endorsements of generalizations about causes by manipulating the priors.

<!--Also, by designing the experiments in the domain of causal language, we further demonstrate the generality of our theory of communicating generalizations.-->

<!-- Going from covariation to causation, as in the classic causal reasoning paradigms, then, actual involves two inferences  -->

<!-- implies that during the events observed (the covariation), causal attribution was withheld: The events were not seen as causal at the time, only in retrospect. -->

<!-- A first pass could be *there exists a causal relationship between blindness and moonshine such that blindness causes moonshine*. -->
<!-- Interestingly, this definition doesn't mention the increased risk in becoming blind from moonshine, but presumably it's more probable than the alternative (not drinking moonshine). -->
<!-- We hypothesize that the semantics of the above statement (a *causal* statement) is underspecified and subject to the same context-sensitivity as generic and habitual language. -->

<!-- We take language that conveys a causal relationship between two classes of things as our third case study: causal language, or *causals*. -->
<!-- For simplicity, we identify the relevant dimension of the scale to be the conditional probability of the effect given the cause, though this representation will have problems for classic cases in causal reasoning.  -->
<!-- Our goal is not to explain the phenomena of causal reasoning *per se*, but rather to adopt a semantics that could apply to the causal domain and test our theory on the language of causal relations. -->
<!-- In particular, we aim to demonstrate empirically that our theory of the language of generalization posits constructs that are causally related to sentence endorsement. -->

<!-- If we see individual events as causal (*token level* causation), then we may reach a point when we organize those events as instances of a category-level causal relation (*type level* causation). -->
<!-- How do we tell others about what we learned? -->
<!-- We use language of the form "C causes E". -->

<!-- Recently, causal attribution of a single event (*token causation*) -->
<!-- Adults and children can infer causal relationships from just a few examples [@Gopnik2000; @Schulz2004] and computational models have been used to formalize the problem of *elemental causal induction*, or learning a causal relationship from observational data [@Griffiths2005; @Kemp2010a]. -->
<!-- Having induced a causal relationship from observational data, people communicate their causal knowledge using language.  -->
<!-- Indeed, when causal relations are nonobvious, causal language helps children as young as 2-years learn those relations [@Bonawitz2010]. -->

## Experiment 3a: Manipulating prevalence priors

In this experiment, we manipulate participants' background knowledge, measuring these beliefs in order to check whether the manipulation was successful. 
Experiment 3b (*causal endorsement*) will then use a very similar experimental procedure in exploring causal generalization language.

<!-- examine causal domains that vary as to whether or not probabilistic causes are plausible and whether or not a background cause of the event is intuitively present. -->
<!-- We introduce participants to the results of several experiments about each domain and see how different experimental results update participants' beliefs about causal power in these domains. -->

<!-- People have theories about how different casusal systems work.  -->
<!-- Physical causation tends to have a deterministic bend to it: If one billiard ball hits a second with a certain force, the second will respond by moving with a certain force.  -->
<!-- If the experiment is repeated exactly, we would expect the same results. -->
<!-- In other domains, probabilistic causes are plausible: Giving a sick animal a certain medication may sometimes make it better and other times it won't.  -->
<!-- Finally, some events could have multiple possible causes: Testing whether or not a drug makes an animal *blink* might be difficult because animals will tend to *blink* even without intervention.  -->

### Method

#### Participants

We recruited `r n.subj.cas` participants from Amazon's Mechanical Turk.
Participants were restricted to those with U.S. IP addresses and who had at least a 95\% work approval rating.
The experiment took on average `r ave.minutes.cas` minutes and participants were compensated \$0.50 for their work.

<!-- The different kinds of system are described in more detail below. -->
<!-- Participants are then told that the team of scientists is conducting experiments with different possible causes (e.g., different kinds of foods). -->

#### Materials

Participants were told a story of a scientific experiment testing different substances to produce an effect (either to make animals sleepy or make plants grow tall). 
Our cover stories were constructed so that the potential cause could have some plausible intuitive mechanism that could give rise to the property (e.g., a naturally occurring herb causing animals to be sleepy). 
The two cover stories can be seen in Table 2.

Participants were then shown "previous experimental results", which followed one of four distributions.
In two of the conditions, participants saw results that came from a single underlying distribution (*common* conditions).
In one of these conditions, all causes produced a strong effect (average efficacy approximately 98\%; the *common strong* condition).
In the second of these conditions, all causes produced a weak effect (average efficacy approximately 20\%; the *common weak* condition).
<!-- The results of the previous experiments were described verbally  -->
<!-- These distributions had the feeling that either there was some background cause which gave rise to the effect (e.g., a simple base rate) or that all substances were equally as efficacious at producing the effect. -->
The two other conditions used distributions in which some experiments resulted in either no or very few successes (i.e., produced 0, 1, or 2 successes), and others that either had strong or weak effects as above. 
These are the *rare strong* and *rare weak* conditions.
<!-- These distributions had the feeling that the effect was somewhat rare and that only certain substances produced any effect. -->

<!-- \begin{table}[] -->
<!-- \centering -->
<!-- \label{tab:causalMaterials} -->
<!-- \resizebox{\textwidth}{!}{ -->
<!-- \begin{tabular}{|p{3cm}|p{6.5cm}|p{6.5cm}|} -->
<!-- \hline -->
<!--                    & Plants                                                                                                                                                                                                  & Animals                                                                                                                                                                                                                               \\ \hline -->
<!-- Cover story        & On this planet, there is a plant called feps and your team wants to figure out how to make these plants grow tall. Your team runs experiments trying to make feps grow tall with different fertilizers. & On this planet, there are animals called cheebas and your team of scientists wants to figure out how to make these animals sleepy. Your team runs experiments trying to make cheebas sleepy with different naturally occurring herbs. \\ \hline -->
<!-- Evidence statement & Your team gave fertilizer B to 100 different feps. Of those 100 treated, 2 feps grew tall.                                                                                                              & Your team gave herb C to 100 different cheebas. Of those 100 treated, 98 cheebas were made sleepy.    \\   \hline                                                                                                                             -->
<!-- \end{tabular} -->
<!-- } -->
<!-- \caption{Cover stories and evidence statements for the two sets of materials used in Expt. 3} -->
<!-- \end{table} -->

```{r causalMaterials, results="asis"}
causalMaterials <- xtable::xtable(data.frame(
  "." = c("Cover story", "Evidence statement"),
  Plants = c("On this planet, there is a plant called feps and your team wants to figure out how to make these plants grow tall. Your team runs experiments trying to make feps grow tall with different fertilizers.",
               "Your team gave fertilizer B to 100 different feps. Of those 100 treated, 2 feps grew tall."),
  Animals = c("On this planet, there are animals called cheebas and your team of scientists wants to figure out how to make these animals sleepy. Your team runs experiments trying to make cheebas sleepy with different naturally occurring herbs.",
               "Your team gave herb C to 100 different cheebas. Of those 100 treated, 98 cheebas were made sleepy.")
), align=c('p{1in}',' |p{1.75in}|', 'p{2in}|','p{2in} |'),
  caption = "Cover stories and example evidence statements for the two sets of materials used in Expt. 3")
bold <- function(x){
paste0('{\\bfseries ', x, '}') }

print(causalMaterials,
    type="latex", 
      comment = F,
      hline.after=-1:nrow(causalMaterials),
      size="\\fontsize{9pt}{10pt}\\selectfont", 
      include.rownames=FALSE,
      sanitize.colnames.function = bold
      )
```

```{r causalExpt, fig.cap="Overview of Experiment 3. A-C: Results of previous experiments are shown one at a time, described in text and displayed in a table. One of the results was lost. Participants are asked to review previous results once all displayed. B: Prior elicitation task: Participants predict the results of the next 5 experiments. D: Causal endorsement task: Results of previously lost experiment are found and participants are asked to evaluate the causal generalization.", fig.width = 10, cache=F}
caus.exp.1 <- ggdraw() + draw_image("figs/causal-expt-1.jpeg", scale = 1) +
  theme(plot.margin = unit(c(2,0,0,0), "pt")) 
caus.exp.2 <- ggdraw() + draw_image("figs/causal-expt-2.jpeg", scale = 1)+
  theme(plot.margin = unit(c(2,0,0,0), "pt")) 
caus.exp.3 <- ggdraw() + draw_image("figs/causal-expt-3.jpeg", scale = 1)+
  theme(plot.margin = unit(c(2,0,0,0), "pt")) 
caus.exp.4 <- ggdraw() + draw_image("figs/causal-expt-prior.jpeg", scale = 1)
caus.exp.5 <- ggdraw() + draw_image("figs/causal-expt-endorse.jpeg", scale = 1)

caus.exp.toprow <- plot_grid(caus.exp.1, caus.exp.2, caus.exp.3, labels = c("A", "B", "C"), nrow = 1, align = 'vh')

title <- ggdraw() + draw_label("Prevalence prior manipulation")
plot_grid(
  plot_grid(title, caus.exp.toprow, ncol=1, rel_heights=c(0.1, 1)),
  plot_grid(add_sub(caus.exp.4, "Prior elicitation"), add_sub(caus.exp.5,"Causal endorsement"), labels = c("D", "E"), nrow = 1),
  nrow = 2)
```

#### Procedure

The experiment was a single trial: Each participant saw only one cover story with one distribution of previous experiments.
<!-- In the first part of the trial, participants learned about different substances' ability to produce an effect ("previous experimental results"). -->
<!-- Participants read a cover story and clicked a button to show the results of successive previous experiments. -->
<!-- For each previous experiment, participants were presented with an *evidence statement* as well as a numerical representation of the results shown in a table above the text. -->
<!-- The table was filled in gradually with each successive experiment.  -->
<!-- The results of one experiment was said to be lost, and a "?" was placed in the table (Figure \@ref(fig:causalExpt)).  -->
<!-- (These lost results would be found in the *causal endorsement* task, Expt. 4b). -->
<!-- Participants were instructed to take a final look at the results of all of the previous experiments before continuing.  -->
<!-- Once participants clicked the "continue" button, they were brought to a new screen where they would make predictions about new substances that have yet to be tested (Figure \@ref(fig:causalExpt)). -->
<!-- Participants saw a list of new substances and slider bars that ranged from 0 to 100.  -->
<!-- The second phase serves as a measurement of participants' feature-probability priors. -->
Participants were told that they were an astronaut-scientist on a distant planet trying to figure out how some system works (i.e., how to make a certain kind of animal sleepy with different herbs or how to make a plant grow tall with different fertilizers). The story for the "sleepy animals" condition read:

\begin{quotation}
You are an astronaut-scientist exploring a distant planet. 
On this planet, there are animals called cheebas and your team of scientists wants to figure out how to make these animals sleepy.
Your team runs experiments trying to make cheebas sleepy with different naturally occurring herbs.
The results are shown below:
\end{quotation}

Participants then click a button to show the results of the experiments.
Experiment results appear one at a time (upon a click), and are described in an *evidence statement* (e.g., "Your team gave herb A to 100 different cheebas. Of those 100 treated, 98 cheebas were made sleepy.") as well as displayed in a table showing the number of successes (e.g., "animals made sleepy") per number of attempts (always 100 per experiment; Figure \@ref(fig:causalExpt)A). 
We described the results of these individual experiments using token-level causal language (e.g., "98 cheebas were made sleepy") to imply that actual causation occurred in these cases.

Participants see the results of eleven experiments, though they are told the results of one experiment were lost and a "?" was placed in the table (Figure \@ref(fig:causalExpt)B). 
(These lost results would be found in the *causal endorsement* task, Expt. 3b).
After participants viewed the results of the 10 "experiments" (and 1 missing experiment), they are told to review the results of the experiments before continuing (Figure \@ref(fig:causalExpt)C).

Upon clicking the continue button, the table of experiment results is removed and participants are told more experiments were conducted.
Participants were asked to predict the results of the next 5 experiments (Figure \@ref(fig:causalExpt)D).
Participants were given five slider bars ranging from 0 - 100, and asked to predict the next five substances (e.g., herbs M, N, P, Q, and R).

After responding, participants then completed an attention check survey where they were asked what the team of scientists were investigating (choosing a response from a drop-down menu with 12 options) and to input one of the numerical results they saw on the previous screen. 
This attention check served to confirm that participants had encoded both relevant aspects of the experiment (the domain and the frequencies).

<!-- #### Materials  -->

<!-- The kind of causal events in question were selected to correspond roughly to different positions in a theoretically meaningful space defined by the model. -->
<!-- Research in elemental causal induction suggests people use causal models that correspond to a mixture probability distributions, where one component of the distribution is a consequence to the intrinsic causal force while the other component is a result of some extrinsic background cause. -->
<!-- So we designed materials that corresponded to domains where various parameters of that mixture model could be manipulated.  -->

<!-- A second dimension of man ipulation corresponds to the extent to which *determinism* is expected in the domain.  -->
<!-- For example, people's theories of physical causation tends to be more deterministic (i.e., favoring probabilities 1 and 0) then causation in the social or psychological domain.  -->

### Results

`r n.subj.cas.failed` participants were excluded from the analysis for failing to answer both of the attention check questions correctly, leaving a total of `r n.subj.cas-n.subj.cas.failed` responses for analysis.
The distributions that resulted from participants predicting the causal efficacy of the new substances are shown in Figure\ \@ref(fig:figure-causals)A.
<!-- \ndg{its unclear in figure caption whether those are emprical, predicted, or bda priors.} -->
As is visually apparent, the empirical prevalence distributions differ between conditions and nicely recapitulate the distributions supplied in the different experimental conditions, suggesting that the manipulation does indeed change participants' representations of what probabilities are likely to occur in each experimental condition. 
This diversity is important because the model of generalizations predicts differences in endorsement---for the same referent prevalence---depending on these priors.
<!-- \ndg{it'd be good to have some kid of stats here. maybe just show that empirical prealence distributions differ between conditions. also say how this matters: as before geeralization model predicts differences in endorsement for same target prevalence depending on priors.} -->
<!-- We hypothesize that with the distributions altered, endorsements of generalizations will similarly be affected. -->

## Experiment 3b: Causal Endorsements

In this experiment, we tested whether the manipulated priors of Expt. 3a are causally related to the endorsement of causal statements.
Most of the experimental design was identical to that of Experiment 3a.

### Method

#### Participants

We recruited `r n.subj.cas.endorse` participants from Amazon's Mechanical Turk.
Participants were restricted to those with U.S. IP addresses and who had at least a 95\% work approval rating.
None of the participants had participated in Experiment 3a. 
The experiment consisted of one trial and took on average `r ave.minutes.cas.endorse` minutes; participants were compensated \$0.25 for their work.

#### Procedure and materials

The materials were the same as in Expt. 3a.
The first part of the experimental trial was the same as in Expt. 3a (the table of "previous experiments"; Figure \@ref(fig:causalExpt)A-C).
Upon continuing beyond the first part of the trial, the table of results and background story were removed from the screen and the participant is told that the results of the "lost experiment" were found. 
The results are reported to the participant in terms of how many out of 100 of the attempts were successful. 
Participants saw one of two reported frequencies: 20\% or 70\% (randomized between-subjects).
Participants were then asked to judge the causal sentence (e.g., "Herb X makes the animals sleepy") by either clicking "Yes" or "No" (Figure \@ref(fig:causalExpt)E). 
This allows us to test whether endorsement of a causal sentence for a given actual frequency is affected by the causal priors induced by our manipulation.
After responding, participants completed the same attention check as Expt. 3a.

```{r}

n.caus.endorse <- length(md.caus.endorse$MAP_h)
r2.caus.rsa <- compute_r2(md.caus.endorse, "MAP_h", "MAP")
mse.caus.rsa <- compute_mse(md.caus.endorse, "MAP_h", "MAP")

```

```{r figure-causals-priors}

fig.causals.priors.data <- d.cas.priors %>%
    select(distribution, response) %>%
    rename(prevalence = response) %>%
    mutate(distribution = factor(distribution, 
           levels = c("rare_weak", "common_weak", "rare_deterministic","common_deterministic"),
           labels = c("Weak, Rare", "Weak, Common","Strong, Rare","Strong, Common"))) %>%
ggplot(.,
       aes( x = prevalence , fill = distribution))+
  geom_histogram(aes(
    y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]
    ),
    bins = 20, color = 'black') + 
    scale_fill_manual(values = c("#859900","#d33682","#dc322f", "#268bd2"))+

  #geom_bar(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) + 
  #geom_density(size = 0.4, aes( y = ..scaled.. ))+
  #scale_fill_solarized()+
  #scale_linetype_manual(values = c(2, 1))+
  facet_wrap(~distribution, scales = 'free', nrow = 2)+
  #scale_y_continuous(limits = c(0, 1), breaks = c(0, 1))+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  guides(fill = F)+
  ylab("Proportion of responses")+
  xlab("Prevalence rating")
```

```{r figure-causals-endorsement, fig.width=7, fig.height=3.5, cache = F}
figure.causals.endorsement <- md.cas.fullmodel.endorsement %>%
  mutate(
    src = factor(src, levels = c( "model", "data")),
    dist = factor(dist, 
           levels = c("rare_weak", "common_weak", "rare_deterministic","common_deterministic"),
           labels = c(
                       "Weak, Rare", "Weak, Common","Strong, Rare","Strong, Common")),
    frequency = factor(frequency, 
                       levels = c(20, 70),
                       labels = c("Referent prevalence = 0.2",
                                  "Referent prevalence = 0.7")
    )) %>%
  ggplot(., aes( x = src, y = MAP, ymin = cred_lower, ymax = cred_upper, 
                 fill = dist #linetype = dist, 
                 #shape = dist,
                #        group = dist#, 
                 #alpha = src
                 ))+
  geom_bar(stat='identity', position = position_dodge(0.9), width = 0.9, color = 'black')+
  geom_errorbar(position = position_dodge(0.9), width = 0.3,alpha = 0.8)+
  #geom_linerange(position = position_dodge(), alpha=0.8)+
  #geom_line(position = position_dodge(8), alpha = 0.8, linetype = 3)+
  #geom_point(position = position_dodge(8), size = 2.25)+
  #theme(axis.text.x = element_text(angle = 90))+
  #geom_hline(yintercept = 0.5, lty = 2, color = 'black', alpha = 0.4)+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  scale_alpha_manual(values = c(1, 0.3))+
  #scale_x_continuous(limits = c(10, 90), breaks = c(20, 70))+
  #scale_fill_solarized()+
  scale_fill_manual(values = c("#859900","#d33682","#dc322f", "#268bd2"))+
  xlab("")+
  ylab("Causal Endorsement")+
  facet_grid(.~frequency)+
  #coord_flip()+
  #coord_fixed(ratio = 90)+
  #guides(fill = F)+
  guides(
    #shape=guide_legend(title="Background distribution"),
    fill=guide_legend(title="Background distribution")
    )+
  theme(
    #plot.margin = unit(c(0,6,0,6), "pt")
    #legend.position = 'bottom', 
    # legend.direction = 'horizontal', 
    # legend.title = element_blank()
  )
```

```{r figure-causals, fig.width=9, cache = F, fig.asp = 0.4, fig.cap="Endorsing generalizations about causes. A: Empirical prevalence prior distributions elicited following prior manipulation cover story (Expt. 3a). B: Endorsement model predictions (left) and human elicited endorsements (right) for four manipulated prevalence prior distributions and two referent prevalence levels."}
prow <- plot_grid( 
  fig.causals.priors.data,# + theme(legend.position="none"),
  figure.causals.endorsement + theme(legend.position="none"),
   align = 'v',
   labels = c("A", "B"),
   #hjust = -1,
   nrow = 1,
  rel_widths = c(0.8, 1)
  #rel_heights = c(1, 1)
  )
#legend <- get_legend(causals.endorsement.insets)

#plot_grid( prow, legend, rel_widths = c(3, .3))
prow
#grid.arrange(figure.causals.endorsement, causals.endorsement.insets, ncol = 1,
             #layout_matrix = cbind(c(1,1,2)))

# ,
#              habituals.endorsement.insets, ncol = 2,
#              layout_matrix = cbind(c(1,1,3), c(2,2,3)))
```

### Results

`r n.subj.cas.endorse.failed` participants were excluded from the analysis for failing to answer both of the attention check questions correctly, leaving a total of `r n.subj.cas.endorse-n.subj.cas.endorse.failed` responses for analysis.
As in our other analyses of endorsement responses, we computed the Bayesian Maximum A-Posteriori (MAP) estimate and 95\% highest probability density interval of the true population probability of endorsing the statement, assuming a uniform prior. 
These are shown for the different experimentally-manipulated priors and referent prevalences in Figure\ \@ref(fig:figure-causals)B.

\ndg{i'm not sure 10A is the best way to show this data, since the lines highlight the differences between causal powers (within conditions), while we care most about differences between conditions.}

As predicted by our model, endorsements for a causal statement were sensitive to the causal power of the target cause and, critically, to the background distribution of other causes. 
When many other causes produced the effect very reliably (*common strong* condition), very few participants endorsed the causal statement for a cause with causal power of 0.2, and were at chance when the causal power was 0.7 (Figure\ \@ref(fig:figure-causals)A).
By contrast, when many other causes failed to produce the effect and those that did were not very reliable (*rare weak* condition; green in figure), at least half of participants endorsed the causal statement for a cause with causal power of 0.2, and were at ceiling when the causal power was 0.7.
The other two conditions (*rare deterministic* and *common weak*) led to endorsements intermediate between these two conditions.
Our model predicted these effects with strong quantitative accuracy ($r^2(`r n.caus.endorse`) = `r r2.caus.rsa`$; MSE = $`r mse.caus.rsa`$).

\ndg{is it worth saying anything about the fact that the effect is even a bit bigger than predicted by the model?}

## Discussion

In our third case study, we applied our model to generalizations about causal events, without any changes.
In this domain, we successfully manipulated participants' beliefs about the expected prevalence of a causal relationship in a domain (Expt. 3a).
This was done using both unimodal (*common weak*, *common strong*) and bimodal (*rare weak*, *rare strong*) distributions. 
In Expt. 3b, we showed that these manipulated priors influenced endorsements of the corresponding causal statements.
In addition to further demonstrating the generality of this theory, these experiments show that the prevalence prior $P(p)$ is causally related to endorsements of generalizations in language.
To our knowledge, these are the first experiments to demonstrate how the prevalence of a feature among other categories can influence endorsing a generalization about a referent category. \ndg{fix last sentence. is this about geeralization? didn't our last experiments show this? is it about causation? huh?}

In these experiments, we used two cover stories that described plausible causal events: herbs making animals sleepy and fertilizers making plants grow tall.
\ndg{which reminds me: need to say something above (3a) about (no) differences between the cover stories?}
We chose these items because there was a plausible causal mechanism that could give rise to the property and these causal events could have ambiguous causal power associated with them (e.g., it is plausible that there are herbs that only weakly make animals sleepy and it is also plausible that there are herbs that almost deterministically make animals sleepy).
These two features of the domains make them particularly amenable to manipulation.
<!-- Inituive theories about these domains are flexible enough to permit such manipulations. -->

It is likely that other domain knowledge would interact with the experimentally-supplied "experimental data" to form a hybrid belief distribution.^[
If this were happening in our domains, we would expect this to show up in the results of Expt. 3a. Participants' predictions about the likely causal power of new causes would be expected to show a mixture of their abstract, intuitive theories and the experimentally supplied data.
]
For example, physical causal systems (e.g., billiard balls hitting each other) could strongly induce near-deterministic notions of causal power, analogous to our *strong* priors conditions. 
Causal systems that demonstrate surprising or *a priori* unlikely effects (e.g., liquids melting concrete) could induce rarity about the existence of a non-zero causal power, analogous to our *rare* prior conditions. 
Our theory would predict that differences in endorsement in these cases would be mediated by differences in the corresponding prevalence prior distributions.