### 2018-10-08 analysis/case1-generics.Rmd
update model results and analyses

---
title: "Analysis for Case 1 (generics)"
output: html_notebook
---

```{r libraries, cache = F}
library(papaja)
library(formatR)
library(rwebppl)
library(xtable)
library(tidyverse)
library(forcats)
library(langcog)
library(coda)
library(ggthemes)
library(ggrepel)
library(jsonlite)
library(gridExtra)
library(lme4)
library(knitr)
library(kableExtra)
library(cowplot)
library(magick)
library(viridis)
library(tidyboot)
theme_set(theme_few())
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
logmeanexp <- function(x){
  x.num <- as.numeric(x)
  xstar = max(x.num)
  return(xstar + log(mean(exp(x.num - xstar))))
}

compute_r2 <- function(df,v1, v2, sigfigs = 3){
  return(format(cor(df[[v1]], df[[v2]])^2, digits = sigfigs))
}

compute_mse <- function(df, v1, v2, sigfigs = 3){
  return(format(mean( (df[[v1]]-df[[v2]])^2), digits = sigfigs))
}

project.path <- "../"
options("scipen"=10) 
```

# Experiment 1a: Generic endorsements

```{r eval = F}
#d1 <- read.csv(paste(project.path, "data/generics/endorsement/",
                    #"truth-judgments-n100.csv", sep = ""))
#d2 <- read.csv(paste(project.path, "data/generics/endorsement/",
              #      "naturalGenerics-trials-formatted.csv", sep = ""))

# write.csv(
#   left_join(
#   d1 %>%
#     mutate(sentence = as.character(sentence),
#            sentence = gsub('&quotechar', '', sentence),
#          sentence = gsub('lyme', 'Lyme', sentence)),
#   unique(d2 %>%
#     mutate(sentence = paste(Category, " ", Property, ".", sep = "")) %>%
#     select(Category, Property, sentence))
# ), file = paste(project.path, "data/generics/endorsement/",
#                     "truth-judgments-n100.csv", sep = ""), row.names = F)
```

```{r generic-endorsement}
d.gen.endorse.catch <- read_csv(paste(project.path, "data/generics/endorsement/",
                    "truth-judgments_catch-trials.csv", sep = ""))

d.gen.endorse <- read_csv(paste(project.path, "data/generics/endorsement/",
                    "truth-judgments-n100.csv", sep = ""))

d.gen.endorse.summary <- left_join(
  d.gen.endorse,
  d.gen.endorse.catch %>% select(workerid, pass)
  ) %>%
  filter(pass == 1) %>%
  rowwise() %>%
  mutate(response = ifelse(response == "agree-key", 1, 0)) %>%
  ungroup() %>%
  group_by(sentence) %>%
  tidyboot_mean(column = response) %>%
  ungroup() %>%
  mutate(sentence = factor(sentence, levels = sentence[order(mean)]))

d.gen.endorse.bayes <- left_join(
  d.gen.endorse, 
  d.gen.endorse.catch %>% select(workerid, pass)
  ) %>%
  filter(pass == 1) %>%
  rowwise() %>%
  mutate(response = ifelse(response == "agree-key", 1, 0),
         sentence = gsub('&quotechar', '', sentence),
         sentence = gsub('lyme', 'Lyme', sentence)) %>%
  group_by(sentence, Property, Category) %>%
  summarize(k = sum(response), n = n()) %>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         low  = qbeta(.025, a, b),
         high = qbeta(.975, a, b),
         MAP_h = (a-1)/(a+b-2),
         mean = a / (a + b)) %>% 
  ungroup()

d.prev <- read.csv(paste(project.path, "data/generics/endorsement/",
                       "naturalGenerics-prior-trials-n57.csv", sep = ""))

d.prev.summary <- d.prev %>%
  mutate(item = paste(Category, " ", Property, ".", sep = "")) %>%
  filter(item %in% d.gen.endorse.bayes$sentence) %>%
  mutate(prevalence = prevalence / 100) %>%
  group_by(Category, Property, item) %>%
  tidyboot_mean(column = prevalence)

d.gen.endorse.bayes.prev <- left_join(
  d.gen.endorse.bayes, 
  d.prev.summary %>% 
    rename(sentence = item, prev = mean, prev_low = ci_lower, prev_high = ci_upper) %>%
    select(-n, -empirical_stat)
) %>%
  mutate(sentence = factor(sentence, levels = sentence[order(MAP_h)]))

# generics.endorsement.spectrum <- ggplot(
#     d.gen.endorse.bayes.prev, 
#     aes(x = sentence, y = MAP_h - 0.5, ymin = low - 0.5, ymax = high - 0.5, fill = prev)
#   )+
#   geom_bar(stat = 'identity', position = position_dodge(), color='black',  alpha = 1)+
#   geom_linerange(position = position_dodge(), size = 1)+
#   coord_flip()+
#   ylab("Human generic endorsement")+
#   xlab("")+
#   scale_fill_viridis(breaks = c(0, 1), limits = c(0, 1))+
#   scale_y_continuous(limits = c(-0.5,0.5), breaks = c(-0.5, 0, 0.5), labels = c(0, 0.5, 1))+
#   guides(color = guide_colorbar(title = 'Prevalence',
#                                ticks = F))+
#   theme(legend.position = 'bottom')
```

Summary of endorsements for true vs. false vs. uncertain generics

```{r generic-endorsement-manipulationcheck}
# MHT's truth judgments of sentences
# t --> true; f --> false; i --> indeterminate/uncertain
all.sentences <- list("Cardinals are red." = "t",
 "Ducks have wings." = "t",
 "Kangaroos have pouches." = "t",
 "Kangaroos have spots." = "f",
 "Leopards are juvenile." = "i",
 "Leopards have spots." = "t",
 "Leopards have wings." = "f",
 "Lions are male." = "i",
 "Lions have manes." = "t",
 "Lions lay eggs." = "f",
 "Mosquitos attack swimmers." = "i",
 "Mosquitos carry malaria." = "t",
 "Mosquitos dont carry malaria." = "f",
 "Peacocks dont have beautiful feathers." = "f",
 "Peacocks have beautiful feathers." = "t",
 "Robins are female." = "i",
 "Robins carry malaria." = "f",
 "Robins lay eggs." = "t",
 "Sharks are white." = "i",
 "Sharks attack swimmers." = "t",
 "Sharks dont attack swimmers." = "f",
 "Sharks have manes." = "f",
 "Sharks lay eggs." = "i",
 "Swans are full-grown." = "i",
 "Swans are white." = "t",
 "Ticks carry Lyme disease." = "t",
 "Ticks dont carry Lyme disease."= "f",
 "Tigers dont eat people."= "f",
 "Tigers eat people." = "t",
 "Tigers have pouches."= "f")

mht.truth.judgments <- data.frame(sentence = names(all.sentences),
           truth_judgment = as.vector(unlist(all.sentences)))

d.gen.endorse.manipulation.check.bayes <- left_join(
  left_join(
    d.gen.endorse, 
    d.gen.endorse.catch %>% select(workerid, pass)
    ),
  mht.truth.judgments ) %>%
  filter(pass == 1) %>%
  rowwise() %>%
  mutate(response = ifelse(response == "agree-key", 1, 0),
         sentence = gsub('&quotechar', '', sentence),
         sentence = gsub('lyme', 'Lyme', sentence)) %>%
  group_by(truth_judgment) %>%
  summarize(k = sum(response), n = n()) %>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         low  = qbeta(.025, a, b),
         high = qbeta(.975, a, b),
         MAP_h = (a-1)/(a+b-2),
         mean = a / (a + b)) %>% 
  ungroup() %>%
  mutate(truth_judgment = factor(truth_judgment, levels = truth_judgment[order(MAP_h)]))

# save(d.gen.endorse.bayes.prev, d.gen.endorse.manipulation.check.bayes,
#      file = "../paper/cached_results/case1_endorseSummary_byItem.RData")

```

# Experiment 1b: Priors

```{r generic-endorsement-priors}
# this data set already excludes the 2nd attempt of 3 participants who completed the experiment a second time

d.gen.endorse.priors <- read.csv(paste(project.path, 
                                       "data/generics/endorsement/",
                                       "naturalGenerics-prior-trials-n57.csv", sep = ""))
gen.endorse.properties <- levels(d.gen.endorse.priors$Property)

genericEndorsementPriorModelHelpers <- '
var eps = 0.01;//Number.EPSILON;
var log = function(x){ return Math.log(x) }
var exp = function(x){ return Math.exp(x) }

var avoidEndPoints = function(x){
  x == 0 ? eps : 
  x == 1 ? 1 - eps : 
  x
}

var betaShape = function(params){
  return {
      a: params.g * params.d,
      b : (1-params.g) * params.d
  }
}

var preprocessedResponses = map(function(d){
  return avoidEndPoints(d / 100)
}, data)
'

structuredPriorModel <- '
var model = function(){
  var phi = uniformDrift({a:0, b: 1, width:0.2});
  var g = uniformDrift({a:0, b: 1, width:0.2});
  var d = uniformDrift({a:0, b: 100, width:5});
  var stableParams = betaShape({g, d})
  var stableComponent = Beta(stableParams)
  var transientComponent = Beta({a:1, b:100})
  mapData({data: preprocessedResponses}, function(d){
    factor( log(
      phi * exp(stableComponent.score(d)) +
      (1 - phi) * exp(transientComponent.score(d))
    ))
  })
  return {g, d, phi}
}
'

unstructuredPriorModel <- '
var model = function(){
  var g = uniformDrift({a:0, b: 1, width:0.2});
  var d = uniformDrift({a:0, b: 100, width:5});
  var stableParams = betaShape({g, d})
  var stableComponent = Beta(stableParams)
  mapData({data: preprocessedResponses}, function(d){
    factor(stableComponent.score(d))
  })
  return {g, d}
}
'
genericEndorsementPriorModel <- paste(genericEndorsementPriorModelHelpers, structuredPriorModel, sep = "\n")

```

```{r generic-endorsement-priors-unbounded, eval = F}
# this data set already excludes the 2nd attempt of 3 participants who completed the experiment a second time
d.gen.endorse.priors <- read.csv(paste(project.path, 
                                       "data/generics/endorsement/",
                                       "naturalGenerics-prior-trials-n57.csv", sep = ""))
gen.endorse.properties <- levels(d.gen.endorse.priors$Property)

unboundedModel <- '
var eps = 0.01;//Number.EPSILON;
var log = function(x){ return Math.log(x) }
var exp = function(x){ return Math.exp(x) }

var avoidEndPoints = function(x){
  x == 0 ? eps : 
  x == 1 ? 1 - eps : 
  x
}

var preprocessedResponses = map(function(d){
  return avoidEndPoints(d / 100)
}, data)

var shapeParams = function(params){
  return {
    a: params.g * params.d,
    b: (1 - params.g) * params.d
  }
}

var model = function(){

  // var numComponents = (1 + poisson(1));
  var numComponents = 5//discrete([1, 1, 1, 1, 1]) + 1
  var mixtureWeights = sort(normalize(repeat(numComponents, function(){
    return uniformDrift({a: 0, b:1, width: 0.2})
  }))).reverse()

  // var mixtureWeights = dirichlet(ones([numComponents, 1]));
  var componentParams = map(function(i){
    var g = uniformDrift({a: 0, b:1, width: 0.2}), d = uniformDrift({a: 0, b:100, width: 20})
    var betaParams = shapeParams({g, d})
    return betaParams
  }, _.range(numComponents))

  mapData({data: preprocessedResponses}, function(d){

    var scr = log(sum(map(function(cParams){
      cParams[1] * exp(Beta(cParams[0]).score(d))
    }, _.zip(componentParams, mixtureWeights))))
//    }, _.zip(componentParams, T.toScalars(mixtureWeights)))))
    factor(scr)
  })

  var componentPredictives = map(function(cParams){
      return beta(cParams)
  }, componentParams)

  return _.fromPairs(
    _.zip(
        map(function(i){ i + "_weight" }, _.range(5)), 
        mixtureWeights
      ).concat(
      _.zip(
        map(function(i){ i + "_predictive" }, _.range(5)), 
        componentPredictives
      )
    )
  )//sort(T.toScalars(mixtureWeights))
};
'

example.generics.prior.properties <- c(
  "dont eat people", "have beautiful feathers",
  "have wings", "are red",
  "carry malaria", "lay eggs", 
  "are female", "have spots"
  )

priorData <- filter(d.gen.endorse.priors, Property == "lay eggs")$prevalence
n_samples <- 10000
lg = 0
m.gen.endorse.priors.unbounded <- webppl(unboundedModel,
                                 data = priorData, data_var = "data",
                                 model_var = "model",
                                 chains = 1,
                                 cores = 1,
                                 inference_opts =
                                   list(method = "MCMC",
                                        #kernel = list(HMC = list(steps = 5, stepSize = 0.001)),
                                        samples = n_samples,
                                        burn = n_samples / 2,
                                        lag = lg,
                                        verbose = T))
```

```{r eval = F}
m.gen.endorse.priors.unbounded %>%
      separate(Parameter, into = c("c", "var")) %>% 
      ggplot(., aes( x=  value, fill = c))+
 geom_histogram(position = position_dodge(), binwidth = 0.05)+
 facet_wrap(~var)+
  scale_fill_solarized()
```

```{r generic-endorsement-priors-bayesFactors, eval = F}
genericEndorsementPriorModelPriorLikelihoodHelpers <- '
var properties = _.uniqBy(_.map(data, "Property"));

var eps = 0.01;//Number.EPSILON;
var log = function(x){ return Math.log(x) }
var exp = function(x){ return Math.exp(x) }

var avoidEndPoints = function(x){
  x == 0 ? eps : 
  x == 1 ? 1 - eps : 
  x
}

var betaShape = function(params){
  return {
      a: params.g * params.d,
      b : (1-params.g) * params.d
  }
}
'

genericEndorsementPriorModelsPriorLikelihood <- '
var transientComponent = Beta({a:1, b:100})

var model = function(){

  var propertyLogLikes = map(function(property){

    var propertyData = _.map(_.filter(data, {Property: property}), "prevalence");
    var preprocessedResponses = map(function(d){
        return avoidEndPoints(d / 100)
    }, propertyData);

    var phi = uniformDrift({a:0, b: 1, width:0.2});
    var g = uniformDrift({a:0, b: 1, width:0.2});
    var d = uniformDrift({a:0, b: 100, width:5});
    var stableParams = betaShape({g, d})
    var stableComponent = Beta(stableParams)

    var logLikes = map(function(d){
      var structuredLL = log(
        phi * exp(stableComponent.score(d)) +
        (1 - phi) * exp(transientComponent.score(d))
      )

      var unstructuredLL = stableComponent.score(d);

      return {structuredLL, unstructuredLL}
    }, preprocessedResponses)

    return {
      structuredLL: sum(_.map(logLikes, "structuredLL")), 
      unstructuredLL: sum(_.map(logLikes, "unstructuredLL"))
    }

  }, properties)

  return {
    structuredLL: sum(_.map(propertyLogLikes, "structuredLL")), 
    unstructuredLL: sum(_.map(propertyLogLikes, "unstructuredLL"))
  }
}
'

```

```{r generic-endorsement-priors-bayesFactorsRun, eval = F}
marginal.prior.likelihoods <- data.frame()

# 100 samples takes 30s
m.gen.endorse.priors.logLike <- webppl(
    paste(genericEndorsementPriorModelPriorLikelihoodHelpers,
          genericEndorsementPriorModelsPriorLikelihood,
          sep = '\n'), 
    data = d.gen.endorse.priors, data_var = "data",
    model_var = "model", 
    inference_opts = list(method = "forward", samples = 1000),
    chains = 3, cores = 3
) %>% 
  mutate(value = as.numeric(value))

min.ll.1 <- min(
  m.gen.endorse.priors.logLike$value[
    is.finite(
      m.gen.endorse.priors.logLike$value
      )
    ]
  )

m.gen.endorse.priors.logLike <- m.gen.endorse.priors.logLike %>% 
  rowwise() %>%
  mutate(value = ifelse(value == -Inf, min.ll.1, value))

save(m.gen.endorse.priors.logLike, file = '../analysis/generics-priors-loglike.Rdata')
```

```{r generic-endorsement-priors-bayesFactorsLoad}
load(file = '../analysis/generics-priors-loglike.Rdata')
marginal.prior.likelihoods <- m.gen.endorse.priors.logLike %>%
  group_by(Chain, Parameter) %>%
  summarize(mll  = logmeanexp(value))

marginal.prior.likelihoods <- marginal.prior.likelihoods %>% 
  spread(Parameter, mll) %>%
  mutate(logBF = structuredLL - unstructuredLL, BF = exp(logBF))

roundedBF1 = round(mean(marginal.prior.likelihoods$logBF)/ 10000)*10000
```

# Model comparisons

## Regression models

```{r generic-regression-models}
d.cv <- read.csv(paste(project.path, "data/generics/endorsement/", "cue-validity-2-freeProduction-trials.csv", sep = ""))

d.target.items <- fromJSON(paste(project.path, "data/generics/endorsement/", "originalStims_wrongDeployment.json", sep = "")) %>%
  mutate(sentence = paste(category, property))

### Preprocessing
# - Force all responses to lower case. 
# - Remove spaces.
# - Fix mosquito mispellings.
# - Count "deertick" as "tick".
# - Remove plural "s".
# 
# Mark if the produced animal matches the generic of interest.

mosquito.mispellings <- c("mosqu", "mesqu", "misqu", "mosiq")

d.target.items <- fromJSON(paste(project.path, "data/generics/endorsement/", "originalStims.json", sep = "")) %>%
  mutate(
    property = gsub("'", "", property),
    Property = gsub("'", "", Property),
    sentence = paste(Category, Property),
    item = paste(category, property))

d.cv.bootstrapped <- data.frame()
resample_n <- length(levels(factor(d.cv$workerid)))
for (i in 1:10){

  d.cv.bsample <- d.cv %>% 
    select(workerid, property, response) %>% 
    spread(property, response) %>% 
    sample_n(resample_n, replace = TRUE) %>%
    gather(property, category, -workerid) %>% 
        group_by(property) %>%
    mutate(n = n(),
           item = paste(category, property)) %>%
    filter(item %in% d.target.items$item) %>%
    group_by(category, property) %>%
    summarize(mentions = n(),
              trials = mean(n), # mean(n) == n, because it's just the number of subjects
              prop = mentions / trials,
              prop = ifelse(is.na(prop), 0.01, prop),
              iteration = i)

  d.cv.bootstrapped <- bind_rows(d.cv.bootstrapped, d.cv.bsample)
}

empiricalLower = function(dist){
  xi <- quantile(dist, 0.025)
  return (xi[["2.5%"]])
}
empiricalUpper = function(dist){
  xi <- quantile(dist, 0.975)
  return (xi[["97.5%"]])
}
empiricalMean = function(dist){
  xi <- quantile(dist, 0.5)
  return (xi[["50%"]])
}

d.gen.cv.summary <- left_join(
    d.target.items %>% select(-sentence),
    d.cv.bootstrapped %>%
      group_by(property) %>%
      summarize(cv_mean = empiricalMean(prop),
                cv_ci_lower = empiricalLower(prop),
                cv_ci_upper = empiricalUpper(prop))
) %>%
  mutate(
    cv_ci_lower = ifelse(is.na(cv_mean), 0.01, cv_ci_lower),
    cv_ci_upper = ifelse(is.na(cv_mean), 0.01, cv_ci_upper),
    cv_mean = ifelse(is.na(cv_mean), 0.01, cv_mean)
  )

# quantile(filter(d.cv.bootstrapped, property == "is red")$prop, c(0.025, 0.5, 0.975))
# 
# d.cv.bootstrapped %>%
#   group_b

# need to filter by items specified with endorsement
# d.gen.cv.summary <- left_join(
#   d.target.items %>% select(-sentence),
#   d.cv %>%
#     rename(category = response) %>%
#     group_by(property) %>%
#     mutate(n = n(),
#            item = paste(category, property)) %>%
#     filter(item %in% d.target.items$item) %>%
#     group_by(category, property) %>%
#     summarize(mentions = n(),
#               trials = mean(n), # mean(n) == n, because it's just the number of subjects
#               prop = mentions / trials)
#   ) %>%
#   mutate(prop = ifelse(is.na(prop), 0.01, prop))

d.gen.endorse.prev.cue <- left_join(
  left_join(
    left_join(
      d.gen.endorse, 
      d.gen.endorse.catch %>% select(workerid, pass)
      ) %>%
      filter(pass == 1) %>%
    rowwise() %>%
    mutate(response = ifelse(response == "agree-key", 1, 0),
         sentence = gsub('&quotechar', '', sentence),
         sentence = gsub('lyme', 'Lyme', sentence),
          sentence = gsub('[.]', '', sentence)
    ),
    d.prev.summary %>% 
      rename(prev_mean = mean,
             sentence = item,
             prev_ci_lower = ci_lower,
             prev_ci_upper = ci_upper) %>%
      ungroup() %>%
      mutate(sentence = gsub('[.]', '', sentence))
    ),
  d.gen.cv.summary %>% select(-item))

glm.gen.endorse.prev <- glm(response ~ prev_mean, 
              data = d.gen.endorse.prev.cue, family = 'binomial')

glm.gen.endorse.prev.cv <- glm(response ~ prev_mean + cv_mean, 
              data = d.gen.endorse.prev.cue, family = 'binomial')

d.gen.endorse.prev.cue.uniq <- unique(select(
    d.gen.endorse.prev.cue, Property, Category, sentence, 
    prev_mean, prev_ci_lower, prev_ci_upper, cv_mean, cv_ci_lower, cv_ci_upper
  ))

d.gen.endorse.prev.cue.uniq.lower <- d.gen.endorse.prev.cue.uniq %>% 
  select(-prev_mean, -cv_mean) %>% 
  rename(prev_mean = prev_ci_lower,
         cv_mean = cv_ci_lower)
d.gen.endorse.prev.cue.uniq.upper <- d.gen.endorse.prev.cue.uniq %>% 
  select(-prev_mean, -cv_mean) %>% 
  rename(prev_mean = prev_ci_upper,
         cv_mean = cv_ci_upper)

d.gen.endorse.regression.prevalence <- left_join(
  d.gen.endorse.bayes,
  d.gen.endorse.prev.cue.uniq %>%
    mutate(
      prediction_mean = predict(glm.gen.endorse.prev, ., type = "response"),
      prediction_ci_lower = predict(glm.gen.endorse.prev, d.gen.endorse.prev.cue.uniq.lower, type = "response"),
      prediction_ci_upper = predict(glm.gen.endorse.prev, d.gen.endorse.prev.cue.uniq.upper, type = "response")
    ) %>%
    mutate(sentence = paste(Category, " ", Property, ".", sep = ""))
) %>% mutate(
  sqErr = (MAP_h - prediction_mean) ^ 2
)

d.gen.endorse.regression.prevalence.cuevalidity <- left_join(
  d.gen.endorse.bayes,
  d.gen.endorse.prev.cue.uniq %>%
    mutate(
      prediction_mean = predict(glm.gen.endorse.prev.cv, ., type = "response"),
      prediction_ci_lower = predict(glm.gen.endorse.prev.cv, d.gen.endorse.prev.cue.uniq.lower, type = "response"),
      prediction_ci_upper = predict(glm.gen.endorse.prev.cv, d.gen.endorse.prev.cue.uniq.upper, type = "response"),
      sentence = paste(Category, " ", Property, ".", sep = "")
      )
) %>% mutate(sqErr = (MAP_h - prediction_mean) ^ 2)

d.gen.endorse.regression <- bind_rows(
  d.gen.endorse.regression.prevalence.cuevalidity %>%
    mutate(src = "regression_Prev_Cuevalidity"),
  d.gen.endorse.regression.prevalence %>% 
    mutate(src = "regression_Prev")
  )

# save(d.gen.endorse.regression.prevalence, 
#      d.gen.endorse.regression.prevalence.cuevalidity, 
#      d.gen.endorse.regression, file = "../paper/cached_results/case1_regressionResults.RData")
load("../paper/cached_results/case1_regressionResults.RData")

r2.gen.n <- length(d.gen.endorse.bayes$sentence)
r2.gen.regression.prev <- round(with(d.gen.endorse.regression.prevalence, cor(MAP_h, prediction_mean))^2,2)
mse.gen.regression.prev <-  round(mean(d.gen.endorse.regression.prevalence$sqErr), 3)

r2.gen.regression.prev.cv <- round(with(d.gen.endorse.regression.prevalence.cuevalidity, cor(MAP_h, prediction_mean))^2, 3)
mse.gen.regression.prev.cv <- round(mean(d.gen.endorse.regression.prevalence.cuevalidity$sqErr), 3)

intermediate.prev.quantiles <-  quantile(d.gen.endorse.regression.prevalence$prev_mean, c(0.25, 0.76))

d.gen.endorse.regression.prevalence.intermediateprev <- d.gen.endorse.regression.prevalence %>%
  filter(
    (prev_mean > intermediate.prev.quantiles[["25%"]]) &
    (prev_mean < intermediate.prev.quantiles[["76%"]])
)
r2.gen.n.intermedprev <- length(d.gen.endorse.regression.prevalence.intermediateprev$sentence)

r2.gen.regression.prev.intermedprev <- round(with(d.gen.endorse.regression.prevalence.intermediateprev, cor(MAP_h, prediction_mean))^2,2)
mse.gen.regression.prev.intermedprev <-  round(mean(d.gen.endorse.regression.prevalence.intermediateprev$sqErr), 3)
```

## RSA models

### model code
```{r rsaHelpers}
rsaHelpers <- '
var probability = function(Dist, x) {
    return Math.exp(Dist.score(x));
}
var targetUtterance = "generic";

var round = function(x){
  return Math.round(x*1000)/1000
}
var utterancePrior = Infer({model: function(){
  return uniformDraw([targetUtterance,"silence"])
}});

var thetaBins = map(round, _.range(0.01, 0.98, 0.01))
// var thetaBins = [
//    0.01, 0.05,0.1,0.15,0.2,0.25,0.3,0.35,0.4,0.45,
//    0.5, 0.55,0.6,0.65,0.7,0.75,0.8,0.85,0.9,0.95
 // ]
var thetaPrior = Infer({model: function(){
 return uniformDraw(thetaBins)
}});

var bins = map(round, _.range(0.01, 0.99, 0.01))

// var bins = [
  // 0.01,0.05,0.1,0.15,0.2,0.25,0.3,0.35,0.4,0.45,
  // 0.5,0.55,0.6,0.65,0.7,0.75,0.8,0.85,0.9,0.95,0.99
// ];

var meaning = function(utt,state, theta) {
  return utt=="generic"? state > theta :
         utt=="generic is false"? state<=theta :
         utt=="silence"? true :
         utt=="some"? state>0.01:
         utt=="most"? state> 0.5:
         utt=="all"? state >= 0.99:
         true
}

var mixture = data.prior.mix[0];
var priorParams = data.prior.params[0];

// var statePrior = Infer({model: function(){
//  var component = flip(mixture);
//  return component ?
//    categorical({
//      vs: bins,
//      ps: map(function(b) {
//        return probability(Beta(priorParams), b) + Number.EPSILON
//      }, bins )
//    }) :
//    categorical({
//      vs: bins,
//      ps: map(function(b) {
//        return probability(Beta({a:1,b:50}), b) + Number.EPSILON
//      }, bins )
//    })
// }});
'
```

```{r rsaInterpretationModels}
no.utterance.model <- '
var noUtteranceInterpreter = function() {
  Infer({model: function(){
   var state = sample(Beta( flip(mixture) ? priorParams : {a:1,b:100}))
    return { state }
 }, method: "forward", samples: 20000, verbose: T})}
'

no.utterance.model.3comp <- '
var componentParams = [
  {a: 1, b:100},
  {a: priorParams.a1, b:priorParams.b1},
  {a: priorParams.a2, b:priorParams.b2}
]
display(data.prior.mix)
var noUtteranceInterpreter = function() {
  Infer({model: function(){
   var state = beta(componentParams[discrete(data.prior.mix)])
    return {
      state: state, 
  }
  }, method: "forward", samples: 20000, verbose: T})}
'

fixed.threshold.model <- '
var fixedThresholdInterpreter = function(threshold) {
  Infer({model: function(){
  // var state = sample(Beta( flip(mixture) ? priorParams : {a:1,b:100}))
  var state = sample(statePrior)
  condition( state > threshold)
    return {
      state: state
  }
 // }, method: "rejection", samples: 20000, burn:5000, verbose: T})}
 }, method: "enumerate"})}
'

uncertain.threshold.model <- '
var uncertainThresholdInterpreter = function() {
  Infer({model: function(){
   var state = sample(Beta( flip(mixture) ? priorParams : {a:1,b:100}))
    factor( Math.log(state) )
    return {
      state: state, 
  }
  }, method: "rejection", samples: 20000, burn:5000, verbose: T})}

'

uncertain.threshold.model.3comp <- '
var componentParams = [
  {a: 1, b:100},
  {a: priorParams.a1, b:priorParams.b1},
  {a: priorParams.a2, b:priorParams.b2}
]
display(data.prior.mix)
display(JSON.stringify(componentParams))
var uncertainThresholdInterpreter = function() {
  Infer({model: function(){
  var component = discrete(data.prior.mix)
  // display(component)
  // display(JSON.stringify(componentParams[component]))
   var state = beta(componentParams[component])
    factor( Math.log(state) )
    return {
      state: state, 
  }
  }, method: "rejection", samples: 20000, verbose: T})}
'
```

```{r simulationCalls}
fixed.threshold.calls <- '
_.fromPairs(map(function(t){return [t, fixedThresholdInterpreter(t)]}, [0.1, 0.3, 0.5, 0.7, 0.9]))
'
no.utterance.call <- ' 
  noUtteranceInterpreter()
'
uncertain.threshold.call <- '
  uncertainThresholdInterpreter()
'
```

### model analysis

```{r generic-RSAmodels, fig.width = 8}
example.generics <- c("Tigers dont eat people", 
                      "Mosquitos carry malaria", 
                      "Robins lay eggs", 
                      "Robins are female",
                      "Leopards have spots")

n_chains <- 4
n_samples <- 100000
burn <- n_samples / 2
lg <- 20

#model_prefix <- "results-generics-jointModel-S1-smntcs_"
#fixed_model_prefix <- "results-generics-jointModel-wNoise-inferFixedThresholdCts-20bins-S1-smntcs_some-"
model_prefix <- "results-generics-jointModel-S1-smntcs_generic-3compPriors-2compRef-"
m.samp <- data.frame()
#m.samp.fixed <- data.frame()
for (i in seq(1, n_chains)){
  mi.gen <- read_csv(paste(project.path,  "models/generics/results/", model_prefix,
                           n_samples, "_burn", burn, "_lag", lg, "_chain", i, ".csv", sep = ""))
  m.samp <- bind_rows(m.samp, mi.gen %>% mutate(chain = i))

#   mi.fixed <- fread(paste(project.path,  "models/generics/results/", fixed_model_prefix,
#                     n_samples, "_burn", burn, "_lag", lg ,"_chain", i, ".csv", sep = ""))
#   m.samp.fixed <- bind_rows(m.samp.fixed, mi.fixed %>% mutate(chain = i))

}

# 
# save(m.samp,
#      file = paste(project.path,  "models/generics/results/", model_prefix, "generic-",
#                     n_samples, "_burn", burn, "_lag", lg, "_",n_chains , "chains.RData", sep = ""))
#

# save(m.samp.fixed,
#      file = paste(project.path, "models/generics/results/results-generics-jointModel-S1-some-inferFixedThresholdCts-noise-100000_burn50000_lag20_2chains.RData", sep = ""))

     # file = paste(project.path,  "models/generics/results/", model_prefix, "some-",
    #                 n_samples, "_burn", burn, "_lag", lg, "_",n_chains , "chains.RData", sep = ""))

## MODEL RESULTS
#load(file = paste(project.path,  "models/generics/results/", "results-generics-jointModel-S1-smntcs_generic-100000_burn50000_lag20_3chains.RData", sep = ""))

# load(m.samp.fixed, 
#      file = paste(project.path, "models/generics/results/results-generics-jointModel-S1-some-inferFixedThresholdDisc-noise-100000_burn50000_lag20_2chains.RData", sep = ""))

load(file = paste(project.path, "models/generics/results/results-generics-jointModel-S1-some-inferFixedThresholdCts-noise-100000_burn50000_lag20_2chains.RData", sep = ""))

#m.samp.fixed <- fread(paste(project.path,  "models/generics/results/", "results-generics-jointModel-wNoise-inferFixedThresholdCts-20bins-S1-smntcs_some-10000_burn5000_lag20_chain1.csv", sep = ""))
# 
# m.samp.fixed <- fread(paste(project.path,  "models/generics/results/", "results-generics-jointModel-wNoise-inferFixedThreshold-20bins-S1-smntcs_some-100000_burn50000_lag20_chain1.csv", sep = ""))
# 
# tail(m.samp.fixed)

fixed.params.posterior <-  m.samp.fixed %>% 
  filter(param %in% c("noise", "fixedThreshold")) %>%
  group_by(param) %>% 
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

m.gen.somemodel.endorsement <- m.samp.fixed %>%
  filter(type == 'endorsement', param == "some") %>%
  group_by(type, param, property, category) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

# m.gen.fullmodel.endorsement <- m.samp.samp %>%
#   rename(type = Parameter, property = Property, category = Category, val = Value, param = Extra) %>%
#   group_by(type, property, category, param) %>%
#   summarize(MAP = estimate_mode(val),
#             cred_upper = hdi_upper(val),
#             cred_lower = hdi_lower(val))

m.gen.fullmodel.endorsement <- m.samp %>%
 filter(type == 'endorsement', param == 'generic') %>%
  # filter(type == 'predictive', param == 'generic') %>%
 # filter(type == 'endorsement') %>%
  group_by(type, param, property, category) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

uncertain.threshold.s1opt <- m.samp %>% 
  filter(type == "param") %>%
  group_by(param) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

fixed.threshold.s1opt <- m.samp.fixed %>% 
  filter(type == "param") %>%
  group_by(param) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

fixed.threshold.posterior <- filter(fixed.params.posterior, param == "fixedThreshold")
generics.fixed.noise.posterior <- filter(fixed.params.posterior, param == "noise")
```

```{r generic-endorsement-referent-prevalence-model, eval = F}
referentPrevalenceModel <- '
var model = function(){
  var g = uniformDrift({a:0, b: 1, width:0.2});
  var d = uniformDrift({a:0, b: 100, width:5});
  var betaParams = betaShape({g, d})
  var referentPrevDist = Beta(betaParams)
  mapData({data: preprocessedResponses}, function(d){
    observe(referentPrevDist, d)
  })
  return {g, d}
}
'
genericEndorsementRefPrevModel <- paste(genericEndorsementPriorModelHelpers, referentPrevalenceModel, sep = "\n")
```

```{r generic-referent-prevalence-model-run, eval = F}
m.gen.refprev.predictive <- data.frame()
n_samples <- 10000
lg = 0

for (p in example.generics ) {

  prevData <- d.prev %>% 
  mutate(sentence = paste(Category," " ,Property, sep = "")) %>%
  filter(sentence == p)

  m.gen.refprev <- webppl(genericEndorsementRefPrevModel,
                                 data = prevData$prevalence, data_var = "data",
                                 model_var = "model",
                                 chains = 1,
                                 cores = 1,
                                 inference_opts =
                                   list(method = "MCMC",
                                        samples = n_samples,
                                        burn = n_samples / 2,
                                        lag = lg,
                                        verbose = T))

    m.gen.refprev.predictive.item <- m.gen.refprev %>%
      group_by(Iteration, Chain) %>%
      spread(Parameter, value) %>%
      rowwise() %>%
      mutate(
        a = g*d,
        b = (1-g)*d,
        prevalence =rbeta(n = 1, shape1 = a,
                                 shape2 = b),
        Category = prevData[1,]$Category,
        Property = prevData[1,]$Property,
        src = "prevAloneModel"
      ) %>% ungroup()

    m.gen.refprev.predictive <- bind_rows(
      m.gen.refprev.predictive,
      m.gen.refprev.predictive.item %>%
        select( src, Category, Property, prevalence)
    )

}
#save(m.gen.endorse.priors.posteriorPredictive, file = '../analysis/generics-priors-posteriorPredictives.Rdata')

load("../analysis/generics-priors-posteriorPredictives.Rdata")
```

```{r gen-prev-forward, eval=F}
#Forward sample Target Prevalence from posterior on parameters (joint inference model)
m.samp.prev.params <- m.samp %>%
    filter(type == "targetPrevalence") %>%
    #filter(type == "withinKind") %>%
    #mutate(parameter = paste(param, property, category, sep = "_")) %>%
    mutate(parameter = paste(param, property, category, component, sep = "_")) %>%
    select(-param, -property, -category, -chain, -type, -component) %>%
    #mutate(item = paste(parameter, property, sep = "_")) %>%
    group_by(parameter) %>%
    mutate(iteration = ave(parameter==parameter, parameter, FUN=cumsum)) %>%
    ungroup() %>%
    separate(parameter, into = c("parameter", "property", "category", "component"), sep= "_")

m.samp.prevalence <- m.samp.prev.params %>%
    group_by(category, property, iteration) %>%
    spread(parameter, val) %>%
    rowwise() %>%
    mutate(
      a = mean*sampleSize,
      b = (1-mean)*sampleSize,
      prevalence = rbeta(n = 1, shape1 = a, shape2 = b)
      ) %>%
    ungroup()

m.samp.prevalence <- m.samp.prev.params %>%
    mutate(parameter = paste(parameter, component, sep = "_")) %>%
    select(-component) %>%
    group_by(category, property, iteration) %>%
    spread(parameter, val) %>%
    rowwise() %>%
    mutate( 
      component = rbinom(n = 1, size = 1, p = mix_1),
      a = ifelse(component == 1, mean_1*sampleSize_1,mean_2*sampleSize_2),
      b = ifelse(component == 1, (1-mean_1)*sampleSize_1, (1-mean_1)*sampleSize_2),
      prevalence = rbeta(n = 1, shape1 = a, shape2 = b)
      ) %>%
    ungroup()

m.samp.prevalence %>% filter(property %in% c("are female", "lay eggs"),
                             category == "Robins") %>%
  ggplot(., aes( x= prevalence))+
  geom_histogram()+
  facet_wrap(~property)
```

```{r example generics}
example.generics.properties <- c("dont eat people", 
                      "carry malaria", 
                      "lay eggs", 
                      "are female",
                      "have spots")
```

```{r gen-prev-dists, eval=F}
gen.prev.dists <- bind_rows(
  d.prev %>% 
  mutate(sentence = paste(Category," " ,Property, sep = "")) %>%
  filter(sentence %in% example.generics) %>%
  mutate(prevalence = prevalence / 100,
         src = 'data') %>% 
    select(-workerid, -trial_type),
  m.samp.prevalence %>% 
    mutate(sentence = paste(category," " ,property, sep = "")) %>%
    filter(sentence %in% example.generics) %>%
    select(property, category, prevalence) %>%
    rename(Category = category, Property = property) %>%
    mutate(src = 'model'),
  m.gen.refprev.predictive
)

ggplot(gen.prev.dists, aes( x = prevalence, color = src))+
  geom_density(aes( y = ..scaled.. ))+
    #  stat_ecdf(size = 1)+
  #geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]),
   #              position = position_dodge())+
  facet_wrap(~Property + Category, nrow = 1)+
  scale_color_solarized()+
  scale_x_continuous(limits = c(-0.05,1.05), breaks = c(0, 0.5, 1)) +
  scale_y_continuous(limits = c(-0.01,1.01), breaks = c(0, 0.5, 1)) +
  theme(strip.text.y = element_text(angle = 0))

```

```{r generic-model-scatters, cache = F}
m.gen.endorse.rsa <- bind_rows(
  left_join(
    d.gen.endorse.bayes,
    m.gen.fullmodel.endorsement %>%
      mutate(sentence = paste(category, " ", property, ".", sep = "")) %>%
      ungroup() %>%
      select(-type, -param)
  ) %>%
    mutate(src = "generics_model"),
  left_join(
    d.gen.endorse.bayes,
    m.gen.somemodel.endorsement %>%
      mutate(sentence = paste(category, " ", property, ".", sep = "")) %>%
      ungroup() %>%
      select(-type, -param)
  ) %>%
    mutate(src = "some_model")
)

r2.gen.rsa.generics <- compute_r2(
  m.gen.endorse.rsa %>% filter(src == "generics_model"), 
  "MAP", "MAP_h", sigfigs = 4
)

r2.gen.rsa.fixed <- compute_r2(
  m.gen.endorse.rsa %>% filter(src == "some_model"),  
  "MAP", "MAP_h", sigfigs = 4
)

mse.gen.rsa.generics <- compute_mse(
  m.gen.endorse.rsa %>% filter(src == "generics_model"), 
  "MAP", "MAP_h", sigfigs = 5
)

mse.gen.rsa.fixed <- compute_mse(
  m.gen.endorse.rsa %>% filter(src == "some_model"),  
  "MAP", "MAP_h", sigfigs = 5
)

d.gen.endorse.rsa.regression <- bind_rows(
  d.gen.endorse.regression,
  left_join(
    m.gen.endorse.rsa %>% 
      rename(prediction_mean = MAP, prediction_ci_lower = cred_lower, prediction_ci_upper = cred_upper),
    d.gen.endorse.regression %>% 
      select(sentence, Property, Category, prev_mean, cv_mean)
  )
) %>%
  mutate(src = factor(src, levels = c( "some_model",
                                      "generics_model",
                                      "regression_Prev", 
                                      "regression_Prev_Cuevalidity"
                                     ),
                      labels = c("Fixed semantics model",
                                 "Uncertain semantics model",
                                 "Prevalence",
                                 "Cue validity + Prevalence"
                                 )))

generics.endorsement.models <- ggplot(d.gen.endorse.rsa.regression, 
                                      aes ( x = prediction_mean, xmin = prediction_ci_lower, xmax = prediction_ci_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = prev_mean ))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.4)+
  geom_errorbarh(alpha = 0.4)+
  geom_point(shape = 21, size = 3)+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  #scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("Model prediction")+
  ylab("Human generic endorsement")+
  facet_wrap(~src, nrow = 2)+
  #theme(legend.position = "bottom")+
  guides(fill = F)
generics.endorsement.models
```

```{r generic-model-insets2, fig.width = 4.75, fig.height = 1.5}
load('../analysis/generics-priors-posteriorPredictives.Rdata')
lower_bins <-seq(0, 0.975, 0.025)
upper_bins <- seq(0.025, 1, 0.025)
mid_bins <- (upper_bins - lower_bins)/2 + lower_bins

round_to_bin <- function(x){
  index <- if (x==0){ 1 } else if (x ==1){ length(mid_bins)} else{
    x <= upper_bins & x > lower_bins
  }
  #if (sum(index) > 1) { print (x) }
  return (mid_bins[index])
}

m.gen.endorse.priors.binned.i <- m.gen.endorse.priors.posteriorPredictive %>% 
  filter(src == "Data") %>%
  #group_by(property) %>%
  #sample_n(28, replace = TRUE) %>%
  #gather(category, response, -workerid, -property) %>%
  rowwise() %>%
  mutate(binned_response = round_to_bin(prevalence)) %>%
  group_by(Property, binned_response) %>%
  count() %>%
  ungroup() %>%
  spread(binned_response, n)

m.gen.endorse.priors.binned.i[is.na(m.gen.endorse.priors.binned.i)] <- 0

m.gen.endorse.priors.binned <- m.gen.endorse.priors.binned.i %>%
  gather(state, n, -Property) %>%
  group_by(Property) %>%
  mutate(prop = (n+1) / sum(n + 1),
       prop = round(prop, 5))

rsaHelpers2 <- '

var roundTo3 = function(x){
  return Math.round(x * 10000) / 10000
}

var probability = function(Dist, x) {
    return Math.exp(Dist.score(x));
}
var targetUtterance = "generic";

var round = function(x){
  return Math.round(x*1000)/1000
}
var utterancePrior = Infer({model: function(){
  return uniformDraw([targetUtterance,"silence"])
}});

var meaning = function(utt,state, theta) {
  return utt=="generic"? state > theta :
         utt=="generic is false"? state<=theta :
         utt=="silence"? true :
         utt=="some"? state>0.01:
         utt=="most"? state> 0.5:
         utt=="all"? state >= 0.99:
         true
}

var bins = _.map(data.prior, "state");

var thetaBins = map2(function(b1, b2){
  return roundTo3((b2-b1)/2 + b1);
}, bins.slice(0, bins.length-1), bins.slice(1))

var priorProbs =  _.map(data.prior, "prop");

var thetaPrior = Infer({model: function(){
 return uniformDraw(thetaBins)
}});

var uncertainThresholdInterpreter = Infer({model: function(){
   var state = categorical({vs: bins, ps: priorProbs})
   factor( Math.log(state) )
    return {
      state: state, 
  }
  }, method: "enumerate"})

var prior = Infer({model: function(){
   var state = categorical({vs: bins, ps: priorProbs})
    return {
      state: state, 
  }
  }, method: "enumerate"})

var returnVal = {uncertainThresholdInterpreter, prior}
returnVal
'

properties <- unique(m.gen.endorse.priors.binned$Property)
gen.listener.predictions <- data.frame()
for (p in unique(m.gen.endorse.priors.binned$Property)){

  priorProbs = m.gen.endorse.priors.binned %>% 
    filter(Property == p) %>%
    ungroup() %>%
    select(state, prop) %>%
    mutate(state = as.numeric(state))

  wp.rs <- webppl(
    program_code = rsaHelpers2, 
    data = list(prior = priorProbs, 
                utt = "generic") ,
    data_var = "data"
    )
  wp.rs.generic <- get_samples(data.frame(wp.rs$uncertainThresholdInterpreter) %>%
                                 rename(prob = probs), 10000)
  wp.rs.prior <- get_samples(data.frame(wp.rs$prior) %>%
                                 rename(prob = probs), 10000)

  gen.listener.predictions <- bind_rows(
   gen.listener.predictions, 
    wp.rs.prior  %>% mutate(property = p, Parameter = "state_Prior"),
    wp.rs.generic  %>% mutate(property = p, Parameter = "state_Posterior")
   )
}

example.generics.properties

example.generics <- c("Tigers dont eat people", 
                      "Mosquitos carry malaria", 
                      "Robins lay eggs", 
                      "Robins are female",
                      "Leopards have spots")

gen.inset.distributions <- bind_rows(
  gen.listener.predictions %>%
    mutate(category = NA),
  d.gen.endorse.priors %>% 
    mutate(Sentence = paste(Category, Property)) %>%
    filter(Sentence %in% example.generics) %>%
    select(prevalence, Property, Category) %>%
    rename(state = prevalence, property = Property, category = Category) %>%
    mutate(Parameter = "speakerBeliefs",
           state = state / 100)
)

category.text.labels <- data.frame(property = c("dont eat people", 
                      "carry malaria", "lay eggs",  "are female", "have spots"),
             category = c("Tigers", "Mosquitos", "Robins", "Robins", "Leopards"),
             x = c(0, 0.3, 0, 0, 0.3),
            y = c(3.8, 4, 4, 4, 4))

library(ggridges)
gen.inset.distributions.refactored <- gen.inset.distributions %>% filter(property %in% category.text.labels$property)

#save(gen.inset.distributions.refactored)
ggplot( gen.inset.distributions.refactored %>%
         mutate(Parameter = factor(Parameter, levels = c("state_Prior", "state_Posterior", "speakerBeliefs"),
                                   labels = c("Null interpretation\n(Prevalence prior)", "Generic interpretation\n(model)", 
                                              "Referent prevalence")),
                property = factor(property, levels = c("have spots", "lay eggs", "are female", "carry malaria", "dont eat people"))),
       aes(x = state, y = Parameter, group = Parameter ))+
  scale_y_discrete(expand = c(0.01, 0)) +   # will generally have to set the `expand` option
  scale_x_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 1))  +    # for both axes to remove unneeded padding
  facet_wrap(~property, nrow = 1)+
  geom_text_repel(data = category.text.labels,
                  aes(label = category, x = x , y = y),
                  inherit.aes = F)+
    #scale_fill_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
    #scale_color_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
  geom_density_ridges(scale = 1, alpha = 0.3, adjust = 1) +
  scale_fill_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    scale_color_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    scale_alpha_manual(values = c(0.6, 0.4, 0))+
    #scale_linetype_manual(values = c(3, 4, 2, 1))+
    scale_linetype_manual(values = c(3, 4, 1))+
    #scale_x_continuous(breaks = c(0, 1), limits= c(0, 1))+
    #scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    xlab("Prevalence") +
    ylab("")+
    theme(legend.position = "bottom", legend.title = element_blank())

```

```{r generic-model-insets, fig.width = 4.75, fig.height = 1.5}
# m.gen.fullmodel.prior.parameters <- m.samp %>%
#   filter(type == "prior", property %in% example.generics.properties) %>%
#   group_by(param, property, category) %>%
#   summarize(MAP = estimate_mode(val),
#             cred_upper = hdi_upper(val),
#             cred_lower = hdi_lower(val))

m.gen.fullmodel.prior.parameters <- m.samp %>%
  filter(type == "prior", property %in% example.generics.properties) %>%
  group_by(param, property, category, component) %>%
  summarize(expval = mean(val),
            MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

# use MAP estimates to generate L(h | generic) & L(h | silence) predictions

# m.gen.fullmodel.prior.parameters.tidy <- m.gen.fullmodel.prior.parameters %>%
#   ungroup() %>%
#   select(param, property, category, MAP) %>%
#   mutate(param = paste(param, category, sep = "_")) %>%
#   select(-category) %>%
#   spread(param, MAP) %>%
#   rename(mix = mixture_NA, 
#          stable_mean = stableFreq_mean, 
#          stable_concentration = stableFreq_sampleSize) %>%
#   mutate( a = stable_mean * stable_concentration, 
#           b = (1 - stable_mean) * stable_concentration)

m.gen.fullmodel.prior.parameters.tidy <- m.gen.fullmodel.prior.parameters %>%
  ungroup() %>%
  mutate(param = paste(param, component, sep = "_")) %>%
  select(param, property, category, expval) %>%
  mutate(param = paste(param, category, sep = "_")) %>%
  select(-category) %>%
  spread(param, expval) %>%
  rename(mixture_0 = mixture_0_NA, 
         mixture_1 = mixture_1_NA, 
         mixture_2 = mixture_2_NA, 
         stable_mean_1 = stableFreq_1_mean, 
         stable_concentration_1 = stableFreq_1_sampleSize,
         stable_mean_2 = stableFreq_2_mean, 
         stable_concentration_2 = stableFreq_2_sampleSize
         ) %>%
  mutate( 
    a_1 = stable_mean_1 * stable_concentration_1, 
    b_1 = (1 - stable_mean_1) * stable_concentration_1,
    a_2 = stable_mean_2 * stable_concentration_2, 
    b_2 = (1 - stable_mean_2) * stable_concentration_2
    )

gen.listener.predictions <- data.frame()

# run interpreter model to generate "hypothetical intepretation" plots
for (p in example.generics.properties){
 priorParams <- m.gen.fullmodel.prior.parameters.tidy %>% filter(property == p)
 # inputData = list(prior = list(params = data.frame(a = priorParams[["a"]],
 #                                                   b = priorParams[["b"]]),
 #                               mix = priorParams[["mix"]]), 
 #                  utt = "generic")
  inputData = list(prior = list(params = data.frame(a1 = priorParams[["a_1"]],
                                                   b1 = priorParams[["b_1"]],
                                a2 = priorParams[["a_2"]],
                                                   b2 = priorParams[["b_2"]]),
                               mix = c(priorParams[["mixture_0"]],
                                       priorParams[["mixture_1"]],
                                       priorParams[["mixture_2"]])), 
                  utt = "generic")

 l0.rs.posterior <- webppl(paste(rsaHelpers, 
                                 uncertain.threshold.model.3comp, 
                                 uncertain.threshold.call, sep = '\n'), 
                           data = inputData, data_var = "data")

 l0.rs.prior <- webppl(paste(rsaHelpers, 
                             no.utterance.model.3comp, 
                             no.utterance.call, sep = '\n'), 
                       data = inputData, data_var = "data")

# l0.rs.posterior <- webppl(paste(rsaHelpers, uncertain.threshold.model, uncertain.threshold.call, sep = '\n'), data = inputData, data_var = "data")
 #l0.rs.prior <- webppl(paste(rsaHelpers, no.utterance.model, no.utterance.call, sep = '\n'), data = inputData, data_var = "data")
 gen.listener.predictions <- bind_rows(
   gen.listener.predictions, 
   l0.rs.prior %>% select(value) %>% mutate(property = p, Parameter = "state_Prior"),
    l0.rs.posterior %>% select(value) %>% mutate(property = p, Parameter = "state_Posterior")
   )
}

# m.gen.fullmodel.target.prevalence <- m.samp %>%
#   filter(type == "targetPrevalence") %>%
#   group_by(param, property, category) %>%
#   summarize(MAP = estimate_mode(val),
#             cred_upper = hdi_upper(val),
#             cred_lower = hdi_lower(val))

# m.gen.speakerBeliefs <- m.samp %>%
#   filter(type == "targetPrevalence") %>%
#   mutate(Sentence = paste(category, property)) %>%
#   filter(Sentence %in% example.generics) %>%
#   mutate(property = factor(property))

m.samp.i <- m.samp %>%
    filter(type == "targetPrevalence") %>%
    mutate(Sentence = paste(category, property)) %>%
    filter(Sentence %in% example.generics) %>%
    mutate(parameter = paste(param, property, category, component, sep = "_")) %>%
    select(-param, -property, -category, -component, -chain, -type) %>%
    group_by(parameter) %>%
    mutate(iteration = ave(parameter==parameter, parameter, FUN=cumsum)) %>%
    ungroup() %>%
    separate(parameter, into = c("parameter", "property", "category", "component"), sep= "_") %>%
    group_by(category, property, iteration) %>%
    spread(parameter, val) %>%
    rowwise() %>%
    mutate(
      a = mean*sampleSize,
      b = (1-mean)*sampleSize,
      val = rbeta(n = 1, shape1 = a, shape2 = b)
      )

m.samp.i.mixture <- m.samp.i %>% 
    filter(component == 1) %>%
    rowwise() %>%
    mutate(c = rbinom(size = 1, n = 1, prob = mix))

m.samp.ij <- left_join(
  m.samp.i %>%
    select(Sentence, property, category, iteration, component, val) %>%
    spread(component, val),
  m.samp.i.mixture %>% select(Sentence, property, category, iteration, c)
) %>% rowwise() %>%
  mutate(prev = ifelse(c == 1, `1`, `2`))

gen.inset.distributions <- bind_rows(
  gen.listener.predictions %>%
    mutate(category = NA),
  m.samp.ij %>%
    select(property, prev) %>%
    rename(value = prev) %>%
    mutate(Parameter = "speakerBeliefs")
)

category.text.labels <- data.frame(property = c("dont eat people", 
                      "carry malaria", "lay eggs",  "are female", "have spots"),
             category = c("Tigers", "Mosquitos", "Robins", "Robins", "Leopards"),
             x = c(0.3, 0.3, 0.47, 0.05, 0.6),
            y = c(0.45, 0.5, 0.45, 0.5, 0.26))

gen.inset.distributions.refactored <- gen.inset.distributions %>%
  mutate(Parameter = factor(Parameter, levels = c("state_Prior",
                                                  "state_Posterior",
                                                  "speakerBeliefs"),
                            labels = c("Prevalence Prior", 
                                       "Prevalence Posterior given generalization", 
                                       "Referent Prevalence")),
         property = fct_relevel(property,
                                "have spots", "lay eggs","carry malaria", 
                                "are female", "dont eat people") )
```

```{r vizualize insets, fig.width = 8, fig.height =4}
load("../paper/cached_results/case1_modelPosteriors.RData")

# save(d.gen.endorse.rsa.regression, gen.inset.distributions.refactored,
#      generics.fixed.noise.posterior, fixed.threshold.posterior,fixed.threshold.s1opt,uncertain.threshold.s1opt,
#      file = "../paper/cached_results/case1_modelPosteriors_moreFlexiblePrior.RData")

save(d.gen.endorse.rsa.regression, gen.inset.distributions.refactored,
     generics.fixed.noise.posterior, fixed.threshold.posterior,fixed.threshold.s1opt,uncertain.threshold.s1opt,
     file = "../paper/cached_results/case1_modelPosteriors.RData")

#load("../paper/cached_results/case1_modelPosteriors.RData")

generics.endorsement.insets <- ggplot(gen.inset.distributions.refactored, 
                                      aes( x = value, fill = Parameter, 
                 color = Parameter, lty = Parameter, alpha = Parameter ))+
  geom_density(aes(y = ..scaled..), adjust = 1.5, size = 1)+
  facet_wrap(~property, nrow = 1)+
  geom_label_repel(data = category.text.labels,
                  aes(label = category, x = x , y = y),
                  inherit.aes = F, color = "#2b83ba")+
    #scale_fill_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
    #scale_color_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
    scale_fill_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    scale_color_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    scale_alpha_manual(values = c(0.6, 0.4, 0))+
    #scale_linetype_manual(values = c(3, 4, 2, 1))+
    scale_linetype_manual(values = c(3, 4, 1))+
    #scale_x_continuous(breaks = c(0, 1), limits= c(0, 1))+
    #scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    xlab("Prevalence") +
    ylab("Scaled probability density")+
    theme(legend.position = "bottom", legend.title = element_blank())

generics.endorsement.insets
```

```{r}
for (p in example.generics.properties){
gen.inset.distributions.refactored %>%
    filter(property == p) %>%

    ggplot(. , aes( x = value, fill = Parameter, 
                 color = Parameter, lty = Parameter, alpha = Parameter ))+
  geom_density(aes(y = ..scaled..), adjust = 1, size = 1)+
  theme_black()+
  facet_wrap(~property, nrow = 1)+
  geom_label_repel(data = category.text.labels %>%
                     filter(property ==p),
                  aes(label = category, x = x , y = y),
                  inherit.aes = F, color = "#b2df8a", fill='black')+
    #scale_fill_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
    #scale_color_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
    scale_fill_manual(values = c("white", "#fb9a99", "#b2df8a"))+
    scale_color_manual(values = c("white", "#fb9a99", "#b2df8a"))+
    scale_alpha_manual(values = c(0.8, 0.5, 0))+
    #scale_linetype_manual(values = c(3, 4, 2, 1))+
    scale_linetype_manual(values = c(3, 4, 1))+
    scale_x_continuous(breaks = c(0, 1), limits= c(0, 1))+
    scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    xlab("Prevalence") +
    ylab("Scaled density")+
    theme(legend.position = "bottom", legend.title = element_blank())
  ggsave(paste(
    "~/Documents/research/talks/generics/mit-bcs-2018-03/figs/generics-insets-",
    p, ".pdf", sep = ""), width = 2.5, height = 2.5)
}

```

```{r generics-endorsement-figure, fig.width = 11, fig.height=8.5, fig.cap="Endorsing generalizations about categories. Top left: Human elicited endorsements for thirty generic sentences reveal a continuuum of endorsements. Top right: Model fits for the uncertain semantics speaker model (upper right), a fixed semantics speaker model (upper left), and regression models based on referent prevalence alone (lower left) and prevalence + cue validity (lower right). Bottom: Five example prevalence priors, listener posteriors upon hearing the generalization, and speaker belief distributions about the referent prevalence. These distributions are inferred using all three data sources from Expt. 1 (see Figure \\ref{fig:genericsModelDiagram} for an overview of the Bayesian data analytic approach).", cache = F}

fig.gen.endorse.data <- generics.endorsement.spectrum +
  theme(plot.margin = unit(c(6,2,0,0), "pt")) #+ ggtitle("prevalence priors")
fig.gen.endorse.models <- generics.endorsement.models +
  theme(plot.margin = unit(c(6,2,0,0), "pt")) #+ ylab("")
fig.gen.endorse.priors <- generics.endorsement.insets +
  theme(plot.margin = unit(c(26,10,0,10), "pt"))

#+ 
    #theme(strip.text.y = element_blank())#+ ggtitle("Interpreter model posterior")
#p4 <- s1.simulations.scatter + theme(plot.margin = unit(c(18,0,6,6), "pt"))

fig.gen.endorse.toprow <- plot_grid(fig.gen.endorse.data,
                                     fig.gen.endorse.models, 
                                    # align = 'v',
                                    #  axis = 'l',
                                    labels = c("A", "B"),
                                    rel_widths = c(0.9, 1))

plot_grid(fig.gen.endorse.toprow,
          fig.gen.endorse.priors,
          labels = c("", "C"), nrow = 2, rel_heights = c(1.1, 0.6)
          )
# prow <- plot_grid( p1 + theme(legend.position="none"),
#            p2 + theme(legend.position="none"),
#            p3 + theme(legend.position="none"),
#            p4,
#            align = 'vh',
#            labels = c("A", "B", "C"),
#            hjust = -1,
#            #nrow = 1,
#            #rel_widths = c(1.2, 3, 2)
#            rel_widths = c(1.2, 3, 1.05, 1.7)
#            )

# grid.arrange(generics.endorsement.spectrum, 
#              generics.endorsement.models,
#              generics.endorsement.insets, ncol = 2,
#              layout_matrix = cbind(c(1,1,3), c(2,2,3)))

```

# Old models

```{r generic-endorsement-priors-savageDickeyModel, eval = F}
savageDickeyModel <-'
var model = function(){
  var phi = uniformDrift({a:0, b: 1, width:0.2});
  var g = uniformDrift({a:0, b: 1, width:0.2});
  var d = uniformDrift({a:0, b: 100, width:5});
  var stableParams = betaShape({g, d})
  var stableComponent = Beta(stableParams)
  var transientComponent = Beta({a:1, b:100})
  mapData({data: preprocessedResponses}, function(d){
    factor( log(
      phi * exp(stableComponent.score(d)) +
      (1 - phi) * exp(transientComponent.score(d))
    ))
  })
  return {phi}
}

//var samples = data.samples[0], burn = data.burn[0], lag = data.lag[0];
// var samples = 2000, burn = burn / 2, lag = 0;
var modelPosterior = Infer({method: "MCMC", samples:10000, burn:5000, lag:0, 
verbose: true}, model)
var modelPrior = Infer({method: "forward", samples:10000}, model)

var pointNull = 1;

var logBFs = map(function(diff){
  var savageDickeyDenomenator = expectation(modelPrior, function(x){return Math.abs(x.phi-pointNull)<diff})
  var savageDickeyNumerator = expectation(modelPosterior, function(x){return Math.abs(x.phi-pointNull)<diff})
  var savageDickeyRatio = savageDickeyNumerator / savageDickeyDenomenator
  return [diff, Math.log(savageDickeyRatio)]
}, [0.01, 0.02, 0.03, 0.04, 0.05])

var returnVals = {
  prior: modelPrior, posterior: modelPosterior, logBFs: logBFs
}
returnVals
'
```

```{r generic-endorsement-priors-bayesFactors-supermodel, eval= F}
genericEndorsementPriorSupermodel <- '
var model = function(){
  var alpha = uniformDrift({a:0, b: 1, width:0.2});
  //var structured = flip();
  var phi = uniformDrift({a:0, b: 1, width:0.2});
  var g1 = uniformDrift({a:0, b: 1, width:0.2});
  var d1 = uniformDrift({a:0, b: 100, width:5});
  // var g0 = uniformDrift({a:0, b: 1, width:0.2});
  // var d0 = uniformDrift({a:0, b: 100, width:5});
  var stableParams1 = betaShape({g: g1, d: d1})//, 
      //stableParams0 = betaShape({g: g0, d: d0});
  var stableComponent1 = Beta(stableParams1)//,
      //stableComponent0 = Beta(stableParams0);

  var transientComponent = Beta({a:1, b:100})
  mapData({data:preprocessedResponses},
    function(d){
    var scr1 = log(phi * exp(stableComponent1.score(d)) +
                    (1 - phi) * exp(transientComponent.score(d)))
    var scr0 = stableComponent1.score(d);
    // display(exp(scr1) + " " + exp(scr0));
    var combined_ll = log(alpha * exp(scr1) + (1 - alpha) * exp(scr0));
    // factor(structured ? scr1 : scr0)
    factor(combined_ll)
  })
  return alpha
}
'
```

```{r generic-endorsement-priors-bayesFactor-supermodelRun, cache = T, eval = F}
for (p in gen.endorse.properties){
  priorData <- filter(d.gen.endorse.priors, Property == p)$prevalence

  m.gen.endorse.priors.supermodel.posterior <- webppl(
    paste(genericEndorsementPriorModelHelpers, genericEndorsementPriorSupermodel, sep = '\n'), 
    data = priorData, data_var = "data",
    model_var = "model",  chains = 2,
                                 cores = 2,
                                 inference_opts = 
                                   list(method = "MCMC",
                                        samples = 2000, 
                                        burn = 2000, 
                                        lag = 2, 
                                        verbose = T)
    )

  qplot(m.gen.endorse.priors.supermodel.posterior$value)
}
```

```{r generic-endorsement-priors-bayesFactors-enumerate, eval = F}
genericEndorsementPriorModelPriorLikelihoodEnumerate <- '
var eps = 0.00000001;

var model = function(){
  var phi = uniformDraw(_.range(eps, 1-eps, 0.1))
  var g = uniformDraw(_.range(eps, 1-eps, 0.05))
  var d = uniformDraw(_.range(eps, 100, 2))
  var stableParams = betaShape({g, d})
  var stableComponent = Beta(stableParams)
  var transientComponent = Beta({a:1, b:100})
  var logLike = sum(map(function(d){
    var scr = log(
      phi * exp(stableComponent.score(d)) +
      (1 - phi) * exp(transientComponent.score(d))
    )
    return scr
  }, preprocessedResponses))
  return logLike
}
'
genericEndorsementUnstructuredModelPriorLikelihoodEnumerate <- '
var eps = 0.00000001;

var model = function(){
  var g = uniformDraw(_.range(eps, 1-eps, 0.2))
  var d = uniformDraw(_.range(eps, 100, 2))
  var stableParams = betaShape({g, d});
  var stableComponent = Beta(stableParams);
  var logLike = sum(map(function(d){
    return stableComponent.score(d)
  }, preprocessedResponses))
  return logLike
}
'
```

```{r generic-endorsement-priors-bayesFactorsRunEnumerate, cache = T, eval =F}
marginal.prior.likelihoods <- data.frame()

for (p in gen.endorse.properties){
  priorData <- filter(d.gen.endorse.priors, Property == p)$prevalence

  m.gen.endorse.priors.logLike.enum1 <- webppl(
    paste(genericEndorsementPriorModelHelpers,
          genericEndorsementPriorModelPriorLikelihoodEnumerate, sep = '\n'), 
    data = priorData, data_var = "data",
    model_var = "model", 
    inference_opts = list(method = "enumerate")
    )

  m.gen.endorse.priors.logLike.enum0 <- webppl(
    paste(genericEndorsementPriorModelHelpers,
          genericEndorsementUnstructuredModelPriorLikelihoodEnumerate, sep = '\n'), 
    data = priorData, data_var = "data",
    model_var = "model", 
    inference_opts = list(method = "enumerate")
    )

  marginal.prior.likelihood1 <- logmeanexp(m.gen.endorse.priors.logLike.enum1$support)
  marginal.prior.likelihood0 <- logmeanexp(m.gen.endorse.priors.logLike.enum0$support)

  marginal.prior.likelihoods <- bind_rows(
    marginal.prior.likelihoods, 
    data.frame(Property = p, structuredLL = marginal.prior.likelihood1, unstructuredLL = marginal.prior.likelihood0)
  )
}

marginal.prior.likelihoods <- marginal.prior.likelihoods %>% mutate(logBF = structuredLL - unstructuredLL, BF = exp(logBF))

```

```{r testingNewViz}

d.prev.zeros <- d.prev %>%
  mutate(nonZero = prevalence > 0) %>%
  group_by(Property) %>%
  tidyboot_mean(column = nonZero)

d.prev.zeros %>% 
  mutate(Property = factor(Property, levels = Property ))
  ggplot(., aes( x = Property, y = mean, ymin = ci_lower, ymax = ci_upper))+
  geom_col()

ggplot(d.prev %>%
         filter(prevalence > 0), 
           aes( x = prevalence))+
           #aes( x = prevalence, color = src, lty = src))+
    geom_density(aes(y = ..scaled..), adjust = 0.5, size = 1)+
    #scale_fill_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    #stat_ecdf(size = 1)+
    #scale_color_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    #scale_alpha_manual(values = c(0.6, 0.4, 0))+
    facet_wrap(~Property)+
    #scale_linetype_manual(values = c(1, 4, 3))+
    scale_x_continuous(breaks = c(0, 100), limits= c(0, 100))+
    scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    #coord_fixed()+
    xlab("") +
    ylab("Normalized probability density")+
    theme(legend.position = "bottom", legend.title = element_blank())
```

```{r generic-endorsement-priors-model-run, eval = F}
m.gen.endorse.priors.posteriorPredictive <- data.frame()
n_samples <- 20000
lg = 0
# example.generics.prior.properties <- c(
#   "dont eat people", "have beautiful feathers",
#   "have wings", "are red",
#   "carry malaria", "lay eggs", 
#   "are female", "have spots"
#   )

for (p in levels(d.gen.endorse.priors$Property)) {
  priorData <- filter(d.gen.endorse.priors, Property == p)$prevalence

    m.gen.endorse.priors.structured <- webppl(
      paste(genericEndorsementPriorModelHelpers, structuredPriorModel, sep = ""),
                                 data = priorData, data_var = "data",
                                 model_var = "model",
                                 chains = 2,
                                 cores = 2,
                                 inference_opts =
                                   list(method = "MCMC",
                                        samples = n_samples,
                                        burn = n_samples / 2,
                                        lag = lg,
                                        verbose = T))

        # m.gen.endorse.priors.structured.predictive <- m.gen.endorse.priors.structured %>%
    #   group_by(Iteration, Chain) %>%
    #   spread(Parameter, value) %>%
    #   rowwise() %>%
    #   mutate(
    #     a = g*d,
    #     b = (1-g)*d,
    #     isPresent = rbinom(1, 1, prob = phi),
    #     prevalence = ifelse(isPresent == 1, 
    #                        rbeta(n = 1, 
    #                              shape1 = a,
    #                              shape2 = b),
    #                        rbeta(n = 1,
    #                              shape1 = 1,
    #                              shape2 = 100)),
    #     Property = p, src = "Structured"
    #   ) %>% ungroup()
    # 
    m.gen.endorse.priors.unstructured <- webppl(
      paste(genericEndorsementPriorModelHelpers, unstructuredPriorModel, sep = ""),
                                 data = priorData, data_var = "data",
                                 model_var = "model",
                                 chains = 2,
                                 cores = 2,
                                 inference_opts =
                                   list(method = "MCMC",
                                        samples = n_samples,
                                        burn = n_samples / 2,
                                        lag = lg,
                                        verbose = T))

    m.gen.endorse.priors.unstructured.predictive <- m.gen.endorse.priors.unstructured %>%
      group_by(Iteration, Chain) %>%
      spread(Parameter, value) %>%
      rowwise() %>%
      mutate(
        a = g*d,
        b = (1-g)*d,
        prevalence = rbeta(n = 1, 
                                 shape1 = a,
                                 shape2 = b),
        Property = p, src = "Unstructured"
      ) %>% ungroup()

    m.gen.endorse.priors.posteriorPredictive <- bind_rows(
      m.gen.endorse.priors.posteriorPredictive,
      bind_rows(
       m.gen.endorse.priors.structured.predictive %>%
          select( src, Property, prevalence),
      m.gen.endorse.priors.unstructured.predictive %>%
          select( src, Property, prevalence),
      data.frame(Property = p,
                 src = "Data",
                 prevalence = priorData/100)
      )
    )

}
# save(m.gen.endorse.priors.posteriorPredictive, file = '../analysis/generics-priors-posteriorPredictives.Rdata')

```

## Appendix C: Comparison of prior parameters inferred using only prior data and in joint-inference model

### Prevalence prior parameters

```{r generic-priors-parameters, eval = F}
m.gen.endorse.priors.posteriorPredictive <- data.frame()
n_samples <- 20000
lg = 0

m.gen.endorse.priors.parameters <- data.frame()
for (p in levels(d.gen.endorse.priors$Property)) {
  priorData <- filter(d.gen.endorse.priors, Property == p)$prevalence

    m.gen.endorse.priors.structured <- webppl(
      paste(genericEndorsementPriorModelHelpers, structuredPriorModel, sep = ""),
                                 data = priorData, data_var = "data",
                                 model_var = "model",
                                 chains = 2,
                                 cores = 2,
                                 inference_opts =
                                   list(method = "MCMC",
                                        samples = n_samples,
                                        burn = n_samples / 2,
                                        lag = lg,
                                        verbose = T))
m.gen.endorse.priors.structured.summary <- m.gen.endorse.priors.structured %>%
  group_by(Parameter) %>%
  summarize(MAP = estimate_mode(value),
            cred_upper = hdi_upper(value),
            cred_lower = hdi_lower(value))

    m.gen.endorse.priors.parameters <- bind_rows(
      m.gen.endorse.priors.parameters,
      m.gen.endorse.priors.structured.summary %>%
        mutate(Property = p)
      )
}

# save(m.gen.endorse.priors.parameters, file = '../analysis/generics-priors-posteriorParameters.Rdata')
```

```{r}
load(file = paste(project.path,  "models/generics/results/", "results-generics-jointModel-S1-smntcs_generic-100000_burn50000_lag20_3chains.RData", sep = ""))

m.gen.endorse.joint.prior.parameters <- m.samp %>%
  filter(type %in% c("prior", "targetPrevalence")) %>%
  group_by(property, category, type, param) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))
```

```{r}
load( file = '../analysis/generics-priors-posteriorParameters.Rdata')
m.gen.endorse.priors.parameters <- left_join(
  m.gen.endorse.priors.parameters,
  m.gen.endorse.joint.prior.parameters %>%
    filter(type == "prior") %>%
    rowwise() %>%
    mutate(Parameter = ifelse(param == "mixture", "phi",
                              ifelse(category == "mean", "g", "d"))) %>%
    ungroup() %>% 
    select(property, Parameter, MAP, cred_upper, cred_lower) %>%
    rename(Property = property,
           MAP_joint = MAP,
           cred_upper_joint = cred_upper,
           cred_lower_joint = cred_lower)
)

ggplot(m.gen.endorse.priors.parameters, aes( x = MAP, xmin = cred_lower, xmax = cred_upper,
                 y = MAP_joint, ymin = cred_lower_joint, ymax = cred_upper_joint ))+
  geom_point()+
  geom_errorbar(alpha = 0.3)+geom_errorbarh(alpha = 0.3)+
  facet_wrap(~Parameter, scales = 'free', labeller = label_parsed)+
  xlab("Inferred values with only prevalence data")+
  ylab("Inferred values in joint-inference model")
```

### Referent prevalence parameters

```{r generic-refprev-parameters, eval = F}
m.gen.endorse.refprev.posteriorPredictive <- data.frame()
n_samples <- 20000
lg = 0

m.gen.endorse.refprev.parameters <- data.frame()

for (p in levels(d.gen.endorse.priors$Property)) {
  categories <- d.gen.endorse.bayes %>% filter(Property == p) %>% distinct(Category)

  for (k in categories$Category) {

    priorData <- filter(d.gen.endorse.priors, Property == p, Category == k)$prevalence

    m.gen.refprev <- webppl(
      paste(genericEndorsementPriorModelHelpers, unstructuredPriorModel, sep = ""),
                                 data = priorData, data_var = "data",
                                 model_var = "model",
                                 chains = 2,
                                 cores = 2,
                                 inference_opts =
                                   list(method = "MCMC",
                                        samples = n_samples,
                                        burn = n_samples / 2,
                                        lag = lg,
                                        verbose = T))

  m.gen.refprev.summary <- m.gen.refprev %>%
    group_by(Parameter) %>%
    summarize(MAP = estimate_mode(value),
              cred_upper = hdi_upper(value),
              cred_lower = hdi_lower(value))

    m.gen.endorse.refprev.parameters <- bind_rows(
      m.gen.endorse.refprev.parameters,
      m.gen.refprev.summary %>%
        mutate(Property = p, Category = k)
      )
    print(k)
  }
  print(p)
}

 #save(m.gen.endorse.refprev.parameters, file = '../analysis/generics-refprev-posteriorParameters.Rdata')
```

```{r}
load(file = '../analysis/generics-refprev-posteriorParameters.Rdata')

m.gen.endorse.refprev.parameters <- left_join(
  m.gen.endorse.refprev.parameters,
  m.gen.endorse.joint.prior.parameters %>%
    filter(type == "targetPrevalence") %>%
    rowwise() %>%
    mutate(Parameter = ifelse(param == "mean", "g", "d")) %>%
    ungroup() %>% 
    select(property, category, Parameter, MAP, cred_upper, cred_lower) %>%
    rename(Property = property,
           Category = category,
           MAP_joint = MAP,
           cred_upper_joint = cred_upper,
           cred_lower_joint = cred_lower)
)

ggplot(m.gen.endorse.refprev.parameters, aes( x = MAP, xmin = cred_lower, xmax = cred_upper,
                 y = MAP_joint, ymin = cred_lower_joint, ymax = cred_upper_joint ))+
  geom_point()+
  geom_errorbar(alpha = 0.3)+geom_errorbarh(alpha = 0.3)+
  facet_wrap(~Parameter, scales = 'free', labeller = label_parsed)+
  xlab("Inferred parameters using only prevalence data")+
  ylab("Inferred parameters in joint-inference model")
```

```{r}
save(m.gen.endorse.priors.parameters, m.gen.endorse.refprev.parameters,
     file = "../paper/cached_results/case1_prevParamDistortion.RData")
```