### 2018-10-11 analysis/model-simulations.Rmd
add elephants model

---
title: "Lang-gen model simulations"
output: html_notebook
---

```{r libraries, cache = F}
library(papaja)
library(formatR)
library(rwebppl)
library(xtable)
library(tidyverse)
library(forcats)
library(langcog)
library(coda)
library(ggthemes)
library(ggrepel)
library(jsonlite)
library(gridExtra)
library(lme4)
library(knitr)
library(kableExtra)
library(cowplot)
library(magick)
theme_set(theme_few())
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
logmeanexp <- function(x){
  x.num <- as.numeric(x)
  xstar = max(x.num)
  return(xstar + log(mean(exp(x.num - xstar))))
}

compute_r2 <- function(df,v1, v2, sigfigs = 3){
  return(format(cor(df[[v1]], df[[v2]])^2, digits = sigfigs))
}

compute_mse <- function(df, v1, v2, sigfigs = 3){
  return(format(mean( (df[[v1]]-df[[v2]])^2), digits = sigfigs))
}

project.path <- "../"
options("scipen"=10) 
```

# RSA model

```{r rsaHelpers}
rsaHelpers <- '
var probability = function(Dist, x) {
    return Math.exp(Dist.score(x));
}
var targetUtterance = "generic";

var round = function(x){
  return Math.round(x*1000)/1000
}
var utterancePrior = Infer({model: function(){
  return uniformDraw([targetUtterance,"silence"])
}});

var thetaBins = map(round, _.range(0.01, 0.98, 0.01))
// var thetaBins = [
//    0.01, 0.05,0.1,0.15,0.2,0.25,0.3,0.35,0.4,0.45,
//    0.5, 0.55,0.6,0.65,0.7,0.75,0.8,0.85,0.9,0.95
 // ]
var thetaPrior = Infer({model: function(){
 return uniformDraw(thetaBins)
}});

var bins = map(round, _.range(0.01, 0.99, 0.01))

// var bins = [
  // 0.01,0.05,0.1,0.15,0.2,0.25,0.3,0.35,0.4,0.45,
  // 0.5,0.55,0.6,0.65,0.7,0.75,0.8,0.85,0.9,0.95,0.99
// ];

var meaning = function(utt,state, theta) {
  return utt=="generic"? state > theta :
         utt=="generic is false"? state<=theta :
         utt=="silence"? true :
         utt=="some"? state>0.01:
         utt=="most"? state> 0.5:
         utt=="all"? state >= 0.99:
         true
}

var mixture = data.prior.mix[0];
var priorParams = data.prior.params[0];

var statePrior = Infer({model: function(){
  var component = flip(mixture);
  return component ?
    categorical({
      vs: bins,
      ps: map(function(b) {
        return probability(Beta(priorParams), b) + Number.EPSILON
      }, bins )
    }) :
    categorical({
      vs: bins,
      ps: map(function(b) {
        return probability(Beta({a:1,b:50}), b) + Number.EPSILON
      }, bins )
    })
}});
'
```

```{r rsaInterpretationModels}
no.utterance.model <- '
var noUtteranceInterpreter = function() {
  Infer({model: function(){
   var state = sample(Beta( flip(mixture) ? priorParams : {a:1,b:100}))
    return { state }
 }, method: "rejection", samples: 20000, burn:5000, verbose: T})}
'

fixed.threshold.model <- '
var fixedThresholdInterpreter = function(threshold) {
  Infer({model: function(){
  // var state = sample(Beta( flip(mixture) ? priorParams : {a:1,b:100}))
  var state = sample(statePrior)
  condition( state > threshold)
    return {
      state: state
  }
 // }, method: "rejection", samples: 20000, burn:5000, verbose: T})}
 }, method: "enumerate"})}
'

uncertain.threshold.model <- '
var uncertainThresholdInterpreter = function() {
  Infer({model: function(){
   var state = sample(Beta( flip(mixture) ? priorParams : {a:1,b:100}))
    factor( Math.log(state) )
    return {
      state: state, 
  }
  }, method: "rejection", samples: 20000, burn:5000, verbose: T})}
'

```

```{r simulationCalls}
fixed.threshold.calls <- '
_.fromPairs(map(function(t){return [t, fixedThresholdInterpreter(t)]}, [0.1, 0.3, 0.5, 0.7, 0.9]))
'
no.utterance.call <- ' 
  noUtteranceInterpreter()
'
uncertain.threshold.call <- '
  uncertainThresholdInterpreter()
'
```

```{r simulationRuns}
priorNames <- c(
  "doesntEatPeople", #"uniformRare", 
  "isFemale", 
  "laysEggs", 
  #"eatsFood", 
  "barks",
  "carriesMalaria",
  "hasSpots"
 # "areSick"
  )
priorShapes <- list(
  isFemale =  list(params = data.frame(a = 10, b = 10), mix = 1), 
  laysEggs =  list(params = data.frame(a = 10, b = 10), mix = 0.4), 
  doesntEatPeople =  list(params = data.frame(a = 50, b = 1), mix = 1), 
  barks =  list(params = data.frame(a = 5, b = 1), mix = 0.4), 
  hasSpots =  list(params = data.frame(a = 5, b = 1), mix = 0.7), 
  carriesMalaria =  list(params = data.frame(a = 2, b = 10), mix = 0.4)
)

sims.fixed.thresholds <- data.frame()
sims.priors <- data.frame()
sims.uncertain.thresholds <- data.frame()

for (p in priorNames){

  inputData = list(prior = priorShapes[[p]])
  rs.fixed.threshold.model <- webppl(paste(rsaHelpers, fixed.threshold.model, fixed.threshold.calls, sep = '\n'), data = inputData, data_var = "data")
  rs.uncertain.threshold.model <- webppl(paste(rsaHelpers, uncertain.threshold.model, uncertain.threshold.call, sep = '\n'), data = inputData, data_var = "data")
  rs.prior <- webppl(paste(rsaHelpers, no.utterance.model, no.utterance.call, sep = '\n'), data = inputData, data_var = "data")

  sims.fixed.thresholds <- bind_rows(sims.fixed.thresholds,
    bind_rows(
      mutate(get_samples(data.frame(rs.fixed.threshold.model$`0.1`) %>% rename(prob = probs), 20000), threshold = 0.1) ,
      mutate(get_samples(data.frame(rs.fixed.threshold.model$`0.3`) %>% rename(prob = probs), 20000), threshold = 0.3),
      mutate(get_samples(data.frame(rs.fixed.threshold.model$`0.5`) %>% rename(prob = probs), 20000), threshold = 0.5),
      mutate(get_samples(data.frame(rs.fixed.threshold.model$`0.7`) %>% rename(prob = probs), 20000), threshold = 0.7),
      mutate(get_samples(data.frame(rs.fixed.threshold.model$`0.9`) %>% rename(prob = probs), 20000), threshold = 0.9)
    ) %>% 
      mutate(PriorShape = p)
    )

  sims.uncertain.thresholds <- bind_rows(sims.uncertain.thresholds, rs.uncertain.threshold.model %>% select(value) %>% rename(state = value) %>% mutate(PriorShape = p))
  sims.priors <- bind_rows(sims.priors, rs.prior %>% select(value) %>% rename(state = value) %>% mutate(PriorShape = p))
}
```

```{r simluationEndorsementModelRuns}
s1.model <- '
var interpreter = cache(function(utterance) {
  Infer({model: function(){
    var state = sample(statePrior)
    var theta = utterance == "generic" ? sample(thetaPrior) : -99
    condition(meaning(utterance, state, theta))
    return state
 }})}, 10000)

var alpha = 2

var endorser = function(featureProb){
  Infer({model: function(){
    var endorsement = sample(utterancePrior)
    var L0 = interpreter(endorsement)
    factor(alpha * L0.score(featureProb))
    return endorsement
  }})
}
// console.log(data.featureProb[0])
probability(endorser(data.featureProb[0]), "generic")
'

exampleGenerics <- c(
  "sharksDontEatPeople",
  "robinsAreFemale", 
  "robinsLayEggs", 
  "kangaroosHaveSpots",
  "dogsBark",
  "mosquitosCarryMalaria"
  )

exampleParameters <- list(
  robinsAreFemale =  list(prior = list(params = data.frame(a = 10, b = 10), mix = 1), featureProb = 0.5), 
  robinsLayEggs =  list(prior = list(params = data.frame(a = 10, b = 10), mix = 0.2), featureProb = 0.5), 
  sharksDontEatPeople =  list(prior = list(params = data.frame(a = 50, b = 1), mix = 1), featureProb = 0.8), 
  dogsBark =  list(prior = list(params = data.frame(a = 5, b = 1), mix = 0.4), featureProb = 0.9), 
  kangaroosHaveSpots =  list(prior = list(params = data.frame(a = 5, b = 1), mix = 0.7), featureProb = 0.1), 
  mosquitosCarryMalaria =  list(prior = list(params = data.frame(a = 1, b = 30), mix = 0.1), featureProb = 0.1)
)

s1.simulations <- data.frame()
for (p in exampleGenerics){

    s1.rs <- webppl(paste(rsaHelpers, s1.model, sep = "\n"), data = exampleParameters[[p]], data_var = "data")
    s1.simulations <- bind_rows(s1.simulations,
                                   data.frame(example = p, endorsement = s1.rs, featureProb = exampleParameters[[p]]["featureProb"] ))

}

s1.simulations.relabeled <- s1.simulations %>%
  mutate(example = factor(example, 
                          levels = c("dogsBark", "kangaroosHaveSpots", "sharksDontEatPeople", "robinsLayEggs", "robinsAreFemale", "mosquitosCarryMalaria"),
                            labels = 
                            c("Dogs bark", "Kangaroos have spots", "Sharks don't eat people", "Robins lay eggs", "Robins are female", "Mosquitos carry malaria")
                            #c("Dogs", "Kangaroos", "Sharks", "Robins ", "Robins", "Mosquitos")
                          ))

```

```{r saveResults}
#save(sims.priors, sims.uncertain.thresholds, sims.fixed.thresholds,s1.simulations.relabeled, file = "../paper/cached_results/modelSims-priors_fixedT_uncertainT_speaker.RData")
```

# Schematic (uniform prior) figure

```{r schematicModel}
schematicModel <- '
var round = function(x){
  return Math.round(x*100)/100
}

var isNegation = function(utt){
  return (utt.split("_")[0] == "not")
};

var avoidEnds = function(x){
  return x >= 1 ? 0.99 : x == 0 ? 0.01 : x
}

var lb = 0, ub = 1, diff = 0.05;
var bins = _.range(lb, ub + diff, diff)

var DiscreteBeta = function(a, b){
  Infer({model: function(){
    categorical({
      vs:bins,
      ps:map(function(x){
        Math.exp(Beta({a, b}).score(avoidEnds(x)))
      }, bins)
    })
  }})
}

var utterances = [
  "gen",
  "null"
];

var utterancePrior = Infer({model: function(){
  return uniformDraw(utterances)
}});

var meaning = function(words, state, threshold){
  return words == "gen" ? state > threshold :
  true
};

var speakerOptimality = 3;
var speakerOptimality2 = 1;

var fixedThetalistener0 = cache(function(utterance, threshold) {
  Infer({model: function(){
    var state = sample(DiscreteBeta(1, 1));
    var m = meaning(utterance, state, threshold);
    condition(m);
    return state;
  }})
}, 10000);

var greaterThanThresholdBins = _.range(lb, ub, diff)
var lessThanThresholdBins = _.range(lb+diff, ub+diff, diff)

var litInterpreter = cache(function(utterance) {
  Infer({model: function(){
    var threshold = uniformDraw(greaterThanThresholdBins)

    var state = sample(DiscreteBeta(1, 1));
    var m = meaning(utterance, state, threshold);
    condition(m);
    return state
  }})
}, 10000);

var speaker1 = cache(function(state) {
  Infer({model: function(){
    var utterance = sample(utterancePrior);
    var L0 = litInterpreter(utterance);
    factor(speakerOptimality*L0.score(state));
    return utterance;
  }})
}, 10000);
'

listenerCall <- '
_.fromPairs(map(function(u){
  var post = litInterpreter(u)
  return [u, post]
}, utterances))
'

speakerCall <- '
_.flatten(
    map(function(s){
      var speakProbs = speaker1(s)
       return {  
          state: s,
          "gen": Math.exp(speakProbs.score("gen")),
          "null": Math.exp(speakProbs.score("null"))
        }
    }, bins)
)
'

literalListenerCall <- '
_.flatten(_.flatten(
map(function(tH){
    var l0posterior = fixedThetalistener0("gen", tH)
    display(expectation(l0posterior))
    map(function(s){
       return {  
          state: s,
          theta:tH, 
          literalPosterior: Math.exp(l0posterior.score(s))
        }
    }, bins)
}, greaterThanThresholdBins)
))
'
```

### Fixed thresholds

```{r schematicFixed}
rs.wp.l0 <- webppl(paste(schematicModel, literalListenerCall,  sep = '\n'))

fig.l0.thresholds <- ggplot(rs.wp.l0, aes( x = state, 
                                           y = literalPosterior, 
                                       group = theta))+
  geom_line(size = 2)+
  geom_vline(aes(xintercept = theta), 
             color = 'darkred', size = 2)+
  #scale_color_solarized()+
  scale_x_continuous(breaks = c(0, 0.5, 1))+
  scale_y_continuous(breaks = c(0, 0.5, 1))+
  #scale_color_continuous(limits = c(0, 1), breaks = c(0, 1))+
  ggtitle("truth-functional threshold = ")+
  ylab("Literal listener posterior probability")+
  xlab("Degree of happiness")+
  facet_wrap(~theta)+
  theme(text = element_text(size = 16))
```

Integrate out threshold (full interpretation model)

```{r schematicUncertain}
rs.listener.wp <- webppl(paste(schematicModel, listenerCall,  sep = '\n'))

rs.listener.wp.tidy <- bind_rows(
  data.frame(rs.listener.wp$gen) %>% 
    mutate(utterance = "gen"),
  data.frame(rs.listener.wp$null) %>% 
    mutate(utterance = "null")
)

rs.listener.wp.tidy.samples <- get_samples(
  rs.listener.wp.tidy %>% rename(prob = probs), 10000)

ggplot(rs.listener.wp.tidy.samples, 
       aes( x = support,fill = utterance, color = utterance))+
  geom_density(alpha = 0.4, size = 1.3)+
  scale_fill_solarized()+
  scale_color_solarized()+
  xlab("prevalence")+
  ylab("Posterior probability density")+
  scale_x_continuous(breaks =c(0, 1))+
  scale_y_continuous(breaks = c(0, 2))+
  #guides(fill = F, color = F)+
  ggtitle("prevalence posterior")

#ggsave("figs/lassiterL1_posteriors_wCost1,5_alpha1.png", width = 6, height = 4)
```

```{r schematicEndorsement}
rs.wp <- webppl(paste(schematicModel, speakerCall,  sep = '\n'))

rs.tidy <- data.frame(rs.wp) %>%
  gather(utt, prob, -state)

fig.thresholds <- ggplot(rs.tidy %>%
                           mutate(utt = factor(utt,
                                               levels=c("null","gen"))), 
                         aes( x = state, y = prob, fill = utt))+
  geom_col( 
           color = 'black')+
  #facet_wrap(~happy_theta)+
  scale_x_continuous(breaks = c(0, 0.5, 1))+
  scale_y_continuous(breaks = c(0, 0.5, 1))+
  ggtitle("truth-functional threshold = ")+
  ylab("Speaker probability of saying 'happy'")+
  xlab("Degree of happiness")+
  theme(text = element_text(size = 16))+
  scale_fill_solarized()

fig.thresholds
#ggsave(fig.thresholds, 
 #      file = "figs/lassiterS1_uttXstateXtheta_wCost.pdf", width = 7, height = 7)
```

```{r}
save(rs.wp.l0, rs.listener.wp.tidy.samples, rs.tidy,
     file = "../paper/cached_results/modelSims-uniform_fixedT_uncertainT_speaker.RData")
```

## Elephants live in africa and asia

```{r}
elephants <- '
var probability = function(Dist, x) {
  return Math.exp(Dist.score(x));
}

var bins = [
  0.01,0.05,0.1,0.15,0.2,0.25,0.3,0.35,0.4,0.45,
  0.5,0.55,0.6,0.65,0.7,0.75,0.8,0.85,0.9,0.95,0.99
];

var theta = 1;
var params75 = {a:45, b: 15};
var params50 = {a:30, b: 30};
var params25 = {a:15, b: 45};
var params100 = {a:50, b: 1};

var allComponents = [params25, params50, params75, params100]
var nullParams = {a:1,b:2000};

var makeComponent = function(params){
  return Categorical({
  vs: bins,
  ps: map(function(b) {
    probability(Beta(params), b) + Number.EPSILON
  }, bins )
})
}

var statePrior = Infer({model: function(){
  var mainPlace = categorical({
    ps:[1, 1, 1, 1],
    vs:[
      {asia: true, africa: false}, {asia: false, africa: true},
      {asia: false, africa: false}, {asia: true, africa: true}
    ]
  });
  return {
    asia: sample(makeComponent(mainPlace.asia ? uniformDraw(allComponents) : nullParams)),
    africa: sample(makeComponent(mainPlace.africa ? uniformDraw(allComponents) : nullParams))
  }
}})

var thetaPrior = Infer({model: function(){
  return {
    asia: uniformDraw(
      [0.01, 0.05,0.1,0.15,0.2,0.25,0.3,0.35,0.4,0.45,
       0.5, 0.55,0.6,0.65,0.7,0.75,0.8,0.85,0.9,0.95]
    ),
    africa: uniformDraw(
      [0.01, 0.05,0.1,0.15,0.2,0.25,0.3,0.35,0.4,0.45,
       0.5, 0.55,0.6,0.65,0.7,0.75,0.8,0.85,0.9,0.95]
    )
  }
}});

var meaning = function(utt, state, theta) {
  state[utt] > theta[utt]
}

var listener0 = function(utterance, worldPriors, thresholdPriors) {
  Infer({model: function(){
    var state = sample(worldPriors)
    var theta = sample(thresholdPriors)
    var m = meaning(utterance, state, theta)
    condition(m)
    condition(state.africa + state.asia <= 1)
    return {state, theta}
  }})
}

var elephantsLiveInAsia = listener0("asia", statePrior, thetaPrior);
var asiaStatePosterior = marginalize(elephantsLiveInAsia,"state");
var asiaThresholdPosterior =  marginalize(elephantsLiveInAsia, "theta");

var elephantsLiveinAfricaAndAsia = listener0("africa", asiaStatePosterior, asiaThresholdPosterior)

var AfricaAndAsiaStatePosterior = marginalize(elephantsLiveinAfricaAndAsia, "state");

var AfricaAndAsiaThresholdPosterior =  marginalize(elephantsLiveinAfricaAndAsia, "theta");

var x = {asiaStatePosterior, AfricaAndAsiaStatePosterior}
x
'
```

```{r}
rs.eleph <- webppl(elephants)

rs.asia <- as.data.frame(rs.eleph$asiaStatePosterior)
rs.asiaafrica <- as.data.frame(rs.eleph$AfricaAndAsiaStatePosterior)

rs.asia.marignals <- bind_rows(
  rs.asia %>%
    rename(state = support.asia) %>%
    group_by(state) %>%
    summarize(prob = sum(probs)) %>%
    mutate(marginal = "asia"),
  rs.asia %>%
    rename(state = support.africa) %>%
    group_by(state) %>%
    summarize(prob = sum(probs)) %>%
    mutate(marginal = "africa")
)

ggplot(rs.asia.marignals, aes(x = state, y = prob))+
  geom_col(position = position_dodge())+
  facet_wrap(~marginal)
```
```{r}
rs.asiaafrica.marignals <- bind_rows(
  rs.asiaafrica %>%
    rename(state = support.asia) %>%
    group_by(state) %>%
    summarize(prob = sum(probs)) %>%
    mutate(marginal = "asia"),
  rs.asiaafrica %>%
    rename(state = support.africa) %>%
    group_by(state) %>%
    summarize(prob = sum(probs)) %>%
    mutate(marginal = "africa")
)

ggplot(rs.asiaafrica.marignals, aes(x = state, y = prob))+
  geom_col(position = position_dodge())+
  facet_wrap(~marginal)
```

# Misc extra

old scatter plot version of endorsements
```{r}
#s1.simulations.scatter <-

  ggplot(s1.simulations.relabeled, aes( x = featureProb, y = endorsement))+
  #geom_jitter(width = 0.1, shape = 21, size = 5)+
  #ylab("Human endorsement probability")+
  #xlab("Frequency of target")+
  geom_point()+
  #geom_label_repel(data = s1.simulations.relabeled,
   #                aes( label = example, fill = example),
  #                     color = 'white',
    #box.padding = unit(0.0, "lines"),
    #point.padding = unit(1e-06, "lines"),
    #segment.color = 'grey37')+
  scale_fill_solarized()+
  scale_x_continuous(limits = c(0, 1),
                     breaks = c(0, 0.5, 1))+
  scale_y_continuous(limits = c(-0.01, 1.01),
                     breaks = c(0, 1),
                     labels = c("False", "True"))+
  coord_fixed(ratio = 1)+
  guides(fill = F)+
  ylab("Endorsement model prediction")+
  xlab("Hypothesized prevalence")

```

```{r eval = F}
all.simulations.refactored <- all.simulations %>%
  mutate(Distribution = paste(Distribution, utterance, sep = "_")) %>%
  filter(Distribution != "Some", Distribution != "Most") %>%
  mutate(Distribution = factor(Distribution, levels = 
                                c( "Prior_generic", "Posterior_Generalization_generic"),
                                labels = c( "Prior/Silence", "Generalization")
                             # c("Prior_some",  "Posterior_Generalization_generic", "Posterior_Generalization_some"),
                             #    labels = c("Prior/Silence",  "Generalization (Uncertain threshold)", "Fixed threshold")
                             ),
         PriorShape = factor(PriorShape, 
                             levels = c("barks", "hasSpots", "doesntEatPeople", "laysEggs", "isFemale", "carriesMalaria"),
                            labels = c("bark", "have spots", "don't eat people", "lay eggs", "are female", "carry malaria")
                            # levels = c("uniform", "barks","laysEggs","carriesMalaria","isFemale"),
                             #labels = c("uniform", "bark","lay eggs","carry malaria","are female")
                                  #"eatsFood","areSick", "uniformRare", 
                                  #"carriesMalaria"
                                  )
         )

l0.simulations.distributions <- ggplot(all.simulations.refactored, aes(x = value, #y = prob, 
                #lty = Distribution, 
                color = PriorShape, fill = PriorShape, group = Distribution, alpha = Distribution, linetype = Distribution))+
    geom_density(aes(y = ..scaled..),
                 size = 0.6,  adjust = 1.5)+
    #geom_bar(stat = 'identity', position = position_dodge(), color = 'black')+
    theme_few() +
    scale_color_solarized()+
    scale_fill_solarized()+
    scale_alpha_manual(values = c(0.7, 0.3))+
    scale_linetype_manual(values = c(1, 3))+
    #scale_fill_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
    #scale_color_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
   #scale_fill_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    #scale_color_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    #scale_linetype_manual(values = c(3, 4, 2, 1))+
    scale_x_continuous(breaks = c(0, 0.5, 1), limits= c(0, 1))+
    scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    xlab("Prevalence of feature in a category") +
    ylab("Scaled probability density") +
  facet_wrap(~PriorShape, scales = 'free')+
  guides(fill = F, color = F)+
  theme(strip.text.y = element_text(angle = 0),
          legend.position = c(0.92, 0.35))

l0.simulations.distributions <- ggplot(all.simulations.refactored, aes(x = value, #y = prob, 
                color = PriorShape, fill = PriorShape, group = Distribution, alpha = Distribution, linetype = Distribution))+
    geom_density(aes(y = ..scaled..),
                 size = 0.6,  adjust = 1.5)+
    theme_few() +
    scale_color_solarized()+
    scale_fill_solarized()+
    scale_alpha_manual(values = c(0.7, 0.3))+
    scale_linetype_manual(values = c(1, 3))+
    scale_x_continuous(breaks = c(0, 0.5, 1), limits= c(0, 1))+
    scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    xlab("Prevalence of feature in a category") +
    ylab("Scaled probability density") +
  facet_wrap(~PriorShape, scales = 'free')+
  guides(fill = F, color = F)+
  theme(strip.text.y = element_text(angle = 0),
          legend.position = c(0.92, 0.35))

```

```{r simulationsOld, eval = F, fig.cap = "Model behavior. Prior distributions over feature-probabilities (Left; grey) and listener model posterior distributions upon hearing an utterance with an uncertain threshold (Middle; red) and a fixed threshold (Right; blue), for five different schematic prior distributions. The uncertain threshold (Middle; red) is hypothesized to correspond to utterances that convey generalizations, while the fixed threshold (Right; blue) is a control model. Shapes of the priors were chosen to intuitivly correspond to properties labeling the distributions.", fig.width = 14, fig.height=6, cache=F}
grid.arrange(l0.simulations.distributions, 
             s1.simulations.scatter, ncol = 3,
             layout_matrix = cbind(c(1,1), c(1,1), c(2, 2)))

```

```{r fim.sims.fixd.thresholds, cache = F}
fig.sims.fixed.thresholds <- ggplot(sims.fixed.thresholds, aes(x = state, fill = PriorShape, color = PriorShape))+
    geom_density(aes(y = ..scaled..), size = 0.6, alpha = 0.3)+
    theme_few() +
    scale_x_continuous(breaks = c(0, 1), limits= c(0, 1))+
    scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    xlab("Prevalence") +
    ylab("") +
    scale_color_solarized()+
    scale_fill_solarized()+
    facet_grid(PriorShape~threshold, scales = 'free')+
    theme(strip.text.y = element_blank()) 
          #axis.title.x = element_text(hjust = 1))
```

```{r figs.sim.posterior, cache = F}
fig.sims.uncertain.threshold <- sims.uncertain.thresholds %>% mutate(PriorShape = factor(PriorShape, 
                             levels = c("barks", "hasSpots", "doesntEatPeople", "laysEggs", "isFemale", "carriesMalaria"),
                            labels = c("bark", "have spots", "don't eat people", "lay eggs", "are female", "carry malaria"))) %>% 
                              ggplot(., aes(x = state, fill = PriorShape, color = PriorShape))+
    geom_density(aes(y = ..scaled..), size = 0.6, alpha = 0.7)+
    theme_few() +
    scale_x_continuous(breaks = c(0, 1), limits= c(0, 1))+
    scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    xlab("Prevalence") +
    ylab("") +
    scale_color_solarized()+
    scale_fill_solarized()+
    facet_grid(.~PriorShape, scales = 'free')+
    theme(strip.text.y = element_blank())#element_text(angle = 0, hjust = 0))
```

```{r fig.sims.endorsement}
s1.simulations.scatter <- ggplot(s1.simulations.relabeled, aes( x = featureProb, y = endorsement, fill = example,
                                                                label = example))+
  geom_col(color = 'black')+
   #geom_text(vjust="inward",hjust="inward")+
  geom_label(data = s1.simulations.relabeled,
                   aes( label = example),
             vjust="inward",hjust="inward",
             color = 'white')+
  # geom_label_repel(data = s1.simulations.relabeled,
  #                  aes( label = example),
  #                      color = 'white',
  #   box.padding = unit(0.0, "lines"),
  #   #point.padding = unit(1e-06, "lines"),
  #   segment.color = 'grey37')+
  scale_fill_solarized()+
  scale_x_continuous(limits = c(0, 1), 
                     breaks = c(0, 1))+
  scale_y_continuous(limits = c(-0.01, 1.01), 
                     breaks = c(0, 1),
                     labels = c("False", "True"))+
  #coord_fixed(ratio = 1)+
  coord_flip()+
  facet_grid(example~.)+
  guides(fill = F)+
  ylab("")+xlab("Referent prevalence")+
  theme(#strip.text.y = element_text(angle = 0, hjust = 0)
        strip.text.y = element_blank()
        #axis.title.x = element_text(hjust = 0)
        #axis.title.y = element_text(vjust = 0)
        )
  #ylab("Endorsement model prediction")+
  #xlab("Hypothesized prevalence")
```

```{r simulationsOld2, fig.cap = "Computational model behavior. A: Prior distributions over the prevalence of a feature in a category for 6 example features B: Different fixed thresholds (columns) give rise to different partial interpretations. C: The interpreter model averages over all thresholds, producing posterior distributions that strongly depend upon the priors. D: Endorsement model predictions, assuming a particular referent-prevalence. Shapes of the priors were chosen to intuitivly correspond to properties labeling the distributions.", fig.width = 12, fig.asp = 0.6, out.width = "95%", cache=F, eval=F}

# p1 <- fig.sims.priors +
#   theme(plot.margin = unit(c(6,0,6,0), "pt")) #+ ggtitle("prevalence priors")
# p2 <- fig.sims.fixed.thresholds +
#   theme(plot.margin = unit(c(6,0,6,0), "pt")) + ylab("")
# p3 <- fig.sims.uncertain.threshold+
#   theme(plot.margin = unit(c(6,6,6,0), "pt"))+ 
#     theme(strip.text.y = element_blank())#+ ggtitle("Interpreter model posterior")
# p4 <- s1.simulations.scatter + theme(plot.margin = unit(c(6,0,6,6), "pt"))
# 
# sim.prevalence.prior.title <- ggdraw() + draw_label("Prevalence priors", fontface='bold')
# sim.prevalence.partial.title <- ggdraw() + draw_label("Partial interpretations given fixed-thresholds", fontface='bold')
# 
# sim.prevalence.posterior.title <- ggdraw() + draw_label("Prevalence posterior (interpretation)", fontface='bold')
# 
# sim.endorsement.title <- ggdraw() + draw_label("Endorsement prediction", fontface='bold', hjust = 0.47)

# prow <- plot_grid( 
#   plot_grid(sim.prevalence.prior.title, 
#             p1 + theme(legend.position="none"), ncol=1, rel_heights=c(0.1, 1)),
#   plot_grid(sim.prevalence.partial.title, 
#             p2 + theme(legend.position="none"), ncol=1, rel_heights=c(0.1, 1)),
#   plot_grid(sim.prevalence.posterior.title, 
#             p3 + theme(legend.position="none"), ncol=1, rel_heights=c(0.1, 1)),
#   plot_grid(sim.endorsement.title, 
#             p4, ncol=1, rel_heights=c(0.1, 1)),
#            align = 'vh',
#            labels = c("A", "B", "C", "D"),
#            hjust = -1,
#            nrow = 1,
#            #rel_widths = c(1.2, 3, 2)
#            rel_widths = c(1.2, 2.9, 1.4, 1.4)
#            )

prow <- plot_grid( 
  plot_grid(sim.prevalence.prior.title, 
            p1 + theme(legend.position="none"), ncol=1, rel_heights=c(0.1, 1)),
  plot_grid(sim.prevalence.posterior.title, 
            p3 + theme(legend.position="none"), ncol=1, rel_heights=c(0.1, 1)),
  # plot_grid(sim.endorsement.title, 
  #           p4, ncol=1, rel_heights=c(0.1, 1)),
           align = 'vh',
           labels = c("A", "B"),
           hjust = -1,
           nrow = 2,
           #rel_widths = c(1.2, 3, 2)
           #rel_widths = c(1.2, 2.9, 1.4, 1.4),
            rel_heights = c(1,1)
           )

# prow <- plot_grid(
#   add_sub(p1 + theme(legend.position="none"), "Prevalence priors"),
#   add_sub(p2 + theme(legend.position="none"), "Partial interpretations given fixed-thresholds"),
#   add_sub(p3 + theme(legend.position="none"), "Full interpretations"),
#   add_sub(p4, "Endorsement"),
#            align = 'vh',
#            labels = c("A", "B", "C", "D"),
#            hjust = -1,
#            nrow = 1,
#            #rel_widths = c(1.2, 3, 2)
#            rel_widths = c(1.2, 2.9, 1.15, 1.7)
#            )

s1.simulations.relabeled
ggdraw(add_sub(fig.sims.priors, 
          expression(
            paste(S("Sharks don't eat people | x = 0.8" ), " = ", 0.4)), size=12, hjust =0))

#prow
```