### 2018-10-15 paper/comgen-paper.Rmd
finish polishing alt theories; expand interim summary to include overview of experiments

---
title             : "The Language of Generalization"
shorttitle        : "The language of generalization"

author: 
  - name          : "Michael Henry Tessler"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "450 Serra Mall, Bldg. 420, Rm. 316, Stanford, CA 94305"
    email         : "mhtessler@stanford.edu"
  - name          : "Noah D. Goodman"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Stanford University"

header-includes:
  - \usepackage{tabularx}
  - \usepackage{multicol}
  - \usepackage{wrapfig}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \usepackage{amsmath}
  - \usepackage{graphicx}
  - \usepackage{subcaption}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{xcolor}

author_note: >
  This paper is currently under review. Comments or suggestions are appreciated and should be directed to MHT.

abstract: > 
    Language provides simple ways of communicating generalizable knowledge to each other (e.g., "Birds fly", "John hikes", "Fire makes smoke"). Though found in every language and emerging early in development, the language of generalization is philosphically puzzling and has resisted precise formalization. Here, we propose the first formal account of generalizations conveyed with language that makes quantitative predictions about human understanding. We test our model in three diverse domains: generalizations about categories (generic language), events (habitual language), and causes (causal language). The model explains the gradience in human endorsement through the interplay between a simple truth-conditional semantic theory and diverse beliefs about properties, formalized in a probabilistic model of language understanding. This work opens the door to understanding precisely how abstract knowledge is learned from language.

keywords          : "genericity, generalization, generics, pragmatics, semantics, Bayesian modeling"
wordcount         : "25000"

bibliography      : ["generics.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
toc               : no

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
---

\newcommand{\denote}[1]{\mbox{ $[\![ #1 ]\!]$}}
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\definecolor{Red}{RGB}{255,0,0}
\definecolor{Green}{RGB}{10,200,100}
\definecolor{Blue}{RGB}{10,100,200}

\newcommand{\mht}[1]{{\textcolor{Blue}{[mht: #1]}}}
\newcommand{\ndg}[1]{{\textcolor{Green}{[ndg: #1]}}}
\newcommand{\red}[1]{{\textcolor{Red}{#1}}}

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = F, fig.path='figs/',
                      echo=FALSE, warning=FALSE, cache=T, message=FALSE, sanitize = T)
```

```{r libraries, cache=F}
library(papaja)
library(formatR)
library(rwebppl)
library(xtable)
library(tidyverse)
library(forcats)
library(langcog)
library(coda)
library(ggthemes)
library(ggrepel)
library(jsonlite)
library(gridExtra)
library(lme4)
library(knitr)
library(kableExtra)
library(cowplot)
library(magick)
library(viridis)
theme_set(theme_few())
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
logmeanexp <- function(x){
  x.num <- as.numeric(x)
  xstar = max(x.num)
  return(xstar + log(mean(exp(x.num - xstar))))
}

compute_r2 <- function(df,v1, v2, sigfigs = 3){
  return(format(cor(df[[v1]], df[[v2]])^2, digits = sigfigs))
}

compute_mse <- function(df, v1, v2, sigfigs = 3){
  return(format(mean( (df[[v1]]-df[[v2]])^2), digits = sigfigs))
}

project.path <- "../"
options("scipen"=10) 
```

<!-- 
possible titles:
- Communicating generalizations
- The language of generalization
- The language of generalization: Probability, vagueness, and interaction
-->

<!-- Generalizations are central to human understanding and language provides simple ways to convey them (e.g., "Birds fly"). -->
<!-- The language of generalization is ubiquitous in everyday discourse and child-directed speech.  -->
<!-- The major issue in formalizing generalization in language is determining which statements are true or which are false. -->
<!-- Using an information-theoretic probabilistic model, we explore the hypothesis that the meaning of these linguistic expressions is *simple but underspecified*, and that general communicative principles can be used to establish a more precise meaning in context.  -->
<!-- To test this theory, we examine endorsements of generalizations about three different domains: categories (*generic language* e.g., "Birds fly"), events (*habitual language* e.g., "John runs"), and causes (*causal language* e.g., "Staring at the sun makes you go blind").  -->
<!-- Across these diverse domains, we find that our model explains the wide variance in human endorsements of language conveying generalization, while simpler models fall short. -->
<!-- Ours is the first formal theory that makes precise, quantitative predictions about human understanding of the language of generalization. -->
<!-- The results demonstrate that the context-sensitivity of such language emerges from the interaction of an underspecified meaning with diverse beliefs about properties.  -->

<!-- Knowledge that extends beyond the present context is crucial to thrive in our open-ended, dynamic world. Yet such knowledge can be difficult to extract from the environment. Fortunately, we are not limited to acquiring generalizations on our own; language allows us to communicate generalizations to each other (e.g., “Asparagus berries are poisonous”, “John hikes”, “Drinking milk makes your bones strong"). In this talk, I’ll argue that three ingredients are necessary to understand generalizations: Probability, vagueness, and context. I formalize these ingredients in an information-theoretic probabilistic model that makes quantitative predictions about human understanding of generalizations. Across diverse domains, I find that the model explains the gradience in multiple dependent measures, while simpler models fall short. This is the first formal theory that makes accurate and precise predictions about human understanding of generalizations in language and is the first step towards building models that can learn abstract knowledge from language. -->

# Introduction

<!-- Learning that an object tends to have a property, an entity tends to exhibit a behavior, or a cause tends to produce an effect can be crucial to thrive in our open-ended world. -->
Knowledge that extends beyond the present context is crucial to thrive in our open-ended, dynamic world. 
Yet such knowledge can be difficult to extract from the environment: The relevant observations may be costly (e.g., learning that a plant is poisonous) or rare (e.g., understanding that lightning strikes tall objects).
Fortunately, we are not limited to acquiring generalizations on our own; language allow us to communicate generalizations to each other.
By sharing generalizable knowledge, we flourish collectively without individually needing to taste potentially-poisonous plants or personally witness lightning strikes.
Being able to flexibly communicate generalizations from one generation to the next supports the faithful transmission of knowledge necessary for culture to cumulatively evolve [@Tomasello1999; @Henrich2015].

<!-- Generalizations can be acquired from the world by drawing inductive inferences from observations [@Baldwin1993; @Xu2008; @Dewar2010]. \ndg{cite older stuff. eg hume} -->
<!-- Cooperative teachers can facilitate inductive learning by supplying helpful observations to the learner [@Csibra2009; @Tomasello1999; @Butler2012]. -->
<!-- At around a child's first birthday, however, the learner gains access to a new, human-unique source of data: language. -->
<!-- Minimally, language fosters generalizations by bringing attention to objects being of the same kind through the use of labeling [e.g., "This is a dog"; @Gelman1986; @Gopnik2000; @Graham2004; @Markman1989; @Schulz2008]. -->
<!-- Objects have many properties though; when speakers attach properties to kind-labels (e.g., "Dogs bark"), listeners hear a generalization. -->
<!-- \ndg{this par doesn't really do much} -->

<!-- Certain linguistic expressions encode generalizations. -->
<!-- Language provides multiple, distinct ways of encoding generalizations. -->
The *language of generalization* covers a diverse swath of natural language expressions.
Generic language conveys generalizations about categories [e.g., "Dogs have four legs"; @Carlson1977; @Cohen1999; @Leslie2007; @Nickel2008] and is the most well-studied case generalizations in language.^[
  Some writers refer to the all of the *language of generalization* as "generic language" or generics.
  In both the empirical and theoretical literatures, however, analysis and experiments often focus only on generalizations about categories. 
  We use the more narrow-scoped terms (e.g., generics, habituals, and causals) to highlight the diversity of semantic types being predicated.
]
In contrast to statements about concrete individuals (e.g., "Rufus has four legs"), generic statements refer to inherently unobservable categories (e.g., the category of \textsc{dog}) and convey information that extends beyond the present context, a fact which children as young as 2 appreciate [@Cimpian2008].
Simple events (e.g., "John ran yesterday") can be generalized into habitual sentences (e.g., "John runs"), and even events of complex inferential types such as actual causal events (e.g., "The fire caused the smoke") can be described in generalization (e.g., "Fire causes smoke").

Understanding the language of generalization is a project with far-reaching implications.
The language of generalization is ubiquitous in everyday conversation, is found in every language [@Behrens2005; @Carlson1995], and conveys rich meanings, impacting motivation [@Cimpian2007], transmitting stereotyped beliefs about social categories [@Rhodes2012], and making meaning from experience [@Orvell2017].
It is highly-prevalent in child-directed speech [@Gelman2008] and its ability to refer to abstractions beyond the present context suggests its centrality to the growth of conceptual knowledge [@Gelman2004].

<!-- Generics are just one case in the language of generalization, however; events can be described in generalization (*habitual language*; e.g., "Mary swims after work.") as well as causal relationships (*causal language*; e.g., "Staring at the sun makes you go blind."). -->

<!-- Generics are ubiquitous in everyday conversation as well as in child-directed speech [@Gelman2008].  -->
<!-- Additionally, generics are the primary way by which speakers discuss social categories, and thus are key to propagating stereotypes [@Gelman2004; @Rhodes2012; @Leslie2015] and impacting motivation [@Cimpian2010]. -->
<!-- Thus, an understanding of the language of generalization that extends to a wide variety of types (e.g., categories, events, causes)  -->

<!-- Thus, an understanding of generalization as a singular phenomenon in language  -->
<!-- promises to provide an understanding of a wide variety of linguistic expressions with rich meanings. -->

<!-- Generic statements ascribing properties to categories are just one kind of generalization found in language.  -->
<!-- Individual entities can also be talked about in generalization, using so-called  -->
<!-- <!-- Recent evidence suggests talking about the other in generalization (using so-called *generic You*) can help make meaning from difficult situations   [e.g., "You never know what will happen on a blind date."; @Orvell2017]. -->
<!-- Communicating about causality can also take the form of communicating a generalization about a causal event  -->
<!-- What these statements have in common is that they convey generalizations, or exhibit *genericity* [@Carlson1977; @Carlson1995]. -->

Despite its ability to convey abstract knowledge, its ubiquity in discourse, and its relative morphosyntactic simplicity, the language of generalization displays subtle context-sensitivities that make it difficult to formalize.
"Robins lay eggs" sounds true and "Robins are female" does not. 
Yet in each case, only 50% of the category has the property (i.e., only the females lay eggs). 
"Mosquitos carry malaria" sounds true despite malaria being present in only a tiny fraction of mosquitos. 
Even more perplexing: "Supreme Court Justices have even social security numbers" is thought to be intuitively a false generalization even if it were the case that on the current bench, nine out of nine justices had even social securtity numbers [i.e., when the sentence "All Supreme Court Justices have even social security numbers" is true; @Cohen1999].
Similar context-sensitivity can be observed with habituals: "Mary climbs mountains" could imply a few mountain climbs a year for Mary, but "John runs" would be infelicitious if John went for a run three times last year.

<!-- When we hear "Dogs bark", it sounds like a universal statement, as in "All dogs bark". -->
<!-- But the property does not apply to all dogs: Basenji dogs are barkless, for example. -->

<!-- The generic could signal something weaker like "Most dogs bark". -->

<!-- However, *most* mosquitos do not carry malaria (in fact, only a tiny fraction do), but "Mosquitos carry malaria" is intuitively true. -->
<!-- *most mosquitos* do not carry malaria, while the statement "Mosquitos carry malaria" is intuitively true. -->
<!-- The only *quantifier* that would permit such a statement would be *some* (e.g., "Some mosquitos carry malaria"). -->
<!-- But weakening the generalization to mean *some* is too permissive: Some robins are female (in fact, half of them are), but endorsing "Robins are female" seems odd. -->
<!-- Even more perplexing: Those very same female robins lay eggs, and "Robins lay eggs" is a perfectly fine thing to say. -->

These observations have led some to conclude that the literal meaning of a generic statement (and by analogy, other generalizations in language) involves more than just the number of members of the category who have the property, otherwise known as the *prevalence* of the feature in the category.
<!-- (or the probability of a member of the category having the property e.g.,  $P(x \text{ is friendly} \mid x \text{ is a dog})$) is not sufficient to make generalizations ("Dogs are friendly") true or false. -->
Instead, theorists have argued that generics are not treatable by the standard tools of truth-functional semantics [@Montague1973], but rather should be thought of semantically as a direct, linguistic manifestation of abstract relations between kinds and properties [@Leslie2008; @Prasada2000; @Prasada2012]. 
<!-- Conceptual accounts of generics emphasize the structure of *generic knowledge* [@Prasada2000], and view generic utterances as the way of expressing special, mental relationships between kinds and properties [@Leslie2008; @Prasada2012].  -->
<!-- Indeed, generics can be true for a range of prevalence or frequency levels [@Khelmani2012; @Prasada2013]. -->
For example, the statements "Bishops move diagonally" or "The Speaker of the House succeeds the Vice President" are true not because of a tendency on behalf of instances of a category to actually uphold the property, but rather the existence of a conceptual relationship (e.g.,  what it means, in the game of chess, to be a bishop).
True generics can be supported by different underlying types of category--property relations [e.g., principled vs. statistical connections; @Prasada2006], and thus support qualitatively different inferences [e.g., "Being striped is one aspect of being a tiger" is generally endorsed, while "Carrying malaria is one aspect of being a mosquito" is not, even while "Tigers are striped" and "Mosquitos carry malaria" are both intuitively true; @Prasada2013].
This conceptual view of generics has been influential in psychology, because it predicts qualitative differences between different kinds of generics. 
<!-- Still, the theory is not sufficiently precise to make any quantitative predictions. -->

Insofar as there is a single class of linguistic expressions that convey generalization, however, there should be something common to them all: a literal meaning that unifies generic, habitual, and causal language.
In this paper, we propose such a semantic core based on prevalence or probability and formalized using the tools of truth-functional semantics [@Montague1973; @Cohen1999].
A semantics based on probability will not be enough to subtle sensitivities to context that the language of generalization exhibit.
We propose that the meaning of generalizations is underspecified or *vague*, and that listeners derive a more precise interpretation in context using world knowledge.

The fact that generics are vague does not preclude them from being treated with formal models.
We draw upon the tools of Bayesian models of cognition [@Tenenbaum2011] to formalize the vagueness and context-dependence of generic language [@Frank2012; @GoodmanLassiter].
The Bayesian model separates the semantics of an utterance conveying a generalization from the world knowledge a listener would use to interpret the utterance, a key theoretical advancement beyond previous accounts.
<!-- By its instantiation in a formal model, this hypothesis has meaningful conditions under which it can be tested. -->
This formal model is the first of its kind to make quantitative predictions about human understanding of generalization in language.
<!-- With these ideas, we create an endorsement model, that describes what makes generalizations in language true or false. -->
<!-- Our model is the first to make quantitative predictions about human endorsements of generalizations in language. -->

The paper is organized as follows. 
In the next section, we describe our computational framework for interpreting generalizations and our precise model for endorsing such statements (i.e., a model of truth judgments).
To illustrate how the model works, we then work through a number of standard examples from the linguistics literature that any model of generics should be able to accomodate.
The third section discusses the relationship of our model to previous accounts of generics from the linguistics and philosophical literatures. 
These introductory sections are followed by three empirical case studies: generalizations about categories (*generic language*), events (*habitual language*), and causes (*causal language*). 
In the case study of generics, we use measure relevant background knowledge and prevalence to predict endorsements of familiar generic statements (e.g., "Robins lay eggs").
In the study of habituals, we measure background knowledge but manipulate the prevalence or frequency of the event to predict endorsements of habitual statements about novel agents (e.g., "John runs" given that he's run a certain number of times in the past).
In this case study, we also further examine the nature of the relevant probabilities for endorsing generalizations, asking whether it matters how often John has run in the past (past frequency) or how often a speaker expects him to run in the future (predictive probability).
Finally, in our last case study, we manipulate background knowledge to show its causal influence on endorsing generalizations about novel causal events (e.g., "Herb X makes wugs sleepy"). 
We compare our model to two previously articulated quantitative models of generics as well as a lesioned-version of our model.
In all cases, we find a very strong agreement of our model's predictions to human elicited endorsements, where the simpler models fall short.
We conclude our paper by clarifying some of the theoretical claims of the model and discuss open-questions for this approach.

<!-- formalizes the hypothesis that the core meaning of a generalization in language is a vague probability and articulates it in a model of language understanding in which relevant background knowledge is formally described. -->
<!-- Having defined a model of interpretation, we then describe an information-theoretic, communicative linking function from interpretation to judgments about the truth or acceptibility of a sentence. -->
<!-- We sketch out a number of worked examples to qualitatively demonstrate the model's behavior and the describe the relation of our model to previous linguistic and philsophical approaches to generics. -->

<!-- The model as articulated predicts fine-grained, quantitative effects of background knowledge, which we both measure and manipulate across our  -->

<!-- This work opens the door to understanding generalizations in richer, pragmatic language contexts and building models that learn abstract knowledge from language. -->

<!-- relevant background knowledge used to understand a generalization in context, and use these components to develop a computational model of endorsement. -->
<!-- The endorsement model's predictions depend upon relevant background knowledge and the prevalence communicated, which we demonstrate via simulation. -->
<!-- Across these three case studies, we both measure and manipulate background knowledge and the prevalence communicated, measuring their effects on endorsement of generalization sentences.  -->

<!-- To do so, we look to probability, the universal currency of belief and a useful representation for human generalization from observations [@Shepard1987; @Tenenbaum2001]. -->
<!-- Given that certain probabilities are the result of inductive generalization, it would make sense that these are the object of communication for the language of generalization. -->
<!-- \ndg{cuttoff / threshold not mentioned before this, need to introduce...} -->
<!-- However, it is not possible to develop a single cut-off for when the generalization should hold, due to examples like "Birds lay eggs" vs. "Birds are female". -->
<!-- Rather than throw-out the quantitative formalization of the threshold semantics, we posit that the threhsold is *underspecified* in the literal meaning of a generalization, and that listeners' background knowledge and the communicative context can be used to derive more precise interpretations in the moment.  -->

<!-- To formalize a model of endorsement (i.e., what makes the statements true or false), we consider the communicative function of a generalization by explicitly modeling the background knowledge a language interpreter would bring to bear when understanding generalizations. -->

<!-- It's been suggested that the mind has cognitive primitives that directly dictate what generalizations are true or false. -->
<!-- One such primitive is the ability to priviledge *characteristic features*: For an animal, their mode of reproduction is a characteristic feature.  -->
<!-- Thus, whatever their mode of reproduction may be will make a true generic statement (e.g., "Robins lay eggs"), regardless of the actual prevalence of the feature. -->
<!-- *Horrifying* or *striking* information is also priviledged and creates true generics, which is why "Mosquitos carry malaria" is true -->

<!-- Specifically, we formalize an account where the core meaning of a generalization in language is a *simple but underspecfied* function of probability and where context helps resolve the underspecification. -->
<!-- In its simplest form, context takes the shape of interlocutors' shared beliefs about the relevant properties or categories under discussion [cf., @Grice1975; @Clark1996; @Levinson2000]. -->
<!-- We then ask what a rational speaker---who can only either endorse the generalization or stay silent---would say in different contexts; this is our formal model of endorsement. -->
<!-- We show how the decision to endorse depends in systematic ways on the shared prior knowledge of the interlocutors. \mht{"interlocutors" may be confusing here...} -->
<!-- We test this theory in three distinct case studies---generalizations about categories, events, and causes---by both measuring and manipulating prior knowledge, showing that it is strongly and causally related to the endorsement of generalizations in the ways predicted by our model. -->

<!-- (cumulative $r^2(N) = 0.XX$). -->
<!-- In the final section of the paper, we discuss the generality of our approach and pave a path to unifying the extant psychological, linguistic, and philosophical literatures on generalization in language. -->
<!-- We also show how our endorsement model is a special case of a speaker model in a rational framework of communication.  -->

# Computational Framework

Generalizations are used to make predictions about events or properties of instances that an agent has yet to experience [@HumeTHN].
People readily predict that the next dog they encounter will have four legs, drinking another cup of coffee will cause jitters, and a new day will find the people that we know doing what they habitually do. 
In each case, we assign a specific exemplar $x$ to a category $k$, and make a prediction that it will have feature $f$.
This prediction can be described by a conditional probability: $P(x \in f \mid x \in k)$, the probability that $x$ will have $f$ given that it is in $k$ (formally, an instance of $k$ will be in the set of things that have $f$), which we will refer to as the *prevalence* and for convenience write as $p = P(f \mid k)$.
Our *prevalence* $p$ is a *prevalence in the mind*, a latent belief that a future instance of a category would have a particular property; others might call this projectibility [@Goodman1955].
The targets of our predictions can vary widely: They may be objects (a dog), events (an instance of coffee-drinking), or more ad-hoc types (a person on a particular day).
The properties also may vary (e.g., having fur, causing jitters, going to the gym).
Yet the mathematical description of the inductive belief is always given by a probability $P(f \mid k)$.

<!-- a useful representation  -->
<!-- How do we come to know things?  -->
<!-- How does one acquire inductive beliefs? -->
Probability is a useful representation for human generalization from observations [@Shepard1987].
If you observe several *wugs* (a novel category) that have two legs, you might infer that all wugs have two legs.
But not all properties have such strong projectibility:
Seeing a wug with broken wings tells you comparatively less about other wugs having broken wings [@Nisbett1983].
Abstract, potentially domain-specific, beliefs about the projectibility of different properties can be represented by a probability distribution over the prevalence $P(p)$ [@Kemp2007].
By assuming some generative process that could produce one's observations $o$ (i.e., a likelihood function $P(o \mid p)$), Bayes' Theorem provides the mathematically correct way to update one's prior beliefs from observations $o$: $P(o \mid d) \propto P(o \mid p) \cdot P(p)$  [@Tenenbaum2001].

<!-- To do so, we look to probability, the universal currency of belief and  -->
<!-- Given that certain probabilities are the result of inductive generalization, it would make sense that these are the object of communication for the language of generalization. -->

<!-- But we do not always have access to  -->
Observational data is not always available, however.
Instead, we must listen to others to learn about properties that are costly to observe (e.g., *staring at the sun makes you go blind*), events that are statistically unlikely (e.g., *lightning strikes tall objects*), or any aspect of the world that we have yet to experience.
Fortunately, language provides simple ways of communicating generalizations.

<!-- Here, $P(d \mid h)$ is the observer's generative model of the observed data.  -->
<!-- It is conditional on $h$ because the probability of various direct observations depends upon the prevalence that the observer assumes.  -->
<!-- Generalization from observational evidence is limited, however, in its capacity to describe how the mind builds complex theories of the world or how culture accumulates across generations. -->
<!-- How do we acquire these generalizations? -->
<!-- For abstract, generalizable knowledge to remain in a culture, it must be faithfully transmitted between individuals and across generations [@Tomasello1999], and language provides simple ways to communicate generalizations. -->
<!-- The human capacity to generalize from observations, deriving $h$ from example $x$s, has been studied extensively in cognitive and developmental psychology [@Nisbett1983; @Baldwin1993; @Heit2000]. -->
<!-- From a very young age, children draw strong inferences about nonobvious object properties from just a few examples [@Baldwin1993; @Gopnik2000;  @XuTenenbaum2007]. -->
<!-- Such inductive generalization can be described by the Bayesian probability calculus: An observer's initial beliefs or state of knowledge---called a prior belief distribution---is updated through experience (i.e., after making observations) and becomes a posterior belief distribution (what they belief *a posteriori*).  -->
<!-- This posterior distribution becomes the observer's prior for the next observation she makes: it thus informs her predictions about unobserved instances. -->

<!-- As a concrete example, consider a learner investigating the feature \textsc{barks} ($f$) for the category \textsc{dog} ($k$). -->
<!-- $h$ describes the learner's subjective probability that the next dog she encounters will bark^[At the moment, we assume the feature \textsc{barks} is an all-or-none Boolean property. Later, we will see how the presence of this feature itself can be thought of as a generalization over events of barking.], or more abstractly, her belief about the barking capacity of dogs. -->
<!-- That is, $h = P(x \text{ barks} \mid x \text{ is a dog})$. -->
<!-- The learner does not know $h$, but has some idea about what $h$ could be; this is represented by $P(h)$, the learner's prior beliefs about how likely it is that $x$ will bark given that $x$ is a dog. -->

<!-- The observer goes around the world, encountering dogs and taking note of how many bark. -->
<!-- This could be represented as a number of positive examples $d$ out of a number of total examples $n$.  -->
<!-- With the weak assumption that the dogs she has observed were all independent and identically-distributed (*i.i.d.*) random samples from the category of dogs, the likelihood of her observations are given by the Binomial distribution.  -->
<!-- That is, $d \sim  \text{Binomial}(n, h)$, where $d$ is the number of dogs she has seen that bark, $n$ is the total number of dogs she has seen, and $h$ is the prevalence of barking among dogs. -->
<!-- This is called a likelihood function: It determines how well a given value of $h$ predicts the observed $d$.  -->
<!-- Given a formal description of the learner's state of knowledge before any observations $P(h)$ and the learner's assumptions about how the data she observed were generated (e.g., the *i.i.d.* assumption), what a learner should believe after having seen the data is proportional to the product of these two terms, as given by Bayes' Rule: -->

<!-- $P(h \mid d)$ is then the strength of the learner's belief in the next dog she encounters barking. -->
<!-- This simple idea of integrating observational evidence with prior knowledge is surprisingly powerful in explaining a multitude of diverse phenomena, from perception to word learning to inductive reasoning [e.g., @Tenenbaum2006; @Tenenbaum2011]. -->
<!-- \ndg{what is this par doing for us? delete?} -->
<!-- Generalization from examples shows complex sensitivites to the kind of observations, and even how those observations came to be observed in the first place. -->
<!-- For instance, observing that an indigenous person living on a remote island *is brown and obese* provides considerably more information about the skin color of other islanders than about their weight [@Nisbett1983]. -->
<!-- This subtlety can be modeled by incorporating heiarchical structure into the prior belief distribution of a Bayesian model, allowing the model to represent and reason flexibly about different kinds of properties [@Kemp2008]. -->
<!-- Observations that are generated by an intentional agent can lead to similarly subtle inferences, like deconfounding otherwise ambiguous causal evidence [@Goodman2009]. -->
<!-- Yet generalizations from observational evidence is limited in its capacity to describe how the mind builds complex theories of the world or how culture accumulates across generations. -->
<!-- Relying upon observations would be particularly problematic, for example, to learn about properties that are costly to observe (e.g., *staring at the sun makes you go blind*) or events that are statistically rare (e.g., *lightning strikes tall objects*). -->
<!-- How do we acquire these generalizations? -->
<!-- For abstract, generalizable knowledge to remain in a culture, it must be faithfully transmitted between individuals and across generations [@Tomasello1999], and language provides simple ways to communicate generalizations. -->

## Communicating generalizations

The language of generalization is easy to spot when a property, that could apply to an individual, is predicated of a category (e.g., "Dogs have four legs": *has four legs* could apply to an individual as in "This dog has four legs").
In the semantics literature on generics, bare plural sentences of this kind are sometimes described as *characterizing sentences* in contrast with *kind-denoting* sentences where the property can only meaningfully apply to the category as a whole (e.g., "Dinosaurs are extinct"; *extinct* cannot apply to an individual dinosaur).
<!-- Dinosaurs are exctinct ~ John is a smoker? -->
Generalization can also manifest when describing instances of an individual [*habitual language*; e.g., "John smokes"; @Carlson1977; @Carlson2005]; in this case, the particular instance being generalized is an instance of an individual (e.g., John at a particular moment in time), which also permits predication (e.g., "John smoked yesterday after dinner").
Verbs like *causes* or *makes* also seem to convey generalization, in this case, about an instance of an actual causal event (*causal language*; e.g., "Fire causes smoke"). 
Psychologists, linguists, and philosophers have long studied the language of generalization, as it appears very simple (e.g., syntactically) yet its meaning is difficult to formalize.^[
  We further distinguish the problem of formalizing a *meaning* for the language of generalization as opposed to the problem of *identifying* generalizations. 
  Syntax alone is neither necessary nor sufficient for a listener to know that the sentence conveys a generalization  (e.g., indefinite singulars can encode generalizations: "A dog has four legs" and bare plurals may not: "Dogs are on my front lawn").
  Our analysis thus begins once a sentence has been disambiguated as conveying a generalization. 
]
We first describe a model for generics, before noting how the basic ideas generalize to habituals and causals.

If generalization from observations can be described by a probability $p$, it is natural to posit that same construct will be at the heart of a semantic theory of the language of generalization (Ingredient 1: Probability).
In semantics, belief updating generally passes through Boolean *truth values* [@Montague1973].
The simplest way to derive a Boolean from a scalar quantity like probability is via a *threshold semantics*: The utterance is true if the relevant scalar value is above a threshold.
For example, the literal meaning of the sentence "Some dogs have four legs" is that there is a non-zero chance that a given dog will have four legs: $\denote{some}(p) := p > 0$.
"Most dogs have four legs" can also be described as a threshold on prevalence (e.g., the chance that a dog will have four legs is greater than 50%): $\denote{most}(p) := p > 0.5$.^[
  These definitions concern the standard semantic truth conditions of quantifiers, not their pragmatic interpretations (e.g., that "some" often implies "not all").
]
Thus, the simplest semantics for a generalization would also be a threshold on the prevalence: $\denote{gen}(p, \theta) := p >\theta$.

The extreme flexibility of generalizations (e.g., "Mosquitos carry malaria"; "Birds lay eggs" vs. "Birds are female") suggests that no fixed threshold would suffice. 
Rather than throw out the threshold-semantics, we posit that the threshold is *underspecified* in the literal meaning and is contextually-determined in a way analogous to how gradable adjectives like *tall* have contextually-determined thresholds [e.g., what counts as tall for a four-year-old is different than what counts as tall for an adult; @Kennedy2007; @Lassiter2013].
We formalize this underspecification of meaning (Ingredient 2: Vagueness) by putting a probability distribution over $\theta$: $P(\theta)$ [@Lassiter2013; @Lassiter2015; cf., @Qing2014].

Finally, the meanings of linguistic expressions manifest in their capacity to convey information from speaker to hearer [@Grice1975; @Clark1996; @Levinson1995].
Thus, the final ingredient to our model is *context* (Ingredient 3), which must be minimally formalized as a listener's prior knowledge about the property $P(p)$ and which we describe in the *Worked Examples* section below.

Putting these three ingredients (probability, vagueness, and context) together, we get the following model of interpreting generalizations in language:

\begin{eqnarray}
L(p, \theta \mid u) &\propto& {\delta_{\denote{u}(p, \theta)}  \cdot P(\theta) \cdot P(p)} \label{eq:L0}
\end{eqnarray}

We denote this probabilistic interpretation model by $L$ to indicate that it is modeling a listener updating their beliefs according to the truth-functional meaning of an utterance $u$.
Formally, the truth-functional meaning is represented by the Kronecker delta function  $\delta_{\denote{u}(p, \theta)}$ that returns $1$ when the utterance is true (i.e., when $p > \theta$) and $0$ otherwise.

\begin{eqnarray}
\delta_{\denote{u_{gen}}(p, \theta)} &\propto  & \begin{cases}
1 & \text{if } p > \theta \\
0 & \text{otherwise}
\end{cases}\label{eq:delta}
\end{eqnarray}

Following @Lassiter2015, we formalize the vagueness in the meaning of generics $P(\theta)$ as a uniform distribution over the support of $P(p)$.

With these ingredients, the literal interpretation model (Eq. \ref{eq:L0}) computes a posterior interpretation distribution on prevalence by considering different possible thresholds $\theta$.
Figure\ \@ref(fig:uniformPriorSimulation) shows the model behavior assuming a uniform prevalence prior $P(p)$, perhaps representing a purely abstract property for which an interpreter has no substantive prior knowledge.
If the threshold were to be very high (e.g., $\theta = 0.75$), only the highest prevalence levels would be consistent with the utterance.
As the threshold decreases, more and more prevalence levels become compatible with the threshold (Figure\ \@ref(fig:uniformPriorSimulation)A).
The interpretation model weights the different thresholds by the probability that each is true, resulting in a posterior interpretation distribution on prevalence that favors higher prevalence levels (Figure\ \@ref(fig:uniformPriorSimulation)B).
This interpretation distribution captures the intuition that generalizations imply high prevalence, though with substantial uncertainty about the precise prevalence.

The model in Eq \ref{eq:L0} describes a general mechanism of belief-updating via an uncertain threshold semantics.
The prevalence prior $P(p)$ is a distribution over probabilities $p$, which represents the output of a generative model of particular domain.
For properties of objects, the generative model is naturally a Bernoulli model (e.g., flipping a coin) and thus the support of $P(p)$ is $[0, 1]$.
For events, the implicit generative model is a Poisson model in which events occur randomly in time.
The probability density of an event occurring at a particular time is governed by the rate parameter for the Poisson model. 
Thus, Eq. \ref{eq:L0} becomes a model of habitual interpretation (e.g., "John runs") if we take the underlying semantic scale to be that of the temporal frequency of an event (e.g., how often we expect an event to occur); the support of $P(p)$ is this case would be $[0, \infty]$.

<!-- The extreme flexibility of generalizations (e.g., "Mosquitos carry malaria"; "Birds lay eggs" vs. "Birds are female") suggests that no fixed threshold would suffice.  -->
<!-- Rather than throw out the threshold-semantics, we posit that the threshold is *underspecified* in the literal meaning and is contextually-determined in a way analogous to how gradable adjectives like *tall* have contextually-determined thresholds [e.g., what counts as tall for a four-year-old is different than what counts as tall for an adult; @Kennedy2007; @Lassiter2013]. -->

<!-- So what threshold $\theta$? should be used for the generalizations?  -->
<!-- Unlike quantified language which have fixed-thresholds, we posit that the language of generalizations is *underspecfied* with respect to the actual truth-functional threshold $\theta$. -->

### Endorsement model

In this paper, we are interested in explaining truth judgments (e.g., "Robins lay eggs" is intuitively true; "Robins are female" is not).
We model the endorsement task (e.g., *true* vs. *false*; *agree* vs. *disagree*) as a *speaker decision* about whether or not to produce the generalization to a naive listener; thus, the endorsement model has two alternative utterances: produce the generalization vs. a null alternative [@Franke2014cogsci; @Degen2014].
For simplicity, we take the null alternative to be an informationless "silent" utterance which is always true: 
$\denote{null}(p, \theta) = \text{True}$.^[
  The null alternative can be realized in at least two other ways: the negation of the generalization (e.g., "It is not the case that Robins lay eggs") or the negative generalization (i.e., "Robins do not lay eggs").
  All results reported are similar for these two alternatives, and we use the alternative of the "silent" utterance for simplicity.
]
This kind of speaker model defined in terms of a listener model formalizes the basic aspects of communicative reasoning and is the simplest instantiation of a Rational Speech Act model [@Frank2012; @Goodman2016].

<!-- If the interpreter is a kind of simple listener model, then we can ask whether an agent would bother to produce the generalization to this listener. -->
<!-- The endorsement model then acts as a kind of speaker, reasoning about how the interpreter model (the listener) would understand the utterance. -->

The endorsement model (called $S$ for speaker) decides whether or not to produce the generalization by reasoning about how the listener model ($L$; Eq. \ref{eq:L0}) would interpret it:
\begin{equation}
S(u \mid p) \propto (\int_{\theta} L(p, \theta \mid u)  \diff \theta)^\lambda
\label{eq:S1}
\end{equation}

The goal of the endorsement model is to align the listener's *a priori* beliefs about prevalence (given by the prevalence prior) with the speaker's beliefs about the prevalence of the feature in the referent category $p$, the *referent prevalence* (e.g., that 50% of robins lay eggs). ^[
  A more general version of this model can relax the assumption that the endorser / speaker has access to a specific prevalence $p$ that it wants to communicate.
  Rather, the endorser may have probabilistic beliefs about the prevalence of the property for the referent category $k$, which would be represented by a distribution over $p$: $P(p \mid k)$. 
  In this situation, we would define the endorsement model decision to be with respect to the expected value of the informativity, which integrates over the endorsement model's belief distribution:
  $S(u \mid k) \propto \exp{(\lambda \cdot {\mathbb E}_{p\sim P_{k}} \ln{ \int_{\theta} L(p, \theta \mid u)} \diff \theta )}$.
  For the empirical case studies described below, these two versions of the model make almost identical predictions.
]
The endorsement model makes an approximately rational (with degree of rationality $\lambda$), information-theoretic decision based on its beliefs about which utterance would best achieve the goal of conveying the referent prevalence $p$.
The semantic threshold $\theta$ is necessary for establishing the truth conditions and deriving an interpretation (used in Eq. \ref{eq:L0}) but is otherwise a nuisance parameter, which the endorsement model integrates out (note that there is no $\theta$ in the left-hand side of Eq. \ref{eq:S1}).
Given that the speaker only has two options (i.e., produce the generalization utterance or stay silent), the endorsement decision comes down to whether or not the referent prevalence is more likely under the prevalence prior distribution $P(p)$ (i.e., the listener's posterior upon hearing silence) or the listener generic interpretation distribution $L(p \mid u)$ given by Eq. \ref{eq:L0} (Figure\ \@ref(fig:uniformPriorSimulation)C).
In the next section, we work through different examples from the linguistics literature, describing the prevalence prior and other relevant model components for each.

<!-- Our endorsement model (Eq. \ref{eq:S1}) provides a mapping from the prevalence in the category being referred to (i.e., how many robins lay eggs)---the *referent prevalence* $p$---to an endorsement probability for the corresponding generalization (e.g., "Robins lay eggs"): $S(u \mid p)$. -->
<!-- This formalism builds on a tradition in formal semantics to try to precisely define the conditions under which a sentence is true or false [@Montague1973]. -->
<!-- We go beyond previous semantic models of generics [e.g., @Cohen1999] by embedding our semantic hypothesis (a simple but vague threshold) in a cognitive model in which relevant contextual knowledge can be formalized. -->
<!-- In this simple model, the relevant contextual knowledge is given by a prior distribution over the prevalence $P(p)$. -->

<!-- By doing so, we are able to describe how context influences generic understanding and are able to make precise, quantitative predictions about human judgments. -->

<!-- The prevalence priors represent abstract beliefs about the properties -->
<!-- To reason through those questions, you may rely upon your intuitive theories of reproduction (e.g., only females lay eggs).  -->
<!-- Each property has a prevalence  -->

<!-- For example, many different kinds of animals *do not bark* (more precisely: have close to 0 exemplars who bark); thus, the prevalence prior for the feature *barks* should have substantial probability mass near 0, because one would expect a randomly-sampled category to have 0 or close to 0 exemplars that bark. -->
<!-- Furthermore, among the categories that do have individuals that bark (e.g., dogs, seals, ...), the probability is quite high that an exemplar will bark; thus, the corresponding prevalence prior would have some probability mass at relatively high prevalences, potentially with some uncertainty (Figure\ \@ref(fig:simulations); column 1). -->
<!-- Abstract beliefs between different kinds of categories could manifest in these prevalence priors in the form of multi-modal distributions; for example, the prevalence prior for *lays eggs* should similarly have substantial probability mass near 0 (since many animal categories do not have egg-layers) but should have a secondary component peaked around 0.5, because among the animals with egg-layers, only the female members of the category have the property (Figure\ \@ref(fig:simulations); column 4). -->

<!-- The uncertain semantic threshold $\theta$ is resolved by the literal interpretation model $Lit$ (Eq. \ref{eq:L0}) the model's background knowledge about the feature in question---the *prevalence prior* $P(p)$---with the uncertain semantics. -->

<!-- Because $P(\theta)$ is assumed to be a uniform distribution across all contexts, a generalization could in principle be felicitous with any referent-prevalence.  -->
<!-- However, the interpretation model's prior beliefs $P(p)$ constrains possible interpretations of the statement, and thus the endorsement.  -->

<!-- Our model makes context-sensitive predictions about endorsing generalizations based on two model components: The prevalence priors $P(p_fk)$ in Eq. \ref{eq:L0} and the referent-prevalence $p$ in Eqs. \ref{eq:S1} and \ref{eq:delta}. -->
<!-- Referent-prevalence is the probability that an individual $x$ of the referent-category (e.g., *dogs*) will have the feature being predicated of the category (e.g., *barks*). -->
<!-- We will return and refine this definition in Case Study 2. -->
<!-- For now, the referent-prevalence is simply a probability that determines the truth-conditional semantics.  -->
<!-- Our hypothesis does not provide a complete mapping from the referent-prevalence directly to endorsement judgments, however, because $\theta$ is underspecified. -->
<!-- These internal models may have domain-specific features but can be described by the domain-general currency of inductive belief: probability. -->

<!-- Since *cue validity* implicitly encodes knowledge of other categories (by summing over $k'$) and the prevalence prior explicitly encodes knowledge of other categories, how are these two constructs related? -->
<!-- In the prevalence prior $P(p_f)$, a category is represented only by its corresponding prevalence (i.e., a sample from the prevalence prior can be thought of as a category $k$, with the only information available about the category being the prevalence of the feature in that category $p$).  -->
<!-- Thus, $P(p) = P(k_{f})$. -->

<!-- cut on 2/9: -->
<!-- The fact that the prevalence prior distribution relates only via its mean value to *cue validity* reveals a potential shortcoming of cue validity: By reducing a distribution to a point estimate, the structure of the distribution is lost.  -->
<!-- This is particularly problematic for prevalence distributions that encode conceptual knowledge via multimodal distributions (e.g., Fig. \@ref(fig:simulations)A; bottom row); the mean of a multimodal distribution is not a good representation of that distribution.  -->
<!-- We return to this point empirically in Case Study 1. -->

<!-- These distributions can be thought of as distributions over categories, displaying only the prevalence for the property in question. -->

<!-- F(left; filled colors) shows six prior distributions over feature-probabilities, representing the relevant background knowledge about each property. -->
<!-- For example,   -->

```{r loadCachedSimulations, cache = F}
load("cached_results/modelSims-priors_fixedT_uncertainT_speaker.RData") # sims.priors, sims.fixed.thresholds, sims.uncertain.thresholds, s1.simulations.relabeled
load("cached_results/modelSims-uniform_fixedT_uncertainT_speaker.RData") # rs.wp.l0, rs.listener.wp.tidy.samples, rs.tidy

s1.simulations.relabeled.2 <- left_join(
  s1.simulations.relabeled %>%  
    mutate(PriorShape = factor(example, 
                               levels = c("Dogs bark",
                                          "Kangaroos have spots",
                                          "Sharks don't eat people",
                                                      "Robins lay eggs",
                                                      "Robins are female",
                                                      "Mosquitos carry malaria"),
                               labels = c("bark", "have spots", "don't eat people", 
                                          "lay eggs", "are female", "carry malaria")),
              category = c("sharks", "robins", "robins", "kangaroos", "dogs", "mosquitos"),
              xlabpos = c(0.8, 0.5, 0.5, 0.325, 0.85, 0.325),
              ylabpos = c(0.12, 0.25, 0.25, 0.12, 0.12, 0.12)
           ),
                       data.frame(src =  c(rep("Prevalence priors", 6), rep("Interpretation posteriors",6 )),
                                  example = with(s1.simulations.relabeled, c(as.character(example), as.character(example))))
 ) %>%
   mutate(featureProb = ifelse(as.character(src) == "Prevalence priors", featureProb, -99),
          xlabpos = ifelse(as.character(src) == "Prevalence priors", xlabpos, -99))
```

```{r uniformPriorSimulation, fig.cap = "Computational model behavior assuming a uniform prior over prevalence. A: Interpretation model posteriors assuming different fixed thresholds (facets). High thresholds rule out more world-states (prevalence levels), leaving fewer world-states among which to distribute the full probability mass. B: Generic interpretation model averages over all thresholds to return a posterior distribution that favors higher prevalence levels in a graded manner. C: Endorsement model predicts higher rates of endorsemens as prevalence levels increase.", fig.width = 13, fig.asp = 0.3, out.width = "100%", fig.align = "center", cache = F}

fig.l0.thresholds <- ggplot(rs.wp.l0 %>% 
                              mutate(theta = round(theta, 2)) %>%
                              filter(theta %in% c(0, 0.25, 0.50, 0.75)), aes( x = state, 
                                           y = literalPosterior, 
                                       group = theta))+
  geom_line(size = 1)+
  geom_vline(aes(xintercept = theta), 
             color = 'darkred', size = 1)+
  scale_x_continuous(breaks = c(0, 1))+
  scale_y_continuous(breaks = c(0, 0.2))+
  ylab("Posterior probability mass")+
  xlab("Prevalence")+
  facet_wrap(~theta, nrow = 1)+
  ggtitle("Fixed-threshold interpretations")

fig.fullInterpretation = ggplot(rs.listener.wp.tidy.samples %>%
                           mutate(utterance = factor(utterance,
                                               levels=c("null","gen"),
                                               labels =c("silence", "generalization"))), 
       aes( x = support,fill = utterance, color = utterance))+
  geom_density(alpha = 0.4, size = 1.3)+
  scale_fill_solarized()+
  scale_color_solarized()+
  xlab("Prevalence")+
  ylab("Posterior probability density")+
  scale_x_continuous(breaks =c(0, 1))+
  scale_y_continuous(breaks = c(0, 1.5))+
  guides(fill = guide_legend(reverse = T), color = guide_legend(reverse = T))+
  ggtitle("Uncertain-threshold interpretation")+
  theme(legend.position = c(0.25, 0.8))

fig.endorsement.thresholds <- ggplot(rs.tidy %>%
                           mutate(utt = factor(utt,
                                               levels=c("null","gen"))), 
                         aes( x = state, y = prob, fill = utt, color = utt))+
  geom_col(size = 1, alpha = 0.4, color= 'black')+
  scale_x_continuous(breaks = c(0, 0.5, 1))+
  scale_y_continuous(breaks = c(0, 0.5, 1))+
  ylab("Endorsement probability")+
  xlab("Prevalence (Speaker beliefs)")+
  ggtitle("Endorsement predictions")+
  scale_fill_solarized()+
  scale_color_solarized()+
  guides(fill = F, color = F)

cowplot::plot_grid(
  fig.l0.thresholds,
  fig.fullInterpretation,
  fig.endorsement.thresholds,
  nrow = 1, labels = c("A", "B", "C"),
  align = 'h',
  rel_widths = c(1.5, 1, 1)
)
```

```{r simulations, fig.cap = "Model simulations assuming different prevalence priors. Top: Prevalence priors for six example features. Shapes of the priors were chosen to intuitively correspond to the properties labeling the distributions. Arrows show referent-prevalence for a target category. Bottom: Interpretation model posterior distributions over prevalence upon hearing a generalization about a novel category. Numbers at bottom correspond to endorsement model predictions for endorsing the generalization for the referent-category whose prevalence is shown in the top facets (e.g., \"Mosquitos carry malaria\").", fig.width = 9, fig.asp = 0.37, out.width = "100%", fig.align = "center", cache=F}

fig.sims.priors <- ggplot(
  bind_rows(sims.priors %>% 
              mutate(src = 'priors'),
            sims.uncertain.thresholds %>% 
              mutate(src = 'posteriors')
            ) %>% mutate(PriorShape = 
                           factor(PriorShape,  
                                  levels = c("barks", "hasSpots", "doesntEatPeople", "laysEggs", "isFemale", "carriesMalaria"),
                            labels = c("bark", "have spots", "don't eat people", "lay eggs", "are female", "carry malaria")),
                         src = factor(src, levels = c("priors", "posteriors"),
                                      labels = c("Prevalence priors", 
                                                 "Interpretation posteriors"))), 
  aes(x = state, fill = PriorShape, color = PriorShape))+
    geom_density(aes(y = ..scaled..), size = 0.6, alpha = 0.7)+
    theme_few() +
    scale_x_continuous(breaks = c(0, 1), limits= c(0, 1))+
    scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    # geom_rect(data = s1.simulations.relabeled.2,
    #           aes(xmin = featureProb - 0.05, xmax = featureProb + 0.05, ymin = 0.0, ymax = 0.15),
    #           fill = 'black', inherit.aes =F)+
    geom_segment(data = s1.simulations.relabeled.2,
              aes(x = featureProb, xend = featureProb, 
                  yend = 0.02, y = 0.2), size = 1,
              fill = 'black', inherit.aes =F, arrow = arrow(length = unit(0.2, "cm"),
                                                            type = 'open'))+
    geom_label(data = s1.simulations.relabeled.2,
                aes(x = xlabpos, y=0.25, label = category),
                label.padding = unit(0.075, "lines"),
                  family = 'Palatino', fontface = 'bold', size = 3.2, inherit.aes =F, force = 15)+
    ylab("Normalized probability density") +
    xlab("Prevalence")+
    scale_color_solarized()+
    scale_fill_solarized()+
    facet_grid(src~PriorShape, scales = 'free')+
    theme(strip.text.y = element_text(angle = 0, hjust =0),
          legend.position = "none", 
          plot.margin=unit(c(0.5,0.5,1.5,0.5),"cm"))

ggdraw(fig.sims.priors) + 
  draw_label(round(unique(select(filter(s1.simulations.relabeled.2, example == "Dogs bark"),endorsement))[[1]], 2), 
             0.12, 0.12, fontface = "italic", fontfamily = "Times") +
  draw_label(round(unique(select(filter(s1.simulations.relabeled.2, example == "Kangaroos have spots"),endorsement))[[1]], 2), 
             0.25, 0.12, fontface = "italic", fontfamily = "Times") +
  draw_label(round(unique(select(filter(s1.simulations.relabeled.2, example == "Sharks don't eat people"),endorsement))[[1]], 2),
             0.36, 0.12, fontface = "italic", fontfamily = "Times") +
  draw_label(round(unique(select(filter(s1.simulations.relabeled.2, example == "Robins lay eggs"),endorsement))[[1]], 2), 
             0.48, 0.12, fontface = "italic", fontfamily = "Times") +
  draw_label(round(unique(select(filter(s1.simulations.relabeled.2, example == "Robins are female"),endorsement))[[1]], 2), 
             0.6, 0.12, fontface = "italic", fontfamily = "Times") +
  draw_label(round(unique(select(filter(s1.simulations.relabeled.2, example == "Mosquitos carry malaria"),endorsement))[[1]], 2), 
             0.72, 0.12, fontface = "italic", fontfamily = "Times") +
  draw_label("Endorsement Prediction", 0.885, 0.12, size = 10)
```

# Worked examples

<!-- The endorsement model defined in the previous section predicts that the variability in endorsements for generic statements and other sorts of generalizations in language can be attributed to two factors: (i) the referent prevalence in the category under discussion $p$ in Eq. \ref{eq:S1} (e.g., $P(x \text{ lays eggs} \mid x \text{ is a robin})$) and (ii) the background knowledge about prevalences---the prevalence prior---$P(p)$ (Eq. \ref{eq:L0}). -->
We explore the predictions of the endorsement model in the context of a few worked examples.
These examples are taken from the linguistics literature on generics and are statements that have been historically challenging for prevalence-based approaches (see Table 1).
We implemented these and all subsequent Bayesian models in the probabilistic programming language WebPPL [@dippl].
All models, analyses, data and links to experiments used in this paper can be found at \url{https://mhtess.github.io}.

Endorsement predictions depend upon the uncertain threshold model's interpretations, which are highly sensitive to background knowledge.
Background knowledge may be richly structured, reflecting intuitive theories about the underlying causes of different kinds of properties [cf., @Leslie2007].
This model posits, however, that the only impact of structured knowledge on truth judgments is in their implications for beliefs about prevalence, formalized in terms of a prevalence prior $P(p)$. 
\ndg{i’d save this for expt sections, and here just say that we posit intuitive priors for illustrations and introduce a method to actually elicit them below}
\mht{without this, we don't have an explanation for what the prevalence prior is, which is important for understanding the examples... i removed stuff about Betas, but check this again...}
The prevalence prior reflects a listener's beliefs about the prevalence of the feature in an unknown or unfamiliar category and be conceptualized in a theory-neutral way as a distribution over the prevalence among alternative categories (Figure\ \@ref(fig:simulations)).^[
  Alternatively, the prevalence prior can be conceptualized as one aspect of an intuitive theory of a domain. 
]
<!-- For example, think about five of your favorite kinds of animal: What percentage of each of them *are female*? What percentage of each of them *lay eggs*? -->
For example, the *lay eggs* prevalence prior should be bimodal, with substantial probability mass near 0-prevalence (since many animal categories do not have egg-layers) and a secondary component peaked around 0.5 (because among the animals with egg-layers, only the female members of the category have the property).
Conversely, the distribution over the prevalence of *being female* is unimodal and centered at 50%-prevalence, because almost all animals have female members in that proportion.
Figure\ \@ref(fig:simulations) (row 1) shows these and other intuitive prevalence priors for different properties (we introduce methods for empirically eliciting these priors in the experimental section).
We now turn our attention to the endorsement model's predictions.

<!-- We simulate how the prevalence priors change the literal interpretation and how they and the referent prevalence impact endorsement using schematic parameter values based on intuition from a few examples in the literature on generic language. -->

<!-- The literal interpretation of a generalization (Eq. \ref{eq:L0}) reflects a balance between an uncertain threshold and shared background knowledge about the feature in question. -->
<!-- TELL US WHAT THEY ARE: BETA DISTRIBUTIONS./ -->

```{r workedExamples, results="asis"}
worked.examples <- data.frame(
  Example = c("1. Dogs bark",
              "2. Kangaroos have spots",
              "3. Robins lay eggs",
              "4. Robins are female",
              "5. Mosquitos carry malaria", 
              "6. Sharks don't eat people",
              "7. Mary handles the mail from Antarcita",
              "8. Supreme Court Justices have even social security numbers",
              "9. Elephants live in Africa and Asia"),
  `referent prevalence` = c("95%", "5%",  "50%", "50%", "5%", "85%", "0 times", "100%", "50%/50%"),
  `intuitive truth value` = c(
    "True", "False",  "True", "False",
    "True", "False","True", "False", "True" 
  ),
  `potential issue` = c(
    "Not all dogs bark",
    "Some kangaroos could have spots",
    "Only female robins (50%) lay eggs",
    "Same number that lay eggs (50%)",
    "Very few mosquitos actually carry malaria",
    "Most sharks don't eat people",
    "Consider there has never been any mail from Antarcita",
    "Consider 100% have even social security numbers",
    "Impossible for most to live in Africa and most to live in Asia"
  )
) %>%
  rename("Referent prevalence" = referent.prevalence,
         "Intuitive truth value" = intuitive.truth.value,
         "Issue" = potential.issue)

worked.examples.tab <- xtable::xtable(worked.examples, 
                       caption = "Example sentences that a theory of generalizations should correctly predict. See details of each example in corresponding section.", 
                       align=c('| p{1.5in}', 'p{2in}| ',' |p{1.25in}|','p{0.75in}|','p{2in} |'))

addtorow          <- list()
addtorow$pos      <- list()
addtorow$pos[[1]] <- c(0)
addtorow$command  <- c(paste("\\hline \n",
                             "\\endhead \n",
                             "\\hline \n",
                             "{\\footnotesize \\em{continued on next page}} \n",
                             "\\endfoot \n",
                             "\\endlastfoot \n",sep=""))

large <- function(x){ paste0('{\\Large ', x, '}') }
bold <- function(x){ paste0('{\\bfseries ', x, '}') }

print(worked.examples.tab, 
      type="latex", 
      comment = F,
       #table.placement = "h", 
      #tabular.environment = "longtable",
      hline.after=-1:nrow(worked.examples.tab),
      size="\\fontsize{9pt}{10pt}\\selectfont", 
      include.rownames=FALSE,
      sanitize.colnames.function = bold
      #add.to.row = addtorow
      )
```

**1. Dogs bark.**

Often the first observation with generics is that they appear to behave like universal quantifiers (e.g., "All") that permit exceptions: Not all dogs bark, but still "Dogs bark" is true.
The uncertain threshold account immediately accomodates this example because there is no single, fixed threshold beyond which a generic statement becomes true.
Rather, listeners have uncertainty about the threshold which leads to graded interpretations (e.g., "Dogs bark" probably means *almost all* dogs bark; Figure\ \@ref(fig:simulations); column 1, row 2).
The endorsement model predicts "Dogs bark" is a rather good generic sentence (Figure\ \@ref(fig:simulations), bottom row) even though there are exceptions to the universal generalization (referent prevalence of barking among dogs shown with arrow in Figure\ \@ref(fig:simulations) row 2).

**2. Kangaroos have spots.**

A second observation regarding generics is that their truth conditions cannot be so lenient so as to always convey *existential* quantification (e.g., "Some").
Indeed, it's intuitively plausible that some kangaroos do have spots but a rational language user might feel awkward to assert the generic "Kangaroos have spots".
The model exhibits this same restraint, predicting a very low endorsement probability for this generic sentence (Figure\ \@ref(fig:simulations); column 2).
The reason is that, given the interpretation model's background knowledge about the property, the statement "Kangaroos have spots" will be interpreted like "Dogs bark": it would mean *almost all* kangaroos have spots. 
The endorsement model, which believes that very few kangaroos have spots, then would rather not endorse the statement because it would lead to a too-strong interpretation and mislead the listener.

**3. and 4. Robins lay eggs vs. Robins are female**

One of the most difficult examples for a theory of generics based on prevalence to handle is the intuitive difference in truth value between the statements "Robins lay eggs" and "Robins are female". 
The former is intuitively true, even though only female robins could lay eggs (and hence, the prevalence is roughly 50%).
However, the same implicit restriction to only females does not seem occur for the latter statement, as "Robins are female" strikes most as strange or false.
Why is "Robins lay eggs" a reasonable utterance while "Robins are female" is not?

The prior distributions over the prevalence of both features are shown in Figure\ \@ref(fig:simulations) (columns 5 and 6).
As described above, the priors are different: Many animals have zero egg-layers (0% prevalence), while the vast majority of animal categories have female members in exactly the same proportion (50%).
Given this background knowledge, the generic interpretation model returns roughly the same interpretation distribution for each hypothetical utterance: In each case, the model believes roughly 50% have the property.
However, only in the case of "Robins lay eggs" does the endorsement model actually assert the generic; it does so because the listener interpretation would be more aligned with the referent prevalence in comparison to the interpretation of the null utterance, which is the prevalence prior.
In contrast, the hypothetical generic interpretation of "Robins are female" is not different from the prevalence prior and hence the generic conveys no new information; here, the generic would not be misleading but uninformative, and hence the model predicts an endorsement probability of 0.5.
Indeed, previous studies on generic endorsements have found "Robins are female" to not be rated as completely false but rather receive an intermediate endorsement level [i.e., *neither true nor false*; @Prasada2013].

**5. Mosquitos carry malaria.**

The statement "Mosquitos carry malaria" is intuitively true in spite of the fact that the vast majority of mosquitos are actually malaria-free.
The prevalence prior for *carries malaria* is highly skewed towards low prevalence levels: many animals do not have malaria-carriers among them and even for those that do, the prevalence is expected to be quite low.
Then, carrying malaria is significantly more true of mosquitos than other animals, an intuition that is often arrived at with this example.
This kind of behavior is related to the construct of *cue validity*---the probability of the feature given the category e.g., $P({is a mosquito} \mid \text{carries malaria})$---which we return to in the next section.
The uncertain-threshold model displays a critical behavior: It can endorse generics when the referent prevalence is very low.

<!-- and then point out that this is related to cue validity, as explored later. (or just wait to introduce cue validity until later.) -->
<!-- ---because the listener interpretation distribution would be more aligned with the referent prevalence than the prior.  -->
<!-- Intuitively, this is because many animals do not have malaria-carriers among them and so in a sense, having malaria is diagnostic for being a mosquito -->
<!-- We return to this relationship in the next section, and provide a precise mathematical derivation of the relationship between the prevalence prior and cue validity in Appendix A. -->

**6. Sharks don't eat people.**

When the prevalence of the feature is very high, it is not necessary for the model to endorse the generic.
"Sharks don't eat people" is predicted to be a somewhat strange utterance, despite the fact that the vast majority of sharks do not eat people.^[
 We create this example as the converse of the classic example "Sharks eat people", predicted to be *true* despite low prevalence [@Leslie2007]. We reverse this example to show how the model deals with a high-prevalence feature for which the generic is predicted to be false or infelicitous.  
] 
<!-- Yet, intuitively, the generic would be misleading (similar in intuition to "Kangaroos have spots"). -->
Because for almost all animal categories, the prevalence of *not eating people* is almost 100% (very few things eat people), interpreting the statement "Sharks don't eat people" would lead one to believe that *no* sharks eat people (i.e., 100% don't eat people), which is too strong.
It is interesting to note that because of the high referent prevalence, the model is less certain of its decision, predicting an endorsement probability around 40% (i.e., somewhat false).
We will test this quantitative prediction in Experiment 1 on generic language.

<!-- Hearing "Ks bark" tells the literal interpreter that *almost all* Ks bark, though there is a good deal of uncertainty in the precise quantity. -->
<!-- "Ks lay eggs" has a strikingly different interpretation: It means "Half of Ks lay eggs" (column 4). -->
<!-- The interpretation model can also derive very weak interpretation, as shown in the  *carries malaria* distribution (column 6). -->
<!-- The endorsement model (Eq. \ref{eq:S1}) predicts endorsement probabilities based on the literal interpretation of the utterance and the referent prevalence (e.g., how many robins lay eggs; shown in arrows in Figure\ \@ref(fig:simulations)A). -->
<!-- Predictions for a few examples are shown in the bottom row of Figure\ \@ref(fig:simulations). -->
<!-- In accord with intuition, the endorsement model predicts "Dogs bark", "Robins lay eggs", and "Mosquitos carry malaria" are all true, despite the referent-prevalences varying substantially between these three examples.  -->
<!-- Intriguingly, the model predicts "Robins are female" (which has the same prevalence as "Robins lay eggs") should receive intermediate endorsement: It is neither true nor false. -->
<!-- To understand this prediction, examine the interpretation distribution for the *are female* property: It is identical with the prior (Figure\ \@ref(fig:simulations); column 5; top vs. bottom row). -->
<!-- The endorsement model decides that is neither better nor worse to produce the generalization than to stay silent, as they both carry the same information content. -->
<!-- This example can be contrasted with cases such as "Sharks don't eat people" and "Kangaroos have spots", which the model thinks would be misleading to say (i.e., better to say nothing). -->

<!-- #### Other philosophical puzzles -->

<!-- In this paper, we aim to balance two competing goals in our experiments: (1) drawing upon familiar, philosophically interesting examples that have been discussed in the vast literature on generic language and (2) introducing novel stimuli that provide targeted quantitative tests of our model vis-a-vis alternative models. -->
<!-- We close by discussing how three previously-discussed, philosophically interesting examples could be addressed in our communicative framework. -->

**7. Mary handles the mail from Antarctica (*yet has never had the opportunity*).**

Imagine there is a job in the local bureaucrat's office to handle the mail from Antarctica and this job is assigned to Mary; to date, however, nobody has ever sent mail to the office from Antarctica [@Cohen1999]. 
In other words, the statement "Mary has handled mail from Antarctica" is false.
"Mary handles the mail from Antarctica", however, is still thought to be intuitively true despite zero actual instances of the event.
This highlights an important ambiguity in the theoretical commitments of the uncertain threshold model: The endorsement model aims to communicate some referent prevalence $p$, but does the prevalence represent the actual, objective frequency in the world (e.g., the number of times in the past that Mary has handled mail from Antarctica) or a subjective, predictive degree of belief in the head (e.g., our prediction that were the appropriate situation to arise, Mary would be handling the mail from Antarctica)?

We posit that Mary the Antarctic mail handler is an extreme case of generalizations expressing *predictive* degrees of belief.
We expect this context sets up the expectation that any future mail coming from Antarctica will be handled by Mary. 
<!-- In Expt. 2c., we experimentally dissociated predictive probability from past frequency, finding that generalization endorsement tracked predictive probability, but not frequency. -->
<!-- We expect the same to be true with Mary: Any future mail coming from Antarctica will be handled by Mary.  -->
Thus the predictive probability that Mary will handle Antarctican mail, should there be some, is high.
Predictive probabilities often track past frequency or actual prevalence in the world, but people's internal models of how the world works can lead them to diverge. 
In this case our understanding of Mary's job leads to a strong predictive probability in the absence of past frequency evidence.
We explore this question experimentally in Exerpiment 2 on habituals.

**8. Supreme Court Justices have even social security numbers.**

Imagine that all current Supreme Court Justice has a social security number which was an *even number*. 
"Supreme Court Justices have even social security numbers" is still considered false, even though the property holds for exactly 100\% of the category [@Cohen1999].
We predict the rejection of even social security numbered Justices is the result of our intuitive theories guiding our subjective predictive probabilities, which feed into the endorsement model.
That is, we predict observers strongly believe there is no causal relation between the evenness of one's social security number and selection for the Supreme Court, and thus would assign a roughly 50\% subjective probability to the *next* justice having an even social security number.
Then, "Supreme Court Justices have even social security numbers" would be rated by our model as similar to "Birds are female", because all professions have roughly the same probability of having employees with even social security numbers.
This example is thus the conceptual opposite of Mary the Antarctic mail handler, where our internal model of the world led us to a strong degree of belief in the property holding in the future; with the Supreme Court Justices, our internal models lead to a relatively weak degree of belief in the property holding in the future.

**9. Elephants live in Africa and Asia.**

<!-- \mht{both this and the "surprisng prediction" above are intutiions/examples we are not going to test empirically (however, here we at least give a model for it)... should these appear elsewhere, or somehow be marked upfront?} -->

Understanding how a semantic representation *composes* is another important test for a theory of generalizations in language.
@Nickel2008 suggests that "Elephants live in Africa and Asia" is troubling for prevalence-based accounts [in particular, majority-quantificational accouts where the generic means *more than half*; @Cohen1999] because the statement should be semantically equivalent to "Elephants live in Africa and elephants live in Asia".
It cannot be the case that most (more than half of) elephants live in Africa and *most* (more than half of) elephants live in Asia, unless we are to posit *international elephants* (i.e., individual elephants who live part-time in Africa and part-time in Asia), which are intuitively implausible.
<!-- However, this is the only interpretation available if the generalization is given a fixed, majority-based quantificational semantics, as posited in most statistical views of generics [e.g., if the generic means *most*; @Cohen1999]. -->

Our theory does not provide a fixed semantics for generics, but an uncertain one which can be updated as more information comes in.
In fact, with prior knowledge suggesting against the existence of international elephants (Figure\ \@ref(fig:elephants)A), our model interprets "Elephants live in Africa and Asia" as meaning that some elephants live in Africa and that different ones live in Asia (Figure\ \@ref(fig:elephants)C).
Interestingly, an incremental parsing of the sentence  (i.e., upon hearing only that "Elephants live in Africa") leads our model to believe that *most*, possibly all, elephants live in Africa (Figure\ \@ref(fig:elephants)B).
When the sentence is completed ("...and Asia"), the model non-monotonically updates its beliefs to something weaker: *some* elephants live in Africa and others in Asia (Figure\ \@ref(fig:elephants)C).
The flexibility of our model accommodates new evidence that might otherwise contradict previous utterances because it maintains uncertainty about the precise meaning of the utterance.
From a compositional perspective, this is highly desirable behavior but we leave the testing of quantitative predictions of this kind for future work.

```{r elephants, fig.cap="Model interpretation of a conjunctive generic (\"Elephants live in Africa and Asia\"). A: Intuitive priors for marginal distributions of prevalence for living in Africa and living in Africa. B: Partial interpretation upon hearing utternace (\"Elephants live in Africa...\"). C: Full interpretation upon hearing the end of the sentence (\"Elephants live in Africa and Asia\"). Uncertainty about the threshold for generic sentences allows the listener to non-monotonoically update its beliefs about the prevalence of the feature in categories.", fig.width = 8, fig.asp = 0.6, out.width = "100%", fig.align = "center", cache=F}
load("cached_results/modelSims-elephants.RData")

fig.eleph.0 <- get_samples(rs.asiaafrica.priors, 10000) %>%
  ggplot(., aes(x = state))+
  geom_density(adjust = 2, fill = 'grey', alpha = 0.3, size = 1.5)+
  facet_wrap(~marginal, ncol = 1)+
  ggtitle('Prevalence priors')+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 1))+
  xlab("")+
  ylab("Probability density")+
  theme(plot.title = element_text(size = 10, hjust = 0.5))

fig.eleph.1 <- get_samples(rs.asia.marignals, 10000) %>%
  ggplot(., aes(x = state))+
  geom_density(adjust = 4, fill = 'grey', alpha = 0.3, size = 1.5)+
  facet_wrap(~marginal, ncol = 1)+
  ggtitle('"Elephants live in Africa..."')+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 1))+
  xlab("Prevalence")+
  ylab("Probability density")+
  theme(axis.title.y = element_blank(),
        plot.title = element_text(size = 10, hjust = 0.5))

fig.eleph.2 <- get_samples(rs.asiaafrica.marignals, 10000) %>%
  ggplot(., aes(x = state))+
  geom_density(adjust = 4, fill = 'grey', alpha = 0.3, size = 1.5)+
  facet_wrap(~marginal, ncol = 1)+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 1))+
  xlab("")+
  ggtitle('"...and Asia"')+
  ylab("Probability density")+
  theme(axis.title.y = element_blank(),
        plot.title = element_text(size = 10, hjust = 0.5))

cowplot::plot_grid(
  fig.eleph.0,
  fig.eleph.1,
  fig.eleph.2,
  nrow = 1,
  rel_widths =c(1,1, 1),
  labels = c("A", "B","C")
)
```

# Relationship to other semantic theories

<!-- We presents the first formal theory of the language of generalization that generates quantitative predictions about human behavioral data.  -->
Our formal theory builds on and relates to a number of extant theories of generics in both the formal semantics and psychological literatures.
The dominant approaches from formal semantics try to describe unified, objective criteria by which to establish if a generic and other generalizations are true or false (i.e., generic *truth conditions*).
These views are typically *statistical* in nature in that they appeal to quantification or the statistics of the world (e.g., how many of the kind have the property) to define those unified, objective criteria.
Many of these theories rely upon mechanisms outside of the truth conditions to make sense of the extreme flexibility exhibited by generics; unfortunately, these extra-semantic mechanisms are not described in sufficient detail to generate precise, quantitative predictions.
Other theorists, primarily in philosophy and psychology, have taken the extreme flexibility of generics to argue against a quantitative theory based on statistics, instead suggesting that abstract, mental representations are directly tied to semantics of generics (e.g., *there is something about being a K which causes it to F*).
Statistical and conceptual theories express the major contrasting views of generic language [@Carlson1995essay].^[
  We use the terms statistical and conceptual to refer to what @Carlson1995essay referred to as "inductive" and "rules and regulations" views, respectively.
]

## Conceptual accounts

Conceptual accounts try to identify the core meaning of a generalization directly with aspects of conceptual structure.
The most influential account in psychology comes from @Leslie2008, which uses the same starting point as our analysis: Generics express generalization. 
@Leslie2008's analysis draws upon insights from the psychological literature on infant generalization to argue that generics tap into an innate, default mechanism that signals the child (or adult) to generalize the property to the kind.
<!-- ^[ -->
<!--   Because the mechanism for understanding generics is innate, in this account, the learning puzzle (vis a vis quantifiers) immediately goes away. -->
<!-- ] -->

The *default generalization* mechanism has three components: the ability to (i) identify whether or not a feature is a *characteristic* property of a kind (e.g., one characteristic property for an animal would pertain to the mode of locomotion for the animal), (ii) identify whether or not a feature is *striking* (e.g., can kill you; "Mosquitos carry the West Nile Virus"), and (iii) segment *counterinstances* (instances of the kind which do not have the feature) into positive and negative counterinstances (i.e., instances of the kind which have some relevant alternative feature and instances of the kind which simply lack the feature, respectively).^[
  This positive vs. negative counterinstance distinction is a rather technical consideration [though not without psychological foundation, see @Leslie2008, p.36] used to account for the difference between examples like "Robins lay eggs" vs. "Robins are female". 
  For "Robins are female", male robins are positive counterinstances because they have an alternative property (being male). 
  For "Robins lay eggs", male robins are negative counterinstances because they simply lack the property (i.e., male robins do not reproduce by some other method).
  We will not discuss this consideration any further other than to note that that the difference between "Robins lay eggs" and "Robins are female" can be explained by a different mechanism (e.g., informativity with respect to a prevalence prior).
  The notion that alternative features may come into play in generic interpretation may also be addressed with prevalence priors constructed with respect to the feature (described in the General Discussion).
]
On this account, generics are true if *some* instances of the kind have the feature when all counterinstances are negative counterinstances and the property is either characteristic or striking. 
If the property is neither characteristic nor striking, then *almost all* instances of the kind must have the feature for the generic to be true (e.g., "Barns are red").
Many of the key factors in this conceptual approach are compatible with our view of prevalence as a predictive probability: What is required is that the distinctions made by conceptual accounts influence predictive probability, which mediates linguistic effects.
Indeed, one version of this mechanism is already suggested by @Leslie2008, citing findings by @Rothbart1978, that a speaker’s perception of the prevalence of the feature can be altered by virtue of its dangerousness or distinctiveness.

### Default generalizations

The conceptual view of generics argues that it is not desirable to define generics in the same terms as quantifiers (e.g., a truth-functional threshold on prevalence), because generics are more basic or more fundamental than quantifiers [@Leslie2008; @Gelman2009].
For example, according to certain measurements, young children have a nearly adult-like understanding of generics far earlier than they do with quantified language [@Gelman2015].
Similarly, adults who are told novel information about categories using quantified language (e.g., "Most spiders shed their skin") and later asked to recall that same information will tend to recall quantified information as generics [e.g., "Spiders shed their skin"; @Leslie2012], but not visa versa.
As a result, generics are thought of as conveying default generalizations, whereas quantified language expresses more sophisticated and specific generalizations.
Rather than be contradictory, our model can be seen as formalizing a default generalization as Bayesian belief updating via a threshold function whose threshold value is contextually-informed via knowledge about properties. 
Alternative formulations of such a default generalization mechanism could be proposed and quantitatively tested against our account.

<!-- Rather than be contradictory to a prevalence-based truth-functional semantics, these observations independently motivate our treatment of the semantics as *underspecified*. -->
<!-- From a probabilistic learning perspective, a child learning the semantics of generics (and quantifiers) need learn  -->
<!-- Probabilistic agents typically have uncertainty about variables of interest in their environment; having certainty about a random variable is a rare and special case. -->
<!-- In a probabilistic representation, then, states of uncertainty are more fundamental than states of uncertainty. -->
<!-- Therefore, if quantified language can be described with *certain* thresholds  (e.g., *some* means greater than *zero*, *most* means greater than *half*), then we would expect generics, as a more fundamental linguistic device, to have an *uncertain threshold*. -->
<!-- Thus, in our framework, generics behave like a more fundamental (i.e., more uncertain) quantifier. -->
<!-- One distinctive feature of generics that this informal analysis does not address is that generics are *intensional* (e.g., having to do with predictions are the future) whereas quantifiers tend to be *extensional* (i.e., concerned with the actual frequencies in the world).  -->
<!-- How and why the intensional/extensional distinction emerges between quantifiers and generics is an outstanding issue for a theory such as ours that tries to unify the linguistic analysis of these two kinds of expressions. -->

### Striking generics

@Leslie2008 posits a special mechanism to treat generics about striking properties (e.g., "Rottweilers maul children"), which seem to be acceptable even when the prevalence of the property is quite low (e.g., very few rottweilers maul children).
Indeed, @Cimpian2010 found participants willing to endorse generics about striking properties (e.g., "Lorches have dangerous feathers") more so than that of neutral properties (e.g., "Lorches have purple feathers") at low levels of prevalence (e.g., when only 30% of the category had the property).
@Leslie2008's argument for accepting striking generics is straightforward: These are relevant properties to know about if you want to survive. 
Our theory posits two mechanisms by which striking properties could influence generic endorsement: the prevalence prior and the referent prevalence.

The first observation is that striking properties are relatively rare in the environment: Most animals do not maul children, eat swimmers, or carry malaria.
Thus, the prevalence prior distributions of these features may be hard to distinguish from properties that are just generally rare and distinctive, as we have assumed in our worked example of "Mosquitos carry malaria".
<!-- ; see also our Expt. 2a, e.g., the distributions for *steals cars* vs. *climbs mountains* for an example with habituals).  -->
@Cimpian2010's experiments did not measure the prevalence prior distribution, but we have found in pilot work that the prevalence prior distribution changes to resemble that of a distinctive property when participants are supplied information about the dangerousness of a property.

The second mechanism by which strikingness could influence generic endorsement in our model is by speakers having a distorted perspective of how prevalent these features are within the referent category.
That is, striking properties may be projected more strongly (i.e., higher predictive probability that future instances will have the property) than neutral properties.^[
  This observation is also made by @Leslie2008 (p.42), who uses it as motivation for elevating striking properties to their special status in her theory.
]
Interestingly, evidence suggests this enhanced projectibility holds for both dangerous properties (e.g., people doing criminal actions) as well as neutral, distinctive properties [e.g., people taller than 6'5", @Rothbart1978 Expts. 2 & 3].
Thus, there is evidence for the influence of strikingness on predictive probability, which we posit is an intermediate representation between conceptual knowledge and generic endorsements.
It is beyond the scope of this paper to directly attempt to empirically distinguish these two potential mechanisms, though our experiments use a number of striking properties for which we measure the prevalence prior and referent prevalence.

 <!-- (Expt. 4) -->
<!-- and explicitly distinctive properties (e.g., "Lorches have distinctive feathers") about equally as often each other and more so than neutral properties (e.g., "Lorches have purple feathers").  -->
<!-- This similar increase in endorsement for striking and distinctive properties could be explained by the prevalence prior $P(p)$ distributions for the two being similar, and distinct from that of the neutral properties. -->
<!-- The first is the prevalence prior. -->
<!-- Thus, strikingness may influence endorsement by either (i) altering the prevalence priors or (ii) influencing speaker's predictions about future prevalence; we do not know of any evidence that shows an effect of strikingness while controlling for these mediating factors. -->
<!-- If there are none, then the effects of strikingness would be accounted for by our prevalence-mediated theory. -->

### Principled and statistical connections

@Leslie2008's construct of characteristic properties is similar to @Prasada2006's notion of a *k-property* (*k* for kind), a property that bears a *principled connection* to the kind. 
A property bears a principled connection to the kind if a (generic) sentence that appeals to the kind to explain the existence of the property makes sense (e.g., "Dogs, by virtue of being the kind of things that they are, are four-legged").
When such sentences are not endorsed (e.g., "Barns, by virtue of being the kind of things that they are, are red") while the simple generic (e.g., "Barns are red") is endorsed, then the property is thought to bear a *statistical connection* to the kind (a so-called *t-property*, for emphasizing that it is the *tokens* having the property which matters for generic endorsement).
@Prasada2013 showed that different kinds of generic statements (e.g., striking properties, characteristic properties with low prevalence) support different kinds of inferences in addition to the difference in endorsements for *by virtue of* statements (e.g., *normative*: "Dogs are supposed to have four legs", *aspect*: "Having four legs is one aspect of being a dog").
Additionally, there is some evidence that the impact of principled connections on interpretations is separate from their influence on beliefs about prevalence [@Prasada2006].

<!-- Where do principled connections come from?  -->
Both @Prasada2006 and @Leslie2008 surmise that a property may qualify as characteristic (or having a principled connection) if there exists an overhypothesis about that property for a relavant superordinate category [e.g., each kind of animal has a means for self-locomotion; @Goodman1955; @Shipley1993].
Overhypotheses are naturally formalized as hierarchical Bayesian models, wherein a learner acquires knowledge at multiple levels of abstraction [e.g., learning from the same event about a particular dog, dogs in general, and animals in general; @Kemp2007].
These Bayesian heirarchical models yield differences in predictive probability that are particularly robust.
Integrating a hierarchical model of kinds and properties with our model of generic language is a natural direction to understanding the computational underpinnings of conceptual relations and generics. 
Different conceptual structures may give rise to roughly the same distributions on prevalence and thus have similar generic endorsement profiles according to our model. 
Yet the inferences that can be drawn from hearing these generics may differ, depending on their interactions with the putative hierarchical knowledge people bring to bear [a la the differences observed for "prevalence-matched items" in @Prasada2006]. 
Building hierarchical models of kinds and properties is a major undertaking in its own regard, but our formalism provides a way of connecting such a model with generic language.

<!-- , and thus, even though the generic can imply the same prevalence, listeners have overall different inferences (e.g., about the *virtue of* sentence). -->

<!-- The focus of this other empirical work is about highlighting the differences among generics, while our focus is to highlight what is common. -->

<!-- Rich, structural mental representations can be formalized in probabilistic models, which can be used to derive quantitative predictions about inferences derived from that knowledge [@Tenenbaum2011].  -->
<!-- For example, @Kemp2008 introduced a model for learning different kinds of structural relationships between objects and properties (e.g., taxonomies, social cliques, dominance relations). -->
<!-- @Goodmanconcepts argue that concepts and intuitive theories can be constructed from elementary random primitives using insights from *probabilistic programmming*.  -->

<!-- our account is really about what is common among generics, and this is about what is differnet -->
<!-- this is an interesting, albeit crude, result. if true, then this is really about different conceptual structures giving rise to roughly the same distributions on prevalence. even though the generic can imply the same prevalence, you have overall different inferences. our account is really about what is common among generics, and this is about what is differnet. -->

<!-- ### Conceptual accounts -->

<!-- Conceptual accounts of generics emphasize the structure of generic knowledge [@Prasada2000], and view generic utterances as the way of expressing special mental relationships between kinds and properties [@Leslie2008; @Prasada2012]. -->
<!-- These accounts start with the perspective that generics express rich, conceptual relationships between kinds and properties. -->
<!-- That is, "Bishops move diagonally" not because of a statistical relationship, but rather, because those are the rules of the game; if you start moving the bishop other ways besides diagonally, you cease to be playing chess. -->

<!-- If a lion lacked the capacity to roar, we might consider a less good example of being a lion because "lions roar". -->

<!-- The most influential conceptual account of generics is from @Leslie2007. -->
<!-- For Leslie, generics are tied to the cognitive system's "default mode of generalization", which infants seem ready to make use of by drawing strong generalizations from small amounts of data [e.g., @Baldwin1993]. -->
<!-- Leslie's "default mode" comes equipped with the ability to single-out *striking properties* (e.g., properties which are dangerous or appalling) as particularly useful aspects of the world to know about. -->
<!-- Hence, "Mosquitos carry malaria" is true because malaria can kill you and is useful to know about. -->

<!-- This accounts predicts that generics that convey striking properties should be endorsed even when they are relatively rare in the category (i.e., low referent-prevalence; e.g., "Sharks attack swimmers"). -->
<!-- Indeed, naturalistic examples (e.g., "Sharks attack swimmers") and artificial examples (e.g., "Lorches have purple feathers")  supplemented with striking information (e.g.,"These feathers are as sharp as needles and can easily get lodged in you, causing massive bleeding") are endorsed when the prevalence is quite low (e.g., 30\% of lorches have purple feathers; @Cimpian2010; see Expt. 1 for naturalistic cases). -->
<!-- In pilot work, using a paradigm similar to the prior manipulation paradigm employed in the causal language experiments (Expt. 3a), we have observed that the prior distribution over the prevalence $P(h)$ change when supplied information about the dangerousness of the feature. -->
<!-- That is, our pilot work suggests that experiments that experimentally induce beliefs about the dangerousness of a feature also manipulate the prevalence priors.  -->
<!-- Thus, our model would also predict these effects, but by way of the prevalence priors rather than some direct connection between generics and striking features. -->
<!-- However, we do not dispute that extra-linguistic factors can play a role in endorsements of generalizations (e.g., a speaker may not *want* a hearer to believe the generalization, which she herself knows to be true, or visa versa; more on this below). -->

<!-- Of course, not all generics convey striking or appalling information (e.g., "Birds lay eggs"). -->
<!-- Leslie's default mode, thus, also distinguishes *negative counter-instances* of a property (e.g., a bird that doesn't lay eggs, such as a male bird) from *positive counter-instances*  (e.g., a hypothetical bird that bears live young) -->
<!-- The conceptual account argues that generics are much less reasonable when *positive* counter-instances exist.  -->
<!-- For example, "Birds are female" seems weird because *being male* is a *positive* counter-instance of *being female*, but since there are no birds that bear live young (i.e., no *positive counter-instances*),  "Birds lay eggs" is fine. -->
<!-- In the current version of our theory, both kinds of counter-instances influence the referent-category feature-probabilities, though only positive counter-instances could impact the statistical details of the comparison class (in particular, when constructed with respect to the feature; see discussion about the Comparison Class). -->
<!-- Thus, the notion of *positive* and *negative* counter-instances may result from a particular way of looking at the constructs posited in our model, namely prevalences within- and across- categories or features. -->

<!-- Finally, generics are not limited to conveying rich, conceptual information; they are also used to describe facts of the world.  -->
<!-- "Ravens are bigger than toasters" is a true statement, but not because of a rich relationship between ravens and toasters. -->
<!-- "Barns are red" because most barns are red. -->
<!-- If farmers around the world decided to paint their barns green, the things they would be painting would still be barns; in that world, we might say "Barns are green". -->
<!-- Thus, conceptual accounts must be supplemented with the proviso that in ceratin situations, the statistics of the world are relevant.  -->

<!-- Our view is that subjective probabilities can be formed either from worldly observations or because of the structure of people's mental representations. -->
<!-- In the last fifteen years, there has been tremendous progress in formalizing  -->
<!-- Rich, structural mental representations can be formalized in probabilistic models, which can be used to derive quantitative predictions about inferences derived from that knowledge [@Tenenbaum2011].  -->
<!-- For example, @Kemp2008 introduced a model for learning different kinds of structural relationships between objects and properties (e.g., taxonomies, social cliques, dominance relations). -->
<!-- @Goodmanconcepts argue that concepts and intuitive theories can be constructed from elementary random primitives using insights from *probabilistic programmming*.  -->

<!-- In this paper, we take a minimalist approach to modeling conceptual structure, incorporating only what is necessary into the interpreter's background knowledge, the prior on prevalence $P(h)$. -->
<!-- In our Bayesian data analysis of background knowledge, we use a mixture model that reflects basic knowledge about kinds and properties (e.g., that most kinds don't have most properites).   -->
<!-- It's plausible that more richly structured knowledge of the kind described by @Prasada2013 and @Leslie2007 can be incorporating into the general probabilsitic framework that we use here. -->

<!-- In Expt. 1a, we discovered that a mixture model is necessary to account for intuitions about the prevalence across different categories. -->
<!-- This minimally structured model  -->

<!-- @Prasada2006 and later @Prasada2013 distinguish between \emph{principled}, \emph{statistical}, and \emph{causal} relations within concepts. -->
<!-- Striking generics (e.g. \emph{Mosquitos carry malaria}) show characteristics of \emph{causal connections} (operationalized using the phrase \emph{There is something about Ks that cause them to F}). -->
<!-- Generics like \emph{Birds lay eggs} (in which only a minority have the property) exhibit *principled* connections (operationalized by endorsement of the phrase \emph{In general, Ks have F}). -->
<!-- The fact that generics about different properties license different kinds of inferences is taken as evidence that the generics themselves represent different kinds of relations. -->
<!-- Statistical information takes a backseat to the conceptual structure. -->

<!-- It is also the first attempt in psychology to try to unify significant swaths of language that are seemingly quite different from one another: generalizations about events, causes, and categories. -->

<!-- -- Leslie (2008): Default inferences, could be similar to the cognitive model described, but we don't make strong metaphysical commitments (e.g., to "negative" properties) -->

<!-- -- "default generalization" -- very close, the listener is doing a generalization, but not generalization based on evidence, but by reasoning about the threshold -->

<!-- Generic language is the simple and ubiquitous way by which generalizations are conveyed between people. -->
<!-- Yet the dramatic flexibility of generic language has confounded psychologists, linguists and philosophers who have tried to articulate what exactly generic statements mean.  -->
<!-- We evaluated a theory of generic language derived from general principles of language understanding using a simple, but uncertain, basic meaning---a threshold on property prevalence. -->
<!-- Our formal model is a minimal extension of the RSA theory of language understanding, together with an underspecified threshold semantics. -->
<!-- The model was able to explain two major puzzles of generics: their extreme flexibility in truth conditions and the contrastingly strong interpretation of many novel generics. -->
<!-- Both of these phenomena were revealed to depend in systematic ways on prior knowledge about properties. -->
<!-- This prior knowledge was revealed through Bayesian model analysis to be structured, providing a promising bridge to conceptual accounts of generic language. -->
<!-- To understand the nature of the underlying prevalence scale, we showed that generic language is about speakers' expectations of future prevalence, and not necessarily what the current state of the world is like.  -->
<!-- Across all experiments, the formal model predicted the quantitative details of participants' judgments with high accuracy. -->

<!-- There have been numerous demonstrations arguing that statistics (e.g., property prevalence) are insufficient to explain generic meaning [@Gelman2002; @Gelman2007; @Cimpian2010; @Cimpian2010c; @Khemlani2012; @Prasada2013]. -->
<!-- In these experiments, the prevalence considered is only the prevalence of the property for the referent-category [e.g., the percentage of birds that lay eggs; @Khemlani2012; @Prasada2013], what we have referred to as \emph{within-kind prevalence}.  -->
<!-- Indeed, this simple statistic also fails to explain our data (Figure \ref{fig:commongenerics}b, right). -->
<!-- Our formal model of pragmatics, by contrast, considers not only within-kind prevalence, but a listener's prior beliefs about prevalence across kinds in order to arrive at a meaning for a generic utterance. -->
<!-- By establishing the validity of a semantics based on prevalence alone, we provide a formalism to learn about categories from generic statements.  -->
<!-- Further, since prevalence is a probability, our model can take information conveyed with a generic and be naturally extended to make predictions about entities in the world or support explanations of events or behavior. -->

## Statistical accounts

### Relative and absolute generics

Our underspecified threshold model has clear antecedents in other statistical accounts, most notably @Cohen1999's theory of generics as a frequency adverb (e.g., "generally").
Cohen treats generics as a class comprising two qualitatively different types: *relative* and *absolute* generics.
*Absolute generics* use a fixed, 50\% threshold on prevalence: $p>0.5$. 
That is, if a particular instance is more likely than not to have the feature, then an absolute generic is true.
By contrast, *relative generics* are true based on a comparison to an alternative set of kinds $Alt(K)$, analogous to the categories that comprise the prevalence prior (the so-called *comparison class*, which we discuss further in the General Discussion).
"Mosquitos carry malaria" is true because an arbitrary mosquito is more likely than an arbitrary member of an alternative kind to have the feature.

In our model, we treat all generics as relative. 
Attested differences in endorsement, then, emerge through the interplay of prior knowledge with our uncertain semantics.^[
  For a related theoretical argument against the "relative" / "absolute" distinction for gradable adjectives, see @Lassiter2015.
]
Further, though Cohen's theory is framed in terms of probabilities, it is a fully deterministic, fixed-threshold account which only makes qualitative predictions about what is true and what is false (i.e., it has a deterministic semantics).
In contrast, we propose a fully probabilistic semantics embedded within a Bayesian model that describes how context resolves the uncertain threshold; our theory is a joint semantic--pragmatic theory.

<!-- In our theory, there is uncertainty about the core semantic meaning, and this uncertainty (though reduced in context) drives the model to make graded predictions about endorsement. -->
<!-- Second, Cohen draws an *a priori* distinction between *Absolute* and *Relative* generics, and provides different mechanisms for each each. -->
<!-- We used examples of both in our stimuli (Expt. 1) and found that our single model handles both without any addition assumptions, casting doubt on the *a priori* distinction between "absolute" and "relative". -->

Cohen's and other statistical theories employ a mechanism (not currently required in our account) known as *domain restriction* to explain the context-sensitivity of generics: contextually restricting the entities that go into the computation of prevalence (i.e., which robins do we look at to compute the probability of laying eggs among robins?).
@Cohen1999 posits that prevalence is calculated by only considering entities that *could have some feature* in a contextually-specified alternative set of features $Alt(F)$.
For example, the property *lays eggs* induces a set of alternatives that are associated with modes of reproduction (e.g., *gives birth to live young*, *undergoes mitosis*, ...).
"Robins lay eggs" [an absolute generic for @Cohen1999] is evaluated by only considering female members of kinds, because only female members can plausibly satisfy one of the other reproductive property alternatives (i.e., the alternative features in $Alt(F)$).
The inferential machinery behind domain restriction ---how to determine $Alt(F)$---relies upon conceptual knowledge, but the details remain obscure [@Carlson1995essay].^[
  However, see @Cohen2004 for a discussion of how his semantic constraints relate to different kinds of generics and different kinds of conceptual representational frameworks used in cognitive science.
]
Our uncertain threshold model can be seen as one particular mechanism by which the domain may be restricted: The structure of the prior distribution over the prevalence of *lays eggs* is a reflection of an intuitive theory of reproduction (i.e., that only females lay eggs) and the uncertain threshold model uses that background knowledge to derive a property-specific interpretation.
There may exist a refactorization of the uncertain threshold model to a fixed threshold where the listener has uncertainty about the relevant domain of restriction.

<!-- Of course, the inferential machinery behind this sort of domain restriction (i.e., what is $Alt(F)$?) requires further theorizing to explain.  -->

### Generic as indexical

@Sterken2015 develops a novel analysis taking the context-sensitivity of generics as primary.
This analysis draws analogy to other, inherently context-sensitive linguistic expressions: *indexicals* (e.g., "this" or "I").
@Sterken2015 uses this analogy to motivate a context-sensitive *quantificational force* as well as mechanism of domain-restriction [of the kind used by @Cohen1999 and others].
Our uncertain threshold semantics can be seen be a particular formalization of @Sterken2015's context-sensitive quantificational force. 
We have not had need for employing domain restriction on the categories, though as noted above, it is a potential avenue for future development.
<!-- Finally, how the context-sensitivity of an indexical differs from the kind of context-sensitivity with an uncertain-threshold is an important question for future research. -->

### "Normal" accounts

A popular alternative view under the statistical banner draws on the intuition that generics often express something normative in the world [@Asher1995; @Pelletier1997; @Nickel2008; @Nickel2016]. 
"Dogs have four legs" is then a good generic not because all dogs have four legs (regrettably, all do not) but were the world to function *normally* (e.g., dogs would not be involved in freak tractor accidents or be born with strange genetic mutations), then *all* dogs would have four legs. 
The idea that our beliefs about what is normal in the world influences our judgments about generalizations has intuitive appeal for rejecting accidentally true generics (e.g., "Supreme Court justices have even social security numbers") and resisting stereotyped language (e.g., "Boys are good at math”).
Our theory does not directly formalize what is *normal*, though we argue that a speaker's beliefs about what is *probable* [which may relate to what is *normal*; see @Icard2017] plays a role in endorsing and interpreting generalizations. 
<!-- We showed in Expt. 2c that a speaker's beliefs about what is likely to be the case in the near future matters above and beyond what is currently true in the world for endorsing generalizations.  -->
<!-- Our notion of prevalence is as a predictive probability, which incorporates background knowledge in order to make predictions about the future.  -->

### Underquantification

A proposal similar to our account concerning underspecification of generics has been made in the computational linguistics literature [so-called *underquantification*; @Herbelot2011].
In their model, generics express an explicit quantified relation, specifically either "Some", "Most", or "All".
This proposal is used to then used to construct a set of features that accurately predicts (relative to human judgments) the quantified relationship expressed by the generic [analogous to a quantifier version of the *implied prevalence* task used in @Gelman2003b; @Cimpian2010]. 
Our semantic theory can be seen as a generalization of *underquantification* to a continuous interval of possible meanings.
This distinction is relevant for the acquisition of the language of generalizations; we do not take as primary the quantified relations (e.g., "Some", "All").
<!-- Formally, whereas the *underquantification* account posits that a generic is *ambiguous* between various meanings, we argue that the generic is *vague*, and has a continuum of possible meanings.  -->
Additionally, by using an underspecified threshold on a scale of probability, our formulation naturally extends to other scales and other kinds of generalizations (e.g., habitual language), where quantified relations like "Most" or "All" are not directly applicable.

### Cue validity

By encoding knowledge about other categories, the prevalence prior distributions in the uncertain threshold model are deeply connected to the construct of *cue validity*, or the probability of the kind given the feature: $P(x \in k \mid x \in f)$ (e.g., one's predictions about whether or not an entity is a mosquito, upon learning that it carries malaria).
Cue validity is believed to play a role in understanding generic language [@Leslie2007; @Khemlani2012; @Prasada2013], but the details remain underspecified.
The strongest view of cue validity is that it operationalizes an alternative hypothesis about the meaning of generic statements: "Mosquitos carry malaria" means "It is mosquitos that carry malaria". 
Indeed, empirically-elicited cue validity has been shown to be highly correlated with endorsements of generics [@Khemlani2012].

Cue validity is *inverse prevalence*; the two are related via Bayes' Rule:
$P( k \mid  f) = \frac{P( f \mid  k) \cdot P( k)}{\sum\limits_{k' \in K} P( f \mid k') \cdot P( k')}$.
Knowledge about other categories $k'$ enters in the denominator to compute cue validity for prevalence.
Indeed this normalizing constant (the denominator) is equal to the expected value (i.e., the mean) of the prevalence prior distribution: $\mathbb{E}[P(p)]$.
Thus cue validity comes from a point estimate of the prevalence prior distribution, and information about cue validity can be derived from the constructs posited in our model. 
We return to the implications of this relationship in the General Discussion.
For a more detailed, mathematical derivation of the relationship between cue validity and prevalence priors, see Appendix A.

## Baseline models for quantitative comparisons

In our empirical studies below, we compare our model to three alternatives.
These alternative models do not represent any of the extant theories of generics described above; no extant theory is sufficiently precise to yield quantitative predictions. 
Instead, these models are designed to interrogate the theoretically-substantive components of our model. 
There are three such components: (a) property knowledge in the form of a prior distribution over prevalence $P(p)$, (b) endorsement as a decision-theoretic process of uttering the generalization vs. not uttering it, and (c) vagueness in the semantics of a generalization (i.e., an uncertain threshold).
In our empirical studies, we compare the uncertain threshold model to an alternative, lesioned model which lacks the vagueness in meaning, assigning a fixed semantics to the generalization (i.e., analogous to a quantified statement) but which has the same prevalence prior and the decision-theoretic architecture. 
There are no correspondingly simple ways to lesion the other two components (property knowledge or speaker decision) that still produces quantitative, context-sensitive predictions.
Instead, we compare our model to two regression models based on empirically elicited referent prevalence and cue validity, which provides some interrogation of the necessity of property knowledge in the form of a full distribution on prevalence.
<!-- Furthermore, for comparison to standard methods in the psychological literature on generics [e.g., @Khemlani2012], we construct alternative regression models based on empirically-elicited quantities (referent prevalence and *cue validity*, described below). -->
In addition to serving as alternatives, these baseline regression models help us understand the statistical properties of our experimental materials and provide a comparison to standard methods in the psychological literature on generics [e.g., @Khemlani2012].

<!-- For comparison to standard methods in the literature on generics [e.g., @Khemlani2012], we construct alternative quantitative models based on empirically-elicited cue validity. -->
<!-- These control models improved the quantitative fit to the endorsement data, but did not match the explanatory power of the uncertain threshold model. -->

# Interim summary and overview of experiments

We have introduced the first quantitative theory of the language of generalization, and discussed the relationship of this account to extant theories of generics.
Above, we presented simulations showing how the model predicts endorsements for classically puzzling generic statments.
These predictions depended upon the background knowledge about the property $P(p)$ as well as the referent prevalence $p$ believed to be true for the category. 
For each of these we chose intuitive values for the parameters of the model (i.e., the prevalence priors and referent prevalence levels).
In what follows, we test this theory empirically for a wide range of generalizations, including generalizations of different types (categories, events, and causes).
We do this by both measuring and manipulating background knowledge and referent prevalence and predicting human endorsements of generalizations.

In Case Study 1, we examine generalizations about categories expressed in generic language. 
We first measure endorsements for thirty generic statements about familiar categories, revealing an entire continuum of endorsements (Expt. 1a). 
We then measure the corresponding prevalence priors and referent prevalence using a prevalence prior elicitation task (Expt. 1b).
We compare the quantitative fits of our model to the three alternative models described above.
This Case Study is an empirical version of several of the worked examples in the section above.

In Case Study 2, we examine generalizations about events expressed in habitual language while manipulating the referent prevalence.
We measure prevalence priors for events of people doing various actions (e.g., people running, hiking, climbing mountains; Expt. 2a).
We then measure endorsements of habituals for statements about novel actors (e.g., "John runs") given referent prevalence information (e.g., "In the last two months, John ran three times."; Expt. 2b).
Finally, we test whether referent prevalence in our model is best thought of as a past frequency (e.g., the number of times John has run in the past) or a prediction about the future (e.g., the number of times a speaker expects John to run in the future) by experimentally manipulating future predictions while keeping constant past frequency (Expt. 2c).
We answer this question by testing the quantitative fits of two versions of our model: one in which the speaker is conveying past frequency and one in which the speaker is conveying their predictions about the future.

Case Study 3 experimentally manipulates the prevalence prior in the domain of causal language (e.g., "Herb X makes animals sleepy").
Expt. 3a measures the (manipulated) prevalence prior, confirming that our manipulation influenced participants' beliefs about the prevalence of the feature across different categories.
Expt. 3b measures the corresponding influence on causal endorsement, finding an effect of the manipulated background knowledge in the way predicted by the uncertain threshold model.

<!-- Other theorists have focused on the diversity among generic statements and looked beyond the semantics to explain the variability between different kinds of generics. -->
<!-- We take the distinction between generics and genericity as roughly coextensive with what has also been described as a statistical vs. conceptual distinction [@Carlson1995essay].^[ -->
<!--   We use the terms statistical and conceptual to refer to what @Carlson1995essay referred to as "inductive" and "rules and regulations" views, respectively. -->
<!-- ] -->
<!-- Conceptual theories often appeal to factors that have an explanatory component (e.g., why the property is present in an instance). ^[ -->
<!--   Such factors are often thought of as outside the domain of truth conditions. -->
<!--   For example, @Nickel2016 refers to genericity as a *semantic mechanism*, and one that could plausibly interact with other semantic mechanisms (e.g., gradability in predicates as in "Giraffes are tall").  -->
<!--   This interaction, @Nickel2016 argues, is analogous to how gold---a chemical element with certain defining properties---can be studied in isolation but may also interact with other metals to produce an alloy.  -->
<!--   We would not expect a theory of gold to be able to predict with high quantitative accuracy the properties of the alloy, without a theory of the other elements involved and their respective interactive effects. Thus, it is important to identify when one has a pure case of genericity on their hands as opposed to an alloy. -->
<!-- ] -->

<!-- Instead, the statement should be glossed as *some elephants live in Africa* and *some elephants live in Asia*, and perhaps taken collectively, *most elephants live in either Africa or Asia* [@Nickel2008].  -->
<!-- This reading of (3) gives it the same semantic content as two statements: "Elephants live in Africa. Elephants live in Asia." -->
<!-- Our model is able to flexibly accomodate this statement by having uncertainty about the truth-conditional threshold.  -->

<!-- With reasonable assumptions about the prior (e.g., that *lives in* is similar to other biological properties from Expt. 1a and assuming international elephants are unlikely *a priori*), upon the hearing "Elephants live in Africa", our model will infer that most elephants live in Africa. -->
<!-- When the model then hears "Elephants live in Asia", the model updates its beliefs to accomodate this second statement,  -->
<!-- This is possible because our model has uncertainty over the semantic thresholds in the generic claim. -->

<!--
**4. Books are paperbacks.**

Sentence (4), strikes hearers as *odd* or perhaps *false*, despite the vast majority of books being paperback. 
Why?
We first observe that the sentence is constructed by pairing a category ("Books") with a property that describes a strict subset (i.e., "paperbacks"). 
We hypothesis that the starngeness of (4) has to do with the role of *predicability* [in the sense of @Keil1979, originally from @Sommers1963] in determining the comparison class.
That is, assume that entities in the relevant comparison classes must be predicable of the property: each entity in each class must either have or not have the proprerty.^[
  Note that this assumption is consistent with the comparison classes we have assumed in our three case studies.
]
Then, "is paperback" can only be informatively said of individual books, because these are the only objects that can plausibly have the property; that is, world knowledge gives us that an elephant cannot be predicated (either true or falsely) by the property "is a paperback".^[
  Note that this idea is related to @Cohen1999's notion of "Relative generics" (described above), which have truth conditions that are determined by a comparison to an alternative set of Ks which is composed of Ks that could satisfy some F in an alternative set of Fs.
]
The relevant aggregations of entities (i.e., of individual books), then, are collections of books (e.g., books of different genres or publication eras among other such sub-categories).
Then, given that "books" is the superordinate category of "kinds of books", the referent-category prevalence (the % of books that are paperbacks) is the mean of the prevalence prior (the average of the % of sub-classes of books that are paperbacks).
Given this prevalence prior and the referent-prevalence, our endorsement model would predict "Books are paperbacks" is neither true nor untrue, in a way analogous to "Robins are female". 
We expect these intuitions to generalize to other examples expressing an analogous relation (e.g., "Pastas are spaghetti", "Animals are humans").

```{r booksArePaperbacks, fig.cap="Sketch of an argument for why 'Books are paperbacks' sounds odd. The feature 'is paperback' can only be predicated (in the sense of @Keil1979) of individual books, which are organized into collections (here, by genre). The prevalence prior is constructed by aggregating these collections of books. Then the prevalence of 'is paperbook' for the category of 'books' can never be higher than the mean of the prior distribution. The endorsement model would give this sentence a rating of 0.5, corresponding to neither true nor false."}
knitr::include_graphics("figs/books-paperbacks.pdf", dpi = 108)
```

**4. Women are submissive.**

Finally, Sentence (4) suggests that a property can be widespread relative to other kinds or groups, as well as being so systemic that we would predict it to be the case in the future (high *predictive prevalence*), and speakers still might find it aversive to actually endorse the generalization [@Haslanger2011]. 
( Due of a variety of socio-cultural reasons, @Haslanger2011 argues that women tend to be submissive (more so than men, and many other social categories), yet still cautions against asserting (5). )
In our **Model Simulations** section, we showed that the generalization often implies the property is widespread, even with relatively uninformed prior beliefs about the property. 
If the generalization conveys the property is in fact more widespread than the speaker herself might believe, this can lead to a distortion of the truth [@Cimpian2010].
The current formulation of our model is relatively ambivalent about the deviations of the listener's beliefs from the speaker's belief (i.e., the speaker doesn't mind that the prevalence that she believes to be true will be exaggerated in the mind of the listener).
However, if states of the world (or, listener's beliefs about states of the world) have different *subjective values*, speakers who take into the account the subjective utilities of the listener may produce language that deviates from what a pure informational transmission speaker might say.
This kind of extension to rational models of communication has already been used to account for a variety of phenomena in *polite language* use [@Yoon2016; @Yoon2017].
Such a value-based utility structure for states of the world for language about social categories may capture intuitions like those about Sentence (4): If the speaker found states of the world where the probability of the feature is high as undesirable (e.g., a world where many women were submissive), then a speaker who considered both the informational content of the utterance and the utility in the states of the world implied by that utterance would too find it reprehensible to make such a generalization.

An alternative information-theoretic analysis of (4) could be made concerning the conceptual structure communicated by generalizations (see earlier discussion of *Prevalence Priors and Conceptual Structure*). 
Suppose there are two possible causes for a person to be submissive: Either it is in their DNA or the person was socialized in such a way as to be submissive. 
Were either of these possible causes to exist for a particular group (e.g., women), then high prevalence of the property would be expected in that group. 
A speaker may intend to communicate not the high prevalence, however, but what she believes is the generative cause of that high prevalence (i.e., either that there is an internal, DNA cause or an external, societal cause).
Suppose the speaker wants to communicate that she believes an external, societal cause is leading to high prevalence of submissiveness.
The decision to endorse (4) depends upon what the speaker believes her listener will infer. 
If she is in a group of close friends discussing the ways in which society shapes behavior, the speaker may opt to endorse (4) because with high probability, she will be understood as endorsing the external, society cause. 
However, if she is worried that some might take her to mean that there's something about women that causes them to be submissive, the speaker may refrain from uttering (4).
Such an explanation would be a purely information transmitting account that does not require any social utility to explain the behavior.

-->

<!--
In sum, we've tried to argue that our information-theoretic communicative model couched in our general cognitive framework might well explain a number of interesting and heretofore puzzling phenomena in genericity as well as make novel predictions about a wider range of language.
-->

<!-- -- Books are paperbacks. (failure of the constrast class) -->
<!-- -- Birds lay eggs and are female vs. Elephants live in Asia and Africa -->
<!-- can use pragmatics to resolve if the property is conjunctive or not.  -->
<!-- -- Women are submissive. (Haslanger, 2011; generics + politeness mechanism, some states have lower value than others) -->
<!-- -- Muslims are terrorists (vs. Mosquitos carry malaria) -->
<!-- -- Mary handles the mail from antarctica (predictive propensity; also highlights domain restriction on events) -->
<!-- - Glass breaks when struck. (so-called "disposition", but here: a generalization about the event of a glass being struck [already we are domain restricted]) -->

<!-- depends upon the probability that utterance vs. silence accurate conveys to a naive listener a. -->
<!-- This probability of successful communication is given by $L(h \mid u, \theta)$ (Eq. \ref{eq:S1}) and depends upon the listener's *a priori* beliefs about the prevalence in question: $P(h)$ (Eq. \ref{eq:L0}), as well as the referent-category prevalence $h$ (Eq. \ref{eq:S1}). -->
<!-- Predictions for endorsement depend upon how well the listener's posterior on $h$  upon hearing a generalization matches an $h$ that the speaker wants to convey in comparison to how well the listener's prior matches (which would result from the speaker staying silent). -->
<!-- For purposes of illustration, we present different hypothetical priors on feature-probabilities $P(h)$, together with listener $L$ posterior distributions on $h$ after hearing utterances corresponding to (i) the underspecified threshold semantics as well as (ii) a comparison model where the threshold is fixed near-0, corresponding intuitively to the quantifier *some* (Figure \ref{fig:simulations}). -->

<!-- This reasoning process favors lower thresholds because they are consistent with more prevalences. -->
<!-- Low thresholds (e.g., $\theta = 0.25$) rule out very few possibilities (only $p \leq 0.25$), and thus an interpretation with a low-threshold still has a lot of uncertainty about the prevalence (Figure\ \@ref(fig:simulations)A, left).  -->
<!-- Higher thresholds  (e.g., $\theta = 0.75$) rule out many possibilities; an interpretation with this threshold has correspondingly less uncertainty about the prevalence (Figure\ \@ref(fig:simulations)A, right). -->
<!-- The flip-side of this is that the endorsement model is less likely to produce an utterance with a low-threshold in mind, because the end result would be, from an information-theoretic perspective, similar to staying silent (a posterior not very different from the prior). -->
<!-- Higher thresholds (e.g., $\theta = 0.9$) rule out more possibilities and thus would be highly informative, but the interpretation model must balance this with the prior probabilities of the prevalences implied by the threshold (Figure\ \@ref(fig:simulations)A).  -->
<!-- The result of this reasoning process is a trade-off between highly informative thresholds and the prior probabilities of the prevalences implied by those thresholds. -->

<!-- The literal interpretation of a generalization about a previously unknown category (e.g., "Feps bark.") is given by Eq. \ref{eq:L0} and shown in  -->

<!-- The lighter distribution is the interpretation distribution given by Eq. \ref{eq:L0}.  -->
<!-- For instance, the interpretation of a generalization about the property \textsc{barks} is that the prevalence is quite high, near 1 with some uncertainty (top left; lighter pink distribution). -->
<!-- This would correspond to a rather strong interpretation (i.e., "Almost all feps bark"). -->
<!-- This is not the case for the \textsc{lays eggs} distribution (a la "Feps lay eggs"; bottom left, green).  -->
<!-- Given the background knowledge about that property----not that many animals lay eggs, and among those for which the feature is present, the probability is about 0.5 (because it's only present in the females)----the interpretation is much more conservative: "About half of feps lay eggs". -->
<!-- The model can even derive very weak interpretations, as is shown in the \textsc{carries malaria} distribution (bottom right, purple). -->
<!-- Again, the literal interpretation depends upon the background knowledge in a deep way. -->

<!-- The endorsement model's predictions for six generic sentences are shown in Figure\ \@ref(fig:simulations) (right). -->
<!-- Given a referent-prevalence $h'$ (e.g., $P(x \text{ lays eggs} \mid x \text{ is a robin})$; Fig.\ \@ref(fig:simulations) right x-axis) and the background knowledge of the interpreter (Fig.\ \@ref(fig:simulations) left), the endorsement model (Eq. \ref{eq:S1}) predicts an endorsement probability (Fig.\ \@ref(fig:simulations) right y-axis). -->
<!-- We see the model correctly  -->

<!-- \mht{For further discussion about the behavior of the model, and these six examples, see Appendix?} -->

<!-- that the behavior of the uncertain-threshold semantics is distinct from both the prior distribution over probability $h$, as well as the inferences derived from a fixed-threshold semantics. -->
<!-- For these simulations, we had specify both the prevalence priors $P(h)$ and the referent prevalence $h'$.. -->
<!-- In the rest of this paper, we both meausure and manipulate $P(h)$ and examine the behavior of the speaker model in three case studies: generalizations about categories, events, and causes. -->

<!-- These different semantic models make different predictions easily seen when considering a uniform prior distribution over $h$ (Figure\ \@ref(fig:simulations); top row). -->
<!-- The fixed-threshold model rules out only the lowest possible feature-probabilities ($h \sim 0$) and returns a posterior very similar to the prior. -->
<!-- The behavior of the generalization is distinct from a fixed-threshold model: The higher the prevalence $h$, the more likely it is after hearing the generalization.^[ -->
<!-- The intuition behind the model's behavior can be gleaned by imagining what the listener model $L$ would believe given different thresholds, and averaging over those possibilities. -->
<!-- If the threshold $\theta$ is very high, only the highest prevalence $h$ would pass the threshold. -->
<!-- If the threshold $\theta$ is slightly lower but still high, the highest prevalences $h$ would still pass the threshold, as would some that are slightly lower. -->
<!-- As the threshold $\theta$ takes on lower and lower values, more and more $h$'s would surpass it. -->
<!-- When $\theta$ is very low, almost all $h$ are consistent with it (akin to fixed-threshold model). -->
<!-- The listener averages over these possibilities: $h$'s with higher values pass more thresholds and the result is that higher $h$'s are more likely *a posteriori*. -->
<!-- ] -->

<!-- Next, we consider a few classic cases from the literature on generics to demonstrate the kind of inferences our listener model would draw from different generic statements. -->
<!-- The strength of a listener's inference about $h$ (i.e., how prevalent is the feature?) depends in systematic ways on the prior belief distribution over $h$. -->
<!-- In Figure\ \@ref(fig:simulations) (row 2), we show a schematic prior distribution that is bimodal with peaks near 0\% and some high prevalence values; this prior could be over the prevalence of *barking* because many kinds of creatures do not bark (i.e., have $h \sim 0$; e.g., cats), but *for those that do have members that bark*, many of them bark (i.e., have some high $h$). -->
<!-- The fixed threshold model (right column) struggles to differentiate those that generally do not have the feature from those that usually have the feature, returning a posterior distribution very similar to the prior.^[ -->
<!-- We model "those that do not generally have the feature" as a Beta distribution with an expected value (mean) close to zero.  -->
<!-- The fact that this component of the prior is not just a Delta function at 0\% reflects the intuitive possibility that some members of a kind who do not generally have the feature (e.g., a cat who barks) could somehow acquire the feature from accidental or transient causes (e.g., a strange genetic mutation). -->
<!-- ] -->
<!-- The uncertain threshold model, however, easily rules out the categories that do not generally have the feature and strongly signals the property is widespread in the category. -->

<!-- How does the model handle the puzzling case of "Robins lay eggs" vs. "Robins are female"? -->
<!-- In each case, we must consider the prevalence prior:  $P(h_\text{lay eggs})$ and $P(h_\text{are female})$. -->
<!-- For being female, no matter what animal category a person might think of, the feature is prevalent in the exact same proprotion (i.e., 50\% of X are female; Figure\ \@ref(fig:simulations) bottom row). -->
<!-- For laying eggs, it really depends: If a person thinks of a falcon, then 50\% of that category lays eggs; if a person thinks of a bear, then 0\% of them lay eggs. -->
<!-- This dependency is reflected in the corresponding $P(h)$, which is a bimodal distribution with peaks near-0\% and near-50\% (Figure\ \@ref(fig:simulations) row 3). -->
<!-- For both priors, the interpretation upon hearing the generalization is the same (i.e., roughly 50\% of the category has the feature; Figure\ \@ref(fig:simulations) rows and 5, middle red panel). -->
<!-- The difference in endorsement comes from the comparison to the prior: For being female, the listener is very likely to arrive at 50\% under her prior, but for laying egges, this is not the case.  -->
<!-- In other words, "Robins are female" has little very informational content as the listener's posterior is nearly identical with her prior (i.e., it doesn't matter if the speaker says the generalization or stays silent). -->
<!-- "Robins lay eggs", however, has a lot of informational content. -->

<!-- Finally, we look at how the uncertain threshold produces an interpretation for the statement "Mosquitos carry malaria", which intuitively should correspond to "Some mosquitos carry malaria". -->
<!-- The prior $P(h_\text{carry malaria})$ assumes that most creatures do not carry malaria and among those that do, the feature is not very widespread. -->
<!-- Here, we see that listener's posterior upon hearing the generalization (the uncertain threshold) is similar to what she learns upon hearing "Some mosquitos carry malaria" (the fixed threshold; Figure\ \@ref(fig:simulations) row 4).  -->
<!-- Thus, a speaker who believes that *very few* mosquitos carry malaria will still endorse "Mosquitos carry malaria". -->

```{r child = 'case1-generics.Rmd'}
```

```{r child = 'case2-habituals.Rmd'}
```

```{r child = 'case3-causals.Rmd'}
```

# General Discussion

The human species is remarkable not only because we can extract useful generalizations from observations of the world, but because we can convey these generalizations to each other succinctly using language.
Generalizations expressed in language (e.g., "John runs.", "Dogs are friendly.", "Fire causes smoke.") are a premier example of how simple statements---statements understood by even the youngest language users---can convey rich meanings and display complex sensitivities to context.
We have argued that the core meaning of such linguistic expressions can be understood by three ingredients: Probability, vagueness, and context. 
That generics are vague does not preclude them from being treated by formal models. 
A vague threshold meaning operating over a speaker's inductive beliefs, formalized by predictive probability, leads to complex, quantitative interactions with background knowledge.

We tested this theory by exploring its implications for the simplest judgments a person can make about a sentence: an endorsement or truth judgment.
Truth judgments are the standard measurement for semantic theories, but are often limited to the intuitions of a single or a small handful of trained theorists, often producing only binary or ternary judgments (i.e., true, false, and possibly indeterminate).
In this work, we measured the truth judgments of naive language users, revealing whole continua of gradience. 
These fine-grained, quantitative measurements pose real challenges for verbal theories of generics, which can only predict qualitative differences.
The framework we propose is sufficiently precise to predict quantitative gradience, and we showed that the gradience in truth judgments depends in systematic ways on an interpreter's prior beliefs about prevalence.
We have provided strict tests of our formal model by both measuring and manipulating listeners' background knowledge as well as comparing our model to a number of alternatives.
In each case, our model provided a strong quantitative fit to human judgements while other models fall short.
In addition, the model accounts for standard linguistic intuitions about a number of philosophically puzzling generic statements (Worked Examples).

The fact that our model applies equally as well to generic and habitual language sheds light on a long-attested relationship.
The analogy of habituality (e.g., "John runs") and genericity (e.g., "Dogs have four legs") has typically been assumed in the literature following the original suggestion by @Carlson1977.
The relationship, however, has been never been empirically tested nor formally described.
This paper presents the first empirical evidence that the reasoning involved in understanding generic and habitual sentences are similar: The same uncertain threshold mechanism explains judgments of both kinds of statements (Expts. 1 & 2).

In our model, a generic statement or other generalization updates a listener's *a priori* beliefs about the prevalence of the property for that category. 
Since the prevalence prior includes the information necessary to compute cue validity (see Appendix A), the cue validity of the feature for the new category is added to the common ground between interlocutors.
Thus, given the knowledge represented in a prevalence prior, generics communicate information both about the prevalence of the feature in the new category (via the literal semantics) as well as the cue validity of the feature for the new category.
For example, if a listener hears "Alligators grow to be 10-feet long", they will update their personal beliefs both about the prevalence of 10-footedness among alligators as well as the probability that an animal is an alligator given that it is 10 feet long. 
To our knowledge, this is the first theory of generics that describes how the informational contribution of a generic relates to cue validity.^[
  We are grateful to an anonymous reviewer for drawing our attention to this. 
]
This also means that our computational framework makes predictions about the cue validity implied by different generalizations, without any further assumptions.

We propose that the language of generalization conveys predictive probabilities.
Communicating probabilities might seem contrary to the long attested failures in reasoning about probability [@Tversky1974].
Our theory suggests that the problems with understanding probabilities observed in classic, cognitive psychology paradigms are a problem in understanding explicit probabilities expressed using *numerical language*, a historically quite recent innovation [cf., @Levinson1995]. 
Rather than conveying probabilities explicitly, the language of generalization conveys them implicitly.
In other words, we argue that the utterance "70% of birds fly" is a precise statement about how many birds fly in the way that "John is 6'3"" is a precise statement about the height of John, whereas "Birds fly" is a vague statement more akin to the utterance "John is tall".
The latter statements are easier to process, understood at an earlier age, and may be more useful for human reasoning. 
Indeed, even infants are actually quite good with reasoning about probabilities, but in ways that are not explicit or tied to the numerical language of probability [@Xu2008; @xu2009statistical].

In the rest of this discussion, we further discuss the contribution of our modeling framework to empirical and theoretical work on the language of generalization, elaborate further on the role of conceptual knowledge in our modeling framework, describe how our model relates to language understanding more generally including theories of vagueness, and sketch an argument for how our formal representation of the language of generalization could provide a relatively simple acquisition problem for a language learner.

<!-- We measured and modeled truth judgments in order to  -->
<!-- We modeled truth judgments as a speaker's decision about whether it would be useful to utter the generalization.  -->
<!-- Endorsement, we have shown,  -->

<!--
NDG: i cut this from below...
This paper puts forth the theory that the language of generalization communicates predictive probabilities from speaker to hearer.
Prevalence in our model is not an observable frequency---the object of other statistical approaches to generics---but a mental representation.
These probabilities encode predictions about the future and are subject to top-down moderation (e.g., by way of intuitive theories).
Predictions about the presence of a feature can be the result of beliefs about internal causes of the feature [e.g., in an essentialist view of natural kinds; @Gelman2003], generative kind representations [@Prasada2009], or merely correlational happenstance.
What matters for their influence on generic endorsement, however, is their influence on subjective, predictive probability.
-->

## A unified model

The statistical vs. conceptual distinction for theories of generics, described earlier in this paper, is roughly co-extensive with the distinction between theories of the truth conditions of generics and the mental phenomenon of *genericity*, respectively [@Nickel2016].
Truth conditions are unified criteria that all generics must satisfy in order to be true.
*Genericity*, on the other hand, is thought to be a non-linguistic, psychological phenomenon previously explained either in terms of pragmatics [@Declerck1991] or meta-semantics [@Leslie2007; @Liebesman2011; @Sterken2015; @Nickel2016].
A full account of generic language must include both a theory of generics (i.e., truth conditions) and a complementary theory of genericity [@Nickel2016].

Our model leverages the insights of formal semantics and computational cognitive modeling to articulate how an agent should update their beliefs based on a generic sentence. 
The Bayesian model provides a clean separation of the semantics of a generalization (represented formally by a threshold function) from world knowledge (represented by the prevalence prior).
We, thus, chart-out a single answer to the semantic question "what do generics mean?", while also formalizing how background knowledge influences generic understanding (a question about genericity).
Previous theoretical accounts have either aimed to account for the context-sensitivity of generics by positing distinct semantic constructs [e.g., *relative* vs *absolute* generics; @Cohen1999] or positing a semantics that cannot be separated from world knowledge [@Leslie2008].
The modeling approach we take here demonstrates how the context-sensitivity of generics can emerge via the interplay of a stable (but vague) semantics and diverse background knowledge.
<!-- That is, our framework provides both a theory of generics and genericity, and their interaction. -->
This model is thus the first step towards unifying conceptual approaches to generics with an inherently quantitative natural language semantics.

<!-- That is, the truth conditions of generics are a function of beliefs formalized with probability, and the relevant probabilities may be informed by rich conceptual structure. -->
<!-- The separation between conceptual structure (which exists in the mind) and prevalence (a quantitative metric for natural language semantics) provides a unified view for the phenomenology of generics. -->

<!--
\mht{unsure about this paragraph... a version of it may be better for the cover letter (if at all)}
\ndg{yeah, this is unecessarily inflamatory. ax it.}
Even the most influential of extant theories generally makes one or a few predictions about "different kinds" of generics, and leave it to intuition to decide for a particular generic statement, what kind of generic it is.
In this paper alone, our theory has made hundreds of predictions about the endorsement of individual generic statements, with high accuracy, derived from a general theory that we posit applies to all generic statements.
In addition, previous accounts are rarely general enough to include generalizations of different semantic types (e.g., *habituals*) and deal only with generics.
For these reasons alone, our account is novel. 
-->

<!-- Thus, in addition to expressing a precise hypothesis about the truth conditions of generics, instantiating it in a cognitive model opens up the possibility of building more elaborate models that learn abstract, generalizable knowledge from language. -->

<!-- We note that regressions model that take into account the referent prevalence and the distinctiveness of the feature (e.g., cue validity) do reasonable jobs at explaining some of the variance in the endorsement data. -->
<!-- These kinds of model, however, do not represent bonafide theories of the semantics of generalizations; they are merely a re-description of the data. -->
<!-- <!-- \ndg{this point is unclear. maybe save it for general discussion?} -->
<!-- A theory of the semantics of a generalization must formalize the conditions under which the generalization is true or false as well as how those truth-conditions relate to endorsement.  -->
<!-- That is precisely what our information-theoretic model does. -->
<!-- xx -->
<!-- The formal Bayesian model we articulate provides a clean separation of the semantics of a generalization from world knowledge. -->

<!-- That is, we provide a route for precise quantitative influence of world knowledge on generic language by way of shared background knowledge between speakers and listeners, which at its base takes the form of a prevalence prior. -->
<!-- Our communicative framework also includes a model of a listener who interprets generalization. -->
<!-- Such a quantitative model makes further predictions about domain-general learning from the language of generalization, which we leave for future work.  -->

<!-- \mht{let me know where, if anywhere, we should mention "double generalizations" (e.g., generics about habitual properties)} -->

<!-- \ndg{ -->
<!-- [these paragraphs were out of place here. perhaps incorporate later...] -->
<!-- } -->
<!-- Our model formalizes a simple meaning for generalizations and investigates the structure of world knowledge that could give rise to the context-sensitivities observed in the data. -->

<!-- Our separation of world knowledge from the semantics of generics provides a substantial, but well-circumscribed role for conceptual structure to influence generic production and comprehension. -->
<!-- We do this without having to posit that conceptual structure and semantics are one in the same. -->
<!-- This is makes it possible to connect with theories of formal semantics as well as applied Natural Language technologies that seeks to draw inferences from information conveyed in generalizations in text [e.g., @Herbelot2011]. -->
<!-- Statistical accounts in particular [@Cohen1999; @Nickel2008] -->

<!-- @Nickel2016 notes that genericity is particularly challenging to study because it intersects with many other phenomena in natural language (e.g., gradable adjectives in "Giraffes are tall"). -->
<!-- We agree, and add to this that genericity intersects with *world knowledge* in subtle ways, and it is critical to accurately describe and measure world knowledge before making claims about the semantics of a generalization. -->
<!-- We used the most basic tests to validate our model: Production was examined in terms of *endorsement*, which can be seen as a special case of production when the only available alterantive utterances are the assertion or its negation. -->
<!-- Comprehension was examined in terms of *implied prevalence*, which in our theory is most basic inference a hearer can make upon hearing a generalization. -->
<!-- Though these are highly abstracted tests of the model, the modeling framework in principle allows for elaboration to a more complete production and comprehension model.  -->
<!-- The model is a communicative model, which includes a component corresponding to *comprehension* of genericity (i.e., a listener model).  -->

<!-- We conclude by sketching out arguments for other known philosophical puzzles in genericity. -->

<!-- Implicit aspects of language use: The comparison class and question under discussion -->

<!-- to explore when and how people choose to communicate their abstract knowledge when they have more options than in a two-alternative forced-choice task. -->
<!-- The view of our endorsement model as a minimal communicative model highlights aspects of our modeling approach for further future development.  -->

<!-- ## Generic Identification -->

<!-- Throughout this paper we treated the bare plural construction as a generic utterance with a threshold semantics: $\denote{\text{K F}}(x, \theta)=x>\theta$. -->
<!-- The bare plural construction can also indicate a specific plural predication. -->
<!-- For example, "Dogs are on my lawn" picks out a specific group of dogs, while "Dogs have fur"  does not [@Carlson1977]. -->
<!-- The problem of \emph{identifying} a generic meaning from a bare plural construction is itself a challenging problem because generic meaning can be signaled using a diverse array of morphosyntactic cues. -->

<!-- @Declerck1991 suggests that generic and non-generic bare plurals can be treated in the same way, and that pragmatic considerations alone may resolve interpretative differences. -->
<!-- Indeed it does appear that knowledge of the properties under discussion (e.g., the state of being on a front lawn; the state of having fur) could facilitate the generic identification process. -->
<!-- Other pragmatic factors, like knowledge of the identity of the speaker (e.g., a teacher vs. a veterinarian), can also disambiguate generic and non-generic meaning [@Cimpian2008]. -->
<!-- Recent work suggests that utterances that fail to refer to specific entities or events could pragmatically imply generic meaning [@Crone2016cogsci]. -->
<!-- Incorporating these insights about generic identification into an information-theoretic, communicative perspective is a natural extension of this work. -->

<!-- Expts. 2c \& 2d that showed using participants' *predictions* about the future as the object of communication in our formal model perfectly tracked habitual endorsement, when the present frequency would make the wrong prediction. -->

<!-- One might also question the feature of our computational model that the speaker is explicitly trying to communicate $h$, her subjective probability.  -->
<!-- We note that if the prevalence prior is structured, as we have argued throughout, then the Question Under Discussion that our speaker model is addressing can also be formalized to be "What component of the prior does the referent-category belong?". This model with a modified QUD makes almost identical quantitative predictions as the simple model without this abstract QUD.  -->

<!-- ### The Problems with Communicating Generalizations -->

<!-- Thus, in order to define a generalization about a category, there must be some corresponding concrete particular instance of that category from which the generalization is formed.  -->
<!-- For example, if an agent observes $n$ instances of \textsc{dog} ($d_1, d_2, ..., d_n$) -->
<!-- For example, if an observer forms a generalization about \textsc{dogs}, she must some way of determining when she is in an instance of the category \textsc{dogs}: She must be able to *individuate*. -->
<!-- In Hume's words, generalizations are predictions about "instances of which we have had no experience resemble those of which we have had experience" [@HumeTHN]. -->

<!-- Generalizations can be made about almost anything. -->
<!-- Here, we restrict our focus to generalizations about categories (whose linguistic expression is known as *generic language*), which have been the primary focus of psychologists, linguists, and philosophers. -->
<!-- We posit that our analysis should expect to generalizations about events (habitual language) as well as generalizations about causal forces (causal language). -->

<!-- Generics express a relation between a kind K (e.g., \textsc{robins}) and a property F (e.g., \textsc{lays eggs}), such that the property can also be said to be applicable of an individual (i.e., the bird in my backyard lays eggs). -->
<!-- Bare plural statements (e.g., \emph{Robins lay eggs}) tend strongly to yield a generic meaning [@Carlson1977], though other forms can express such a meaning sometimes (e.g., \emph{A mongoose eats snakes.}). -->

<!-- Given that generics express a property that can be applied to individuals, it would seem intuitive that the number of individuals with the property would be what makes the statement true or false. -->
<!-- Counter-examples like \emph{Mosquitos carry malaria} and \emph{Birds lay eggs} v. \emph{Birds are female} stifle such intuition.  -->

<!-- ### Statistical Accounts of Generics -->

<!-- Statistical accounts take the \textbf{property prevalence} to be fundamental: \emph{Birds lay eggs} means \emph{Birds, in general, lay eggs}.  -->
<!-- Of course, birds do not in general lay eggs (it's only the adult, female ones that do). -->
<!-- The primary way of dealing with such issues is to posit domain restrictions ("implicitly, we are only talking about the females") when there are "salient partitions" [@Carlson1995]. -->
<!-- The most fully-developed theory on this front is due to @Cohen1999.  -->
<!-- Let's first introduce some notation: -->

<!-- ### Conceptual Accounts of Generics -->

## The role of conceptual knowledge

Our theory assumes that listeners and speakers employ real-world knowledge of objects and events in order to use and interpret generalizations.
It is this real-world knowledge (presently formalized in terms of a prevalence prior) where issues about domain specific knowledge, interpretation of properties [@Prasada2006], over-hypotheses [@Leslie2008], and essentialism [@Gelman2003] can all play a role.
<!-- It was not our present goal to model this aspect of interlocutors' conceptual knowledge. -->
<!-- Instead,  -->
We hypothesize that all of this knowledge feeds into participants' predictions, measured through judgments about prevalence.
<!-- which we model in a precise, quantitative manner. -->
This framework, thus, brings precision and clarity to the *output* of these cognitive processes and conceptual understanding: intuitions about prevalence.
Applying the same precision and clarity to the more abstract aspects of conceptual knowledge including property knowledge, essentialism, etc. is an important next step for this line of work. 
Probabilistic causal models [@pearl1988probabilistic; @Gopnik2003theory] and their generalization in probabilistic programs [@Goodmanconcepts] are obvious starting places to look for such a formalism for conceptual structure and higher-order abstractions.

 <!-- You bring precision and clarity to the judgments by capturing the output of the cognitive processes that generate prevalence.    But you tend to downplay or at least not emphasize that importance and sometime sounds as if you think you have an alternative view that argues against some of the other perspectives.    I think this is worth clarifying. -->

<!-- (at least in the form of a *prevalence prior*) against which generalizations are interpreted. -->
<!-- In our modeling work, we found that prevalence prior distributions were well-modeled by a mixture-model with two mixture components. -->
<!-- Such a prior distribution implies categorical distinctions between different kinds of categories, perhaps by virtue of having (or lacking) certain causal mechanisms that can bring about the target feature. -->
<!-- Actual conceptual knowledge, needed to interpret generalizations, is most likely richer than our simple two-component prevalence distributions. -->

<!-- Conceptual structure, including higher-order abstractions, can be formalized in probabilistic causal models [@pearl1988probabilistic; @Gopnik2003theory] and their generalization in probabilistic programs [@Goodmanconcepts]. -->
<!-- We believe such probabilistic concept representations can be adopted as a formal analog to intuitive theories of natural kinds and their properties [@Gelman2003; @Keil1992]. -->
<!-- We do not assume a particular view of natural kinds in our model, but note that viewing natural kinds as richly structured categories with clusters of non-obvious properties is compatible with our information-theoretic communicative model.  -->
<!-- The important commonality among different domain-specific background theories is that they each have a statistical aspect (supporting prediction via beliefs about prevalence), which language can interface with. -->
<!-- Thus, our theory provides a separation between conceptual knowledge and the semantics of generalizations; the interface between the two is predictive probability or *prevalence*.  -->

The fact that conceptual knowledge is not a part of our formal semantics does not imply that such knowledge is unrelated to communicating generalizations.
In fact, we saw in Expt. 2c that it was participants predictions about the future that matter for the prevalence computation involved in endorsement.
Intuitive theories guide these predictions, and might provide new insight into classic puzzles of generics.
We described in the Worked Examples how our model treats "Supreme Court Justices have even social security numbers" as infelicitous even when 100% of justices have even SSNs, because a speaker's subjective probability that the *next justice* would have an even SSN is likely 50%.
However, were we to learn much more surprising information---for instance, if every Supreme Court Justice in U.S. history had a social security number which was a *prime number* (a much lower probability outcome)---the shear suspiciousness could compell an observer to revise their theory of the domain (appealing perhaps to a conspiracy), update their subjective probability of future instances, and endorse such a generic.

Not only does conceptual knowledge influence interpretation via beliefs about prevalence, but conceptual knowledge may very well be what a speaker intends to communicate.
Like other models in the probabilistic pragmatics tradition, our model distinguishes the relevant variables for the truth-functional semantics (e.g., a threshold and a prevalence) from those that impact the speaker's utility in producing the utterance. 
A speaker's utility comes from addressing an implicit goal of communication, sometimes referred to as the *Question Under Discussion* or QUD [@roberts1996qud].
The separation of the QUD from the literal, truth-functional denotation of the utterance is an important theoretical distinction made explicit in our formal modeling approach [@Goodman2016]. 
For simplicity, the model presented here assumes the QUD is about prevalence.
However, the model makes nearly identical predictions if the QUD concerns abstract parameters of the prevalence distributions [e.g., if the speaker intends to convey that "K is the kind of thing that Fs"; cf., @Prasada2006], something much closer to communicating conceptual understanding.
This observation may prove useful for constructing models of speakers whose goals are to convey abstract relations between kinds and properties.

Conceptual knowledge concerning the strikingness or dangerousness of properties and their relation to generic language has been of particular interest to psychologists and philosophers since @Leslie2007.
We described before that the strikingness of certain properties plausibly influences prevalence priors or a speaker's predictions about future prevalence.
We have pilot data suggested these constructs are indeed altered as a result of introducing information about the dangerousness of features to participants [e.g., as in @Cimpian2010].
More empirical work is needed to understand the conditions under which strikingness influences generic understanding above and beyond the probabilistic constructs we have posited in our theory.
<!-- If such effects exist, our modeling framework could be elaborated with a non-informational speaker utility [see e.g., @Yoon2016; @Yoon2017]. -->

<!-- The model, however, has the flexibility to address richer communicative goals [also known as the *Question Under Discussion* or QUD; @roberts1996qud]. -->

<!-- Future work should investigate formalizations of attested differences in background knowledge about properties [e.g., @Prasada2013's *principled*, *statistical*, and *causal* connections] and their corresponding influence on generic endorsement via prevalence priors.  -->

<!-- \ndg{ [why is this par here. and what does it mean??] -->
<!-- The rational model of communication that we have developed in this paper makes explicit a speaker's assumptions about the goals of communication, sometimes referred to as the *Question Under Discussion* or QUD [@roberts1996qud].  -->
<!-- Richer communicative contexts will surely need to elaborate the QUD to fit more appropriately with the informational goals of the context in which generalizations are produced. -->
<!-- This observation is promising for the future goal of integrating communicative goals about kinds or conceptual structure into the information-theoretic communicative framework presented here. -->
<!-- } -->

<!-- Primarily investigated through generic language (category generalizations), this perspective holds that generics are the linguistic outlet of the mind's default mode of representing kinds [@Leslie2007] and "...an adequate account of the meanings of generic sentences requires an understanding of the ways in which we represent connections between kinds and properties [p.406; @Prasada2013]." -->
<!-- @Prasada2006 and later @Prasada2013 distinguish properties that have *principled connections* to kinds, *statistical connections*, and *causal connections*.  -->

<!-- Leslie suggests information that is striking (e.g., "Tigers eat people") is ecologically useful and thus permitted to be a generalization, even if the property is not widespread. -->
<!-- Prasada has argued that *characteristic* properties (e.g., "Diapers are absorbent") and *statistical* properties (e.g., "Diapers are white") are  -->
<!-- Gelman outlines how generics tend to express *essential* qualities that are relatively timeless and enduring.  -->
<!-- Where in the probability-based semantics could such conceptual distinctions come into play? -->

<!-- Inferential role semantics -->
<!-- We take the view the causal relations only does anything because it implies statistical regularities, and these statistical regularities are what our linguistic model depends on. -->

<!-- - what are habituals like under Leslie? -->

<!-- probabilistic treatment of kinds vs. of causality -->
<!-- conceptual representation vs. causal structure -->

<!-- rational rules: concepts are relatively definitional (with some probabilistic structure on those definitions) -->
<!-- - leslie: my kinds are your probabilistic concept representations -->
<!-- - the reason why particular generics are licensed. not that there are kinds and that generics communicate about them. -->
  <!-- - fundamentally communicative element to what the prevalence is and why a generic would sound good in a praticular case (information relative to the prevalence prior) -->

<!-- - Ellen: causality implies a mechanism, not just a statistical relation. -->
<!-- - are there parts of the  -->
<!-- - essentialism: different generative models that imply the same statistics with different causal relations (but orthogonal) -->

## The comparison class

<!-- ^[ -->
<!--   This behavior might suggest a potential area for model revision.  -->
<!--   Intuitively, "Xs are female" should mean "All Xs are female". -->
<!--   In this paper, our empirical interest concerns endorsements or truth judgments; at this point, the interpretation model is only an intermediate representation useful for deriving predictions about endorsement. -->
<!-- ] -->

The probabilistic communicative model we introduce in this paper assumes shared background about the statistics of an event or property in question, represented by the prior belief distribution over the prevalence $P(p)$.
We constructed prevalence priors for a property, event, or cause by considering other possible categories having the property, people doing the action, or causes producing the effect.
In Expt. 3, we empirically demonstrated the influence of other categories on endorsement of generalizations.
Collectively, these other kinds, people, or causes form *comparison classes* against which the referent-category is evaluated.
Thus, the prevalence prior $P(p)$ is actually a conditional distribution constructed with respect to some comparison class $C$: $P(p \mid C)$.

In this paper, we have assumed particular values of $C$ for our three case studies (*animals*, *people*, and *possible causes*, respectively).
We think the choice of these classes is intuitive, but it is a limitation of this work that we do not derive these choices from more general information-theoretic considerations.
The problem of choosing a comparison classes, however, is not unique to the language of generalization, but is a problem for any theory of *vague* or *underspecified* language [e.g., gradable adjectives like *tall* and vague quantifiers like *many*; @Bale2011; @Solt2009; @Qing2014; @Schmidt2009; @Solt2012]. 
We have begun to explore how pragmatic reasoning and world knowledge can flexibly adjust the comparison class to appropriately suit the context [e.g., how a "short" basketball player is *short for a basketball player* while a "tall" basketball player is *tall for a person*; @Tessler2017].
It remains to be seen how such principles could operate on richly structured, hierarchical knowledge about categories and properties that would be important for interpreting generic language.

The employment of comparison classes in our model provides additional flexibility towards modeling different communicative goals.
The comparison classes used in our case studies were constructed with respect to the category (*other animals*, *other people*, *other possible causes*).
Thus, our endorsement model assumes that the QUD was "*what* has this feature?".
Any part of a sentence, however, can be brought into focus (e.g., by prosody) and turned into an answer to a QUD: "*Dogs* have four legs" (as opposed to other animals), "Dogs *have* four legs" (as opposed to eating or doing other things with four legs), "Dogs have *four* legs" (as opposed to two legs, three legs, six legs), "Dogs have four *legs*" (as opposed to other limbs/body parts), "Dogs *have four legs*" (as opposed to having other features).
It has long been noted that the same generalization can be used to address multiple QUDs [@KrifkaGenericBookFocus], and we posit that differences in interpretation are the result of multiple distinct comparison classes competing for influence.
<!-- "*Lawyers* care about the law." suggests a category-wise comparison class perhaps in a pedagogical context (i.e., "Lawyers *(as opposed to doctors, firefighters, ...)* care about the law.") while "Lawyers care about *the law*" suggests a feature-wise comparison class perhaps produced by a sardonic speaker (i.e., "Lawyers care about the law *(as opposed to justice, doing what is morally right, ...)*"). -->
Prosody provides a good cue towards resolving the comparison class, and pragmatic reasoning is likely relevant here as well [@Declerck1986; @Tessler2017].
In this work, we focused on category-wise comparison classes purely for methodological convenience.
Future work should investigate the factors that give rise to feature-wise interpretations of generalizations (e.g., "Dogs *have four legs*", as opposed to having other features), other QUDs generics can address, how the relevant comparison classes are constructed, and how these inferences interact with threshold inference.

<!-- Adult judgments of the felicity of gradable adjectives like *tall* or *dark* depend upon fine-grained details of the statistics of the comparison class [@Qing2014; @Schmidt2009; @Solt2012] in ways directly analogous to the context-sensitivity highlighted here for the language of generalizations. -->
<!-- An important problem for comparison classes for both gradable adjectives and generalizations is that the comparison class often goes unsaid (e.g., "John is tall for a person" is often produced as just "John is tall"; "Robins lay eggs relative to other animals" is just "Robins lay eggs"). -->
<!-- In this paper, we do not include such a mechanism in our model; instead made the comparison class clear to participants in order to study the remaning context-sensitivity of generics. -->

<!--
[NDG: i cut this because it is unecassary speculation.]
Uncertainty about the comparison class adds yet more opportunity to the language of generalization for disagreement, miscommunication, or deceit.
Even if interlocutors roughly coordinate on a particular comparison class (e.g., agreeing that they are talking *relative to other animals*), they may fail to coordinate on the prevalences for other categories in the comparison class.
That is, our model makes the prediction that the acceptance or rejection of controversial statements (e.g., "Muslims are terrorists") depends upon a speaker's beliefs about how often other categories exhibit the property (e.g., neo-nazis committing acts of terrorism).
Understanding the contextual influence on the construction of comparison classes should provide insight into failures of communication.
\mht{ellen: TRUE,  I AGREE WITH YOU AND THIS IS PROBABLY OKAY.  BUT IT THINK IT WOULD BE BETTER TO SAY SOMETHING SUBSTANTIVE;  FOCUSSING ON HOW KNOWLEDGE OF THE COMPARISION CLASS SHOULD PROVIDE INSIGHTS INTO FAILURES TO COMMUNICATE SUCCESSFULLY.}
-->

<!-- Thus, we conclude generic, habitual, and causal language can be seen as a special case of *vague language*.  -->
<!-- The construction of a comparison class adds another layer of flexibility into the use of generalizations in language. -->
<!-- Objects can be conceptualized and categorized in multiple ways, giving rise to multiple, logically-possible comparison classes (e.g., a robin is a bird, is an animal, is an object).  -->
<!-- We have begun doing work investigating how pragmatic reasoning can help constrain this open-ended problem [@Tessler2017]. -->

<!-- ### Question under discussion -->

<!-- However, the choice in constructing a comparison class can be thought of as the result of addressing a particular QUD. -->
<!-- The QUD is thus another layer of flexibility that makes understanding generalizations so challenging. -->
<!-- In the minimal contexts employed in our experiments, this is a reasonable assumption. -->
<!-- There is a different assumption we could have made, however: Another clear way to construct the comparison class is with respect to other features.  -->
<!-- This *feature-wise* comparison class would correspond to addressing the implicit QUDs: *what does \textsc{person} do?*, *what effects does \textsc{cause} bring about?*, and *what features does \textsc{kind} have?* -->

<!-- Throughout our experiments in this work, we have focused on sentences about animals.  -->
<!-- In addition to being the main focus of past theoretical and empirical work, focusing on animals is methodologically convenient as the comparison class for generics about animals is quite naturally \emph{animals}. -->
<!-- When we look beyond generics about animals, deciding what goes into a comparison class becomes less clear. -->
<!-- There are some hints that the comparison class can be derived with respect to the property [@Keil1979], but may involve pragmatic reasoning as well. -->
<!-- For example, the statement "iPhones are useful" could be in comparison to other forms of technology (like a desktop computer), while "iPhones are heavy" could really only be informative relative to other handheld devices. -->

<!-- The incorporation of a comparison class into the study of generic language might help elucidate other puzzles concerning generics. -->
<!-- Recent work in philosophy and linguistics, for instance, suggest generic language is context-sensitive [@Nickel2008; Sterken2015]. -->
<!-- @Nickel2008 points out that \emph{Dobermans have floppy ears} may be true in the context of a discussion of evolutionary biology but that \emph{Dobermans have pointy ears} is true in a discussion of dog breeding. -->
<!-- Our theory provides a hint from where to begin to understand this context sensitivity: the comparison class. -->
<!-- Different conversational contexts could bring to mind different comparison classes, in a way analogous to the context-sensitivity of  gradable adjectives (e.g., \emph{tall}).  -->
<!-- Hearing that "Abigail is tall" means different things is Abigail is 20 years old or if she is 4.  -->
<!-- Future work will be needed to explore whether a pragmatic inference approach is also relevant to establishing the comparison class, and what background knowledge about properties, categories, and context is relevant. -->

## Genericity and vagueness

We have argued that the language of generalizations are vague descriptions of prevalence or frequency in a way analagous to how gradable adjectives like "tall" are vague descriptions of some underlying degree scale (e.g., height). 
However, vague predicates like "tall" exhibit a number of additional phenomena that we do not suppose generics or generalizations would necessarily exhibit.
For example, vague predicates admit borderline cases (e.g., a person who is neither tall nor not tall). 
It is *a priori* unclear whether one could construct borderline cases with generics or habituals without sounding like a contradiction (e.g., "John both runs and doesn't run").
Additionally, Sorites paradoxes can be constructed out of arguments using vague predicates, and it is not clear that generics or habituals could be used to construct such arguments.
No experimental data yet bears on these ideas, but would be useful for further understanding the relationship between generalizations and vague predicates.

For now, we understand both genericity and vagueness as analogous linguistic phenomena insofar as they can both be thought of as the result of a listener being uncertainty as to the exact truth conditions (e.g., an uncertain threshold).
We note that this sort of "lexical uncertainty" is a kind of *parameter learning* problem (e.g., the listener knows the form of the truth conditions, but not the values of some variables) which can be formally distinguished from other sorts of uncertain truth conditions better modeled as a *structure learning* problem (e.g., lexical ambiguity).
Further work should be done to better understand the kinds of context-sensitive linguistic phenomena that are modeled as parameter vs. structure learning problems.

Our view of understanding generalizations involving reasoning about an uncertain threshold is reminiscent of @Williamson1992's epistemicist view on vagueness.
This view holds that when a speaker utters a vague statement, there exists an objective set of criteria determining whether the entity in question satisfies the vague predicate.
Vagueness emerges from uncertainty about the details of those objective criteria (e.g., the precise value of a threshold). 
This uncertainty in turn could be the result of a population of speakers using slightly different criteria, thus shifting the criteria or threshold.
Listeners then will not know exactly what kind of speaker they are dealing with *a priori* and normatively should maintain uncertainty about the truth conditions. 
In our model, world knowledge and general communicative principles help a listener resolve the value of threshold [see @Lassiter2015 for an extended discussion of resolving the value of a free semantic variable in context].

## Acquiring the language of generalization

<!-- The linguistic outlet for generalizations about categories---so called, *generic language*---has received tremendous attention from psychologists, linguists, and philosophers. -->
<!-- Generic language is one of the earliest emerging forms of complex, compositional language. -->
<!-- Somewhere between two and three years of age, children recognize that generics convey a generalization about a category, not directly tied to concrete instances in a scene [@Cimpian2008]. -->

Perhaps the most surprising aspect of the language of generalization is how difficult it is to formalize, given how common it is in child-directed and child-produced speech [@Gelman1998; @GelmanEtAl2004].
<!-- from a formal perspective is that generic language is far from trivial to characterize.  -->
<!-- Generalizations are common in child-directed and child-produced speech [@Gelman1998; @GelmanEtAl2004] and are believed to be central to the growth of conceptual knowledge [@Gelman2004], -->
Generics are often contrasted with quantifier language (e.g., "some", "most", "all"), whose truth conditions are easy to formalize but which pose difficulty for young children to acquire [@Brandone2014; @Gelman2015].
<!-- If generics convey something about the probabilty of a feature, we would expect them to be in some way comparable to quantifier statments (e.g., "some", "most", "all", ...). -->
<!-- The extreme flexibility of generic meaning (e.g., "Birds lay eggs" vs. "Birds are female") stands in stark contrast to its early emergence in development. -->
<!-- In fact, quantified statements, whose formal meaning is much more straight-forward (e.g., "Some" means more than zero), emerge *later* in development [@Brandone2014; @Gelman2015], and are often confused for generic statements [@Leslie2012a]. -->
@Leslie2008 argues that the necessary complexity of a formal account of generics and the simplicity with which young children acquire generics implies that the normal tools for describing the semantics of quantified utterances (e.g., a truth-functional threshold) are inappropriate for generics.^[
  This argument is simplistic in that it ignores evidence that children's struggle with quantifiers stems from pragmatic issues [@Musolino2006].
]

Rejecting the tools of truth-functional semantics for generalizations would be throwing the baby out with the bathwater: In fact, our model of generics can be viewed as *simpler* from a learning perpsective than a model of quantifier semantics.
With some reasonable assumptions about the hypothesis space of semantic meanings, the acquisition of a threshold-based truth-functional meaning $\denote{u}(p, \theta) := p > \theta$ requires learning three distinct aspects of meaning: 
(1) the *dimension* being described (i.e., prevalence $p$),
(2) the *polarity* of the relation (i.e., $>$ vs. $<$), and 
(3) the value of the *threshold* (e.g., $\theta = 0$ for "some", $\theta = 0.5$ for "most").^[
  This particular framing of the acquisition process ignores the potentially different impact of pragmatic reasoning on generics vs. quantifiers. 
]
If a learner first acquires the dimension and the polarity (i.e., generics have to do with prevalence in some positive way), a rational learner should then represent uncertainty over possible thresholds $\theta$.
For quantifier semantics, a learner would then need to learn the context-invariant value of the threshold $\theta$, but for generics, she could have an adult-like semantics for generics by maintaining uncertainty about the threshold.
That is, the language of generalization is learned once aspects (1) and (2) are understood.
Thus, under our framework, generics should not only be learned before quantifiers but could also facilitate the acquisition of quantifiers because of their shared logical form.
<!-- In our model, $\theta$ is left underspecified in the semantics and is inferred in context. -->
<!-- This informal analysis predicts that a child would first acquire the language of generalizations, and from that, carve out more precise meanings for quantified language, consistent with the developmental data. -->

There is a secondary argument for why our semantics presents an easy learning problem.
<!-- The uniform prior over the threshold $\theta$ in the truth-functional semantics (as we have assumed in our model) provides a secondary argument for why acquiring our semantics for generics should be easy. -->
We have assumed the truth functional threshold $\theta$ comes from a uniform distribution over the unit interval $[0, 1]$, which is mathematically equivalent to a *continuously-valued* or *soft semantics* wherein the degree to which the utterance is true is proportional to the degree itself, in this case prevalence $p$: $\int_{0}^{1} \delta_{p > \theta} \diff \theta = p$.
The model for generic interpretation (Eq. \ref{eq:L0}) then becomes: $Lit(p \mid u) \propto p \cdot P(p)$
This *soft semantics* (corresponding intuitively to a meaning like "the higher the probability, the better", or simply "more is better") is perhaps the simplest quantitative semantics one could posit.
The difficulty in acquiring the meaning of quantifiers, then, is a difficulty in recognizing a fixed-threshold semantics as a special case of this *more is better* semantics.
We leave for future work the precise implementation of such an acquisition model.

# Conclusion

It might seem paradoxical that a part of language that is so common in communication and central to learning should be vague. 
Shouldn't speakers and teachers want to express their ideas as crisply as possible?
To the contrary, underspecification can be efficient, given that context can be relied upon to resolve uncertainty in the moment [@Piantadosi2012].
In our work, context takes the form of shared beliefs between speakers and listeners. 
By leveraging this common ground, the language of generalizations provides a powerful way to communicate and learn abstract knowledge, which would otherwise be difficult or costly information to acquire through direct experience.

<!-- The dark side of this flexibility is the potential for miscommunication or deceit: A speaker might assert a generalization that she herself would not accept, conveying a too-strong impression to a naive listener. -->
<!-- Our model predicts this potential particularly for properties which, when present, are widespread in a category; biological properties are believed to have this distribution, but many properties of social categories may as well [@Cimpian2011a; @Cimpian2012b; @Rhodes2012]. -->
<!-- Disagreements are also predicted when interlocutors fail to share background assumptions, which could manifest as differences in the prior distributions on prevalence or what goes into the comparison class. -->
<!-- Core aspects of political disagreement then (e.g., as to whether "Humans cause global warming") could be derived from differences in the estimated causal responsibility of the category in question (e.g., what impacts humans are having), of *other forces* (e.g., plate tectonics), as well as the comparison class of other forces (e.g., what are the alternative causes?).  -->
<!-- This is an important area for future research. -->

Categories are inherently unobservable. 
You cannot see the category \textsc{dog}, only some number of instances of it.
Yet we easily talk about these abstractions, conveying hard-won generalizations to each other and down through generations.
<!-- The theory presented here gives one explanation of how we do so, providing a computational perspective on how we communicate generalizations and how beliefs play a central role in the meaning of words. -->
The theory presented here provides the first computational perspective on how we communicate generalizations, illustrating how beliefs play a central role in understanding the meaning of words.

\newpage

# References

<!-- \setlength{\parindent}{-0.1in}  -->
<!-- \setlength{\leftskip}{0.125in} -->
<!-- \noindent -->

<div id = 'refs'></div>

\newpage

```{r child = 'appendix-cvDerivation.Rmd'}
```

\newpage
```{r child = 'appendix-cueValidity.Rmd'}
```

\newpage

```{r child = 'appendix-bda.Rmd'}
```

\newpage

```{r child = 'appendix-habitualsItems.Rmd'}
```